---
title: "script_assessIPMprediction"
author: "Zheng Yao"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(AICcmodavg)
library(lme4)
library(tidyverse)
library(dplyr)
library(glue)
library(zoo)
library(MASS)
install.packages(c("foreach", "doParallel"))
library(foreach)
library(doParallel)
library(MuMIn)
```

lambda_j \| lambda_g,s_j s= site effect j=B1, ... GSP-LI, n=8\
lambda_j want to estimate the site effect

#load in necessary data. These were made from previous scripts

```{r}
load("/Users/a123/Desktop/TOM/VFT/model assessment/data/model_export_mar_2026.RData")    ## this environment contains vital rate functions plus the raw data used to build them.

```

# global IPM

```{r}
#size bounds for IPM
TBF_long$logsize0 <- (log(TBF_long$size0+0.1))
TBF_long$s.logsize0 <- scale(TBF_long$logsize0) 
TBF_long$s.TBF <- scale(TBF_long$TBF) 
TBF_long$s.TSF <- scale(TBF_long$TSF) 
TBF_long$sq.TSF <- (TBF_long$TSF)^2
TBF_long$s.sq.TSF <- scale(TBF_long$sq.TSF) 
IPM_bounds <- range(TBF_long$logsize0, na.rm = TRUE)  # our IPM will be based on logsize as a size metric. 
no_bins <- 100
bin_edges <- seq(IPM_bounds[1], IPM_bounds[2], length.out= no_bins+1)
bin_mids <- zoo::rollmean(bin_edges,2)

log.fr <- log(1.1)
# define mean and sd of seedling sizes
rows_with_new <- TBF_long[grepl("\\bnew", TBF_long$comm1), ]
rows_with_new <- rows_with_new[!grepl("no new", rows_with_new$comm1), ] # delete "no new"
rows_with_new <- rows_with_new[!grepl("no news", rows_with_new$comm1), ] # delete "no new" 
rows_with_new <- rows_with_new[!grepl("not possible to see new plants", rows_with_new$comm1), ]
rows_with_new <- rows_with_new[!grepl("No news", rows_with_new$comm1), ]
rows_with_new <- rows_with_new[!grepl("new lvs", rows_with_new$comm1), ]
rows_with_new <- rows_with_new[!grepl("new nail", rows_with_new$comm1), ]

mean_seedling_logsize <- mean(rows_with_new$logsize1, na.rm=T) # get the mean logsize of seedings   ### the mean seedling size is very large
var_seedling_logsize <- var(rows_with_new$logsize1,na.rm=T)# get the var in logsize of seedings  

### build empty matrices to store IPM output. Here, I'm interested in the effect of fire on lambda, so I'm setting average conditions for all other climate variables
possible_scenarios <- 
  as.data.frame(matrix(NA, nrow= 22, ncol = 28))
names(possible_scenarios) <- c("FRI","s.TBF","s.T_LA", "s.T_LH",
  "s.sq.T_LH",
  "s.sq.P_LA",
  "s.T_LC",
  "s.T_LD",
  "s.P_LD",
  "s.sq.T_LA",
  "s.P_LA",
  "s.P_LW",
  "s.P_LH",
  "s.T_RA",
  "s.sq.T_RA",
  "s.P_RA",
  "s.T_RH",
  "s.P_RH",
  "s.T_RC",
  "s.T_RW",
  "s.T_RD",
  "s.sq.T_RC",
  "s.P_RW",
  "s.P_RD",
  "s.sq.T_RH",
  "s.sq.P_RA",
  "s.sq.P_RW",
  "s.sq.P_RD")


possible_scenarios$FRI <- rep(1:11, 2) # we want FRI to range from 1-11
possible_scenarios$s.TBF <- scale(rep(c(0, 10), each= 11) ,# raw TBF values
                                  center= attr(TBF_long$s.TBF,"scaled:center"), 
                                  scale= attr(TBF_long$s.TBF,"scaled:scale")) # for each set of FRI's, we want to set a short TBF (of 3) and a long TBF (of 11)

possible_scenarios[, c(  "s.T_LA",  "s.T_LH",  "s.sq.T_LH",  "s.sq.P_LA",  "s.T_LC",  "s.T_LD",  "s.P_LD",  "s.sq.T_LA",  "s.P_LA",  
                         "s.P_LW",  "s.P_LH",  "s.T_RA",  "s.sq.T_RA",  "s.P_RA",  "s.T_RH",  "s.P_RH",  "s.T_RC",  "s.T_RW",  "s.T_RD",
                         "s.sq.T_RC","s.P_RW","s.P_RD", "s.sq.T_RH",  "s.sq.P_RA",  "s.sq.P_RW",  "s.sq.P_RD")] <- 0 # setting all climate variable to their mean

my_sites <- unique(TBF_long$site) 
no_reps <- 1000 # or somethign similar

########### extract model components for later predict() ###############
sur_model_terms <- terms(survival_mod)
sur_term <- attr(sur_model_terms, "term.labels")
sur_term <- reformulate(sur_term)

gr_model_terms <- terms(growth_mod)
gr_term <- attr(gr_model_terms, "term.labels")
gr_term <- reformulate(gr_term)

prep_model_terms <- terms(prep_mod)
prep_term <- attr(prep_model_terms, "term.labels")
prep_term <- reformulate(prep_term)

crep_model_terms <- terms(crep_mod)
crep_term <- attr(crep_model_terms, "term.labels")
crep_term <- reformulate(crep_term)

vargrowth_model_terms <- terms(vargrowth_mod)
vargrowth_term <- attr(vargrowth_model_terms, "term.labels")
vargrowth_term <- reformulate(vargrowth_term)

rec_model_terms <- terms(recruit_mod)
rec_term <- attr(rec_model_terms, "term.labels")
rec_term <- reformulate(rec_term)

##################################################################
###################### initiate IPM  ##############################
##################################################################
all_pop_growth_rates <- matrix(data= NA, nrow= dim(possible_scenarios)[1], ncol= no_reps)
n_cores <- detectCores()
cl <- makeCluster(n_cores - 1)
registerDoParallel(cl)
j_pop_gr_rate <- rep(NA, dim(possible_scenarios)[1]) ### changed to no_scenarios

all_pop_growth_rates <- foreach(
  j = 1:no_reps,
  .combine = cbind,
  .packages = c("MASS", "lme4", "popbio", "dplyr")
) %dopar% {
  rsur_coefs <- coefficients(survival_mod) # first, assign all "randomy selected" coefficients to the mean estimated coefficeints
  rsur_coefs[!names(coefficients(survival_mod)) %in% paste("site", my_sites, sep="")] <- MASS::mvrnorm(mu= coefficients(survival_mod), Sigma = vcov(survival_mod))[!names(coefficients(survival_mod)) %in% paste("site", my_sites, sep="")]
  #rsur_coefs <- lme4::fixef(survival_mod)
  #rsur_coefs <-  MASS::mvrnorm(mu= lme4::fixef(survival_mod) , Sigma = vcov(survival_mod))
  rgr_coefs <- lme4::fixef(growth_mod)
  rgr_coefs <-  MASS::mvrnorm(mu= lme4::fixef(growth_mod) , Sigma = vcov(growth_mod)) # no site effect in this model
  rprep_coefs <- coefficients(prep_mod)
  rprep_coefs[!names(coefficients(prep_mod)) %in% paste("site", my_sites, sep="")] <- MASS::mvrnorm(mu= coefficients(prep_mod), Sigma = vcov(prep_mod))[!names(coefficients(prep_mod)) %in% paste("site", my_sites, sep="")]
  rcrep_coefs <- lme4::fixef(crep_mod)
  rcrep_coefs <-  MASS::mvrnorm(mu= lme4::fixef(crep_mod) , Sigma = vcov(crep_mod)) # no site effect in this model
  rvar_coefs <- lme4::fixef(vargrowth_mod)
  rvar_coefs <-  MASS::mvrnorm(mu= lme4::fixef(vargrowth_mod) , Sigma = vcov(vargrowth_mod)) # no site effect in this model
  rrec_coefs <- lme4::fixef(recruit_mod)
  rrec_coefs <-  MASS::mvrnorm(mu= lme4::fixef(recruit_mod) , Sigma = vcov(recruit_mod)) # no site effect in this model
  #rrec_coefs <-  MASS::mvrnorm(mu= lme4::fixef(recruit_mod) , Sigma = vcov(recruit_mod)) 
  
  for (i in 1: dim(possible_scenarios)[1]) {
    
    i_scenario <- possible_scenarios[i,] # which combination of FRI, TBF, and climate do I want to use
    
    TSFs <- 0:(i_scenario$FRI-1) # TSF range from 0 to (FRI-1) such that an FRI of 1 will always have TSF= 0
    i_scenario <- i_scenario[rep(seq_len(nrow(i_scenario)), length(TSFs)), ]
    i_scenario$TSF <- TSFs
    i_scenario$s.TSF <- scale(TSFs, center= attr(TBF_long$s.TSF,"scaled:center"), 
                              scale= attr(TBF_long$s.TSF,"scaled:scale"))
    i_scenario$s.sq.TSF <- scale(TSFs^2, center= attr(TBF_long$s.sq.TSF,"scaled:center"), 
                                 scale= attr(TBF_long$s.sq.TSF,"scaled:scale"))
    
    mx_prod <- diag(no_bins+1) # setting up an identity matrix before you enter the loop
    
    for (ii in TSFs){
      i_ii_scenario <- i_scenario[which(i_scenario$TSF==ii),]
      data_for_prediction <- i_ii_scenario[rep(seq_len(nrow(i_ii_scenario)), length(bin_mids)), ]
      data_for_prediction$s.logsize0 <-   scale(bin_mids, center= attr(TBF_long$s.logsize0,"scaled:center"), 
                                                scale= attr(TBF_long$s.logsize0,"scaled:scale"))
      data_for_prediction_rec <- cbind(i_ii_scenario)
      
      predicted_sur <- predicted_gr <-predicted_pr <- predicted_cr <- predicted_vg <- matrix(data=NA, nrow= length(my_sites), ncol= length(bin_mids))
      predicted_rec <- rep(NA, length(my_sites))
      
      ### make empty mxes before entering the site loop
      
      predicted_sur_allsites <-predicted_gr_allsites<-predicted_pr_allsites <-predicted_cr_allsites<-predicted_vg_allsites <-matrix(NA, nrow = length(my_sites), ncol = no_bins, dimnames = list(my_sites, NULL))
      predicted_rec_allsites <- matrix(NA, nrow = length(my_sites), ncol = 1, dimnames = list(my_sites, NULL))
      
      
      for (s in my_sites){ #looping through the sites
        s_data_for_prediction <- data_for_prediction
        s_data_for_prediction_rec <- data_for_prediction_rec
        s_data_for_prediction$site <- s
        s_data_for_prediction_rec$site <- s
        
        s_data_for_prediction$site     <- factor(s_data_for_prediction$site,  levels = my_sites)   ## it is important to add the levels=my_sites here to ensure that the model matrix is built correctly
        s_data_for_prediction_rec$site <- factor(s_data_for_prediction_rec$site, levels = my_sites)
        
        s_data_for_prediction <- s_data_for_prediction %>% mutate(across(c("s.logsize0", "s.sq.TSF", "s.TSF", "s.TBF"), as.numeric))   ### change col type to numeric to avoid matrix col type
        # Zheng-- you will have to change the predict() call in the below to be a custom predict function that uses the random coefficients generated earlier in the code (e.g., rsur_coefs)
        
        ### build model matrices and use mx multiplication to obtain predicted values. Don't forget to transform if needed!!!!!!!!!
        
        m_sur <- model.matrix(sur_term ,data=s_data_for_prediction)
        p2_sur <- rsur_coefs %*% t(m_sur) # check on whether this coef(survival_mod is a column or a row!!!!!!!!
        p2_sur <- plogis(p2_sur)   ## transform to binomial 
        predicted_sur_allsites[s,] <- p2_sur
        
        m_gr <- model.matrix(gr_term, data=s_data_for_prediction)
        p2_gr <- rgr_coefs %*% t(m_gr) # check on whether this coef(survival_mod is a column or a row!!!!!!!!
        predicted_gr_allsites[s,] <- p2_gr
        
        m_prep <- model.matrix(prep_term,data=s_data_for_prediction)
        p2_prep <- rprep_coefs %*% t(m_prep) # check on whether this coef(survival_mod is a column or a row!!!!!!!!
        p2_prep <- plogis(p2_prep)   ## transform to binomial 
        predicted_pr_allsites[s,] <- p2_prep
        
        m_crep <- model.matrix(crep_term  ,data=s_data_for_prediction)
        p2_crep <- rcrep_coefs %*% t(m_crep) # check on whether this coef(survival_mod is a column or a row!!!!!!!!
        predicted_cr_allsites[s,] <- p2_crep
        
        m_vg <- model.matrix(vargrowth_term,data=s_data_for_prediction)
        p2_vg <- rvar_coefs %*% t(m_vg) # 
        predicted_vg_allsites[s,] <- exp(p2_vg)  
        
        m_rec <- model.matrix(rec_term , data=s_data_for_prediction_rec)
        p2_rec <- rrec_coefs %*% t(m_rec) # check on whether this coef(survival_mod is a column or a row!!!!!!!!
        predicted_rec_allsites[s,1] <- p2_rec*log(1.1)            #### account for composite rec variable (rec/log.fr)
        
        
        # for mixed models, the above functions will predict site-specific results (i.e., will incorporate random effects)
        rm(s_data_for_prediction) # this part removes the site-specific dataframe used in this loop, so you do not accidentaly use the previous sites' dataframe instead of the current one-- in case there was an error somehow
      }
      predicted_sur <- apply(predicted_sur_allsites, MARGIN= 2, FUN= mean) # average size-specific vital rates across sites
      predicted_gr <- apply(predicted_gr_allsites, MARGIN= 2, FUN= mean)
      predicted_pr <- apply(predicted_pr_allsites, MARGIN= 2, FUN= mean)
      predicted_cr <- apply(predicted_cr_allsites, MARGIN= 2, FUN= mean)
      predicted_vg <- apply(predicted_vg_allsites, MARGIN= 2, FUN= mean)
      predicted_vg[which(predicted_vg<0)] <- 0                 #### to prevent negative variance
      predicted_rec <- mean(predicted_rec_allsites)
      
      gmx <- matrix(NA, length(bin_mids),length(bin_mids)) ## create empty gmx before entering the no_bins loop
      
      # make the TSF-specific matrix
      for (ss in 1:no_bins) {
        growcdf <- pnorm(bin_edges,predicted_gr[ss],sqrt(predicted_vg[ss])) # Yes. Vargrowth=(pred - obs)^2
        grows <- growcdf[2:length(bin_edges)]-growcdf[1:(length(bin_edges)-1)]
        if(sum(grows)>0){grows <- grows/sum(grows)
        gmx[,ss] <- grows
        } else if (sum(grows)==0 & grow_vals[ss]< min(bin_edges)) { 
          gmx[,ss] <- c(1, rep(0, length(binmids)-1))  
        } else if (sum(grows)==0 & grow_vals[ss]> max(bin_edges)) { 
          gmx[,ss] <- c(rep(0, length(binmids)-1), 1)  } else { # deals with complete eviction-- OUT OF THE BOUNDS  of the kernel
            gmx[,ss] <- NA} 
        # this if statement breaks the code (puts NA's into the matrix) if the sum of the PDF is zero (which happens if all the probability is outside of the size bounds)
      } # end ss loop
      
      # make the surv*growth mx
      survgmx <- gmx*t(matrix( rep(predicted_sur,no_bins),no_bins)) # Yes, this is doing element-wise multiplication 
      reprow <-  predicted_sur*predicted_pr*predicted_cr # prob that you survive, reproduce, then number fruits|reproduce; this multiplcation does it element by element
      
      mx <- matrix(0, no_bins+1, no_bins+1)
      mx[2:(no_bins+1), 2:(no_bins+1)] = survgmx
      mx[1,2:(no_bins+1)] = reprow # FRUIT ROW
      
      sdlng_cdf <- pnorm(bin_edges,mean_seedling_logsize,sqrt(var_seedling_logsize)) 
      sdlng_pdf <- sdlng_cdf[2:length(bin_edges)]-sdlng_cdf[1:(length(bin_edges)-1)]
      if(sum(sdlng_pdf)>0){
        sdlng_pdf <- sdlng_pdf/sum(sdlng_pdf)} else {
          sdlng_pdf <- NA} 
      
      mx[2:(no_bins+1),1]  = predicted_rec *sdlng_pdf 
      
      mx_prod <- mx_prod %*% mx # in the first iteration of the loop, mx just becomes mx_prod
      
    } # closes ii (TSFs) loop
    
    j_pop_gr_rate[i] <- (popbio::lambda(mx_prod))^(1/length(TSFs)) 
    # multiply them together to get a matrix, etc. 
  }
    j_pop_gr_rate
  
}


######################################
###################################### global IPM results, rep=1000 
######################################

all_pop_growth_rates <- cbind(possible_scenarios, all_pop_growth_rates)
names(all_pop_growth_rates) <- c(names(possible_scenarios), paste("rep", 1:no_reps, sep = ""))

rep_cols <- grep("^rep", names(all_pop_growth_rates), value = TRUE)

mean_lambda_across_scenarios <- all_pop_growth_rates %>%
  rowwise() %>%
  mutate(
    mean_lambda = mean(c_across(all_of(rep_cols)), na.rm = TRUE),
    lower_bound = quantile(c_across(all_of(rep_cols)), 0.025, na.rm = TRUE),
    upper_bound = quantile(c_across(all_of(rep_cols)), 0.975, na.rm = TRUE)
  ) %>%
  dplyr::select(FRI,s.TBF,mean_lambda,lower_bound, upper_bound) %>%
  mutate(TBF=NA) %>%
  ungroup()

mean_lambda_across_scenarios$TBF[1:11] <- "Short TBF"
mean_lambda_across_scenarios$TBF[12:22] <- "Long TBF"

##########. graph 
##########

plt1 <- ggplot(data= mean_lambda_across_scenarios, aes(x= FRI, y= mean_lambda)) +
  geom_line(aes(color= TBF)) + 
  geom_ribbon(aes(ymin = mean_lambda_across_scenarios$lower_bound, ymax = mean_lambda_across_scenarios$upper_bound, fill= TBF), alpha = 0.1) + 
  #ylim(0,1) +
  #xlim(c(0, 12)) +
  labs(y= "Growth Rate", x = "Fire Return Interval",color = NULL, fill = NULL) +
  theme_bw() + theme(legend.position='none') +
  theme(legend.position = "right") +
  scale_x_continuous(breaks = 0:10) +
  theme(text = element_text(size = 16)) 
plt1



```

## use global IPM to predict regions

#### Croatan (history of fire suppression)

```{r}

all_pop_growth_rates_Cro <- matrix(data= NA, nrow= dim(possible_scenarios)[1], ncol= no_reps)
n_cores <- detectCores()
cl <- makeCluster(n_cores - 1)
registerDoParallel(cl)
j_pop_gr_rate_Cro <- rep(NA, dim(possible_scenarios)[1]) ### changed to no_scenarios

all_pop_growth_rates_Cro <- foreach(
  j = 1:no_reps,
  .combine = cbind,
  .packages = c("MASS", "lme4", "popbio", "dplyr")
) %dopar% {
  rsur_coefs <- coefficients(survival_mod) # first, assign all "randomy selected" coefficients to the mean estimated coefficeints
  rsur_coefs[!names(coefficients(survival_mod)) %in% paste("site", my_sites, sep="")] <- MASS::mvrnorm(mu= coefficients(survival_mod), Sigma = vcov(survival_mod))[!names(coefficients(survival_mod)) %in% paste("site", my_sites, sep="")]
  #rsur_coefs <- lme4::fixef(survival_mod)
  #rsur_coefs <-  MASS::mvrnorm(mu= lme4::fixef(survival_mod) , Sigma = vcov(survival_mod))
  rgr_coefs <- lme4::fixef(growth_mod)
  rgr_coefs <-  MASS::mvrnorm(mu= lme4::fixef(growth_mod) , Sigma = vcov(growth_mod)) # no site effect in this model
  rprep_coefs <- coefficients(prep_mod)
  rprep_coefs[!names(coefficients(prep_mod)) %in% paste("site", my_sites, sep="")] <- MASS::mvrnorm(mu= coefficients(prep_mod), Sigma = vcov(prep_mod))[!names(coefficients(prep_mod)) %in% paste("site", my_sites, sep="")]
  rcrep_coefs <- lme4::fixef(crep_mod)
  rcrep_coefs <-  MASS::mvrnorm(mu= lme4::fixef(crep_mod) , Sigma = vcov(crep_mod)) # no site effect in this model
  rvar_coefs <- lme4::fixef(vargrowth_mod)
  rvar_coefs <-  MASS::mvrnorm(mu= lme4::fixef(vargrowth_mod) , Sigma = vcov(vargrowth_mod)) # no site effect in this model
  rrec_coefs <- lme4::fixef(recruit_mod)
  rrec_coefs <-  MASS::mvrnorm(mu= lme4::fixef(recruit_mod) , Sigma = vcov(recruit_mod)) # no site effect in this model
  #rrec_coefs <-  MASS::mvrnorm(mu= lme4::fixef(recruit_mod) , Sigma = vcov(recruit_mod)) 
  
  for (i in 1: dim(possible_scenarios)[1]) {
    
    i_scenario <- possible_scenarios[i,] # which combination of FRI, TBF, and climate do I want to use
    
    TSFs <- 0:(i_scenario$FRI-1) # TSF range from 0 to (FRI-1) such that an FRI of 1 will always have TSF= 0
    i_scenario <- i_scenario[rep(seq_len(nrow(i_scenario)), length(TSFs)), ]
    i_scenario$TSF <- TSFs
    i_scenario$s.TSF <- scale(TSFs, center= attr(TBF_long$s.TSF,"scaled:center"), 
                              scale= attr(TBF_long$s.TSF,"scaled:scale"))
    i_scenario$s.sq.TSF <- scale(TSFs^2, center= attr(TBF_long$s.sq.TSF,"scaled:center"), 
                                 scale= attr(TBF_long$s.sq.TSF,"scaled:scale"))
    
    mx_prod <- diag(no_bins+1) # setting up an identity matrix before you enter the loop
    
    for (ii in TSFs){
      i_ii_scenario <- i_scenario[which(i_scenario$TSF==ii),]
      data_for_prediction <- i_ii_scenario[rep(seq_len(nrow(i_ii_scenario)), length(bin_mids)), ]
      data_for_prediction$s.logsize0 <-   scale(bin_mids, center= attr(TBF_long$s.logsize0,"scaled:center"), 
                                                scale= attr(TBF_long$s.logsize0,"scaled:scale"))
      data_for_prediction_rec <- cbind(i_ii_scenario)
      
      predicted_sur <- predicted_gr <-predicted_pr <- predicted_cr <- predicted_vg <- matrix(data=NA, nrow= length(my_sites), ncol= length(bin_mids))
      predicted_rec <- rep(NA, length(my_sites))
      
      ### make empty mxes before entering the site loop
      
      predicted_sur_allsites <-predicted_gr_allsites<-predicted_pr_allsites <-predicted_cr_allsites<-predicted_vg_allsites <-matrix(NA, nrow = length(my_sites), ncol = no_bins, dimnames = list(my_sites, NULL))
      predicted_rec_allsites <- matrix(NA, nrow = length(my_sites), ncol = 1, dimnames = list(my_sites, NULL))
      
      
      for (s in my_sites[3:4]){ #looping through the sites
        s_data_for_prediction <- data_for_prediction
        s_data_for_prediction_rec <- data_for_prediction_rec
        s_data_for_prediction$site <- s
        s_data_for_prediction_rec$site <- s
        
        s_data_for_prediction$site     <- factor(s_data_for_prediction$site,  levels = my_sites)   ## it is important to add the levels=my_sites here to ensure that the model matrix is built correctly
        s_data_for_prediction_rec$site <- factor(s_data_for_prediction_rec$site, levels = my_sites)
        
        s_data_for_prediction <- s_data_for_prediction %>% mutate(across(c("s.logsize0", "s.sq.TSF", "s.TSF", "s.TBF"), as.numeric))   ### change col type to numeric to avoid matrix col type
        # Zheng-- you will have to change the predict() call in the below to be a custom predict function that uses the random coefficients generated earlier in the code (e.g., rsur_coefs)
        
        ### build model matrices and use mx multiplication to obtain predicted values. Don't forget to transform if needed!!!!!!!!!
        
        m_sur <- model.matrix(sur_term ,data=s_data_for_prediction)
        p2_sur <- rsur_coefs %*% t(m_sur) # check on whether this coef(survival_mod is a column or a row!!!!!!!!
        p2_sur <- plogis(p2_sur)   ## transform to binomial 
        predicted_sur_allsites[s,] <- p2_sur
        
        m_gr <- model.matrix(gr_term, data=s_data_for_prediction)
        p2_gr <- rgr_coefs %*% t(m_gr) # check on whether this coef(survival_mod is a column or a row!!!!!!!!
        predicted_gr_allsites[s,] <- p2_gr
        
        m_prep <- model.matrix(prep_term,data=s_data_for_prediction)
        p2_prep <- rprep_coefs %*% t(m_prep) # check on whether this coef(survival_mod is a column or a row!!!!!!!!
        p2_prep <- plogis(p2_prep)   ## transform to binomial 
        predicted_pr_allsites[s,] <- p2_prep
        
        m_crep <- model.matrix(crep_term  ,data=s_data_for_prediction)
        p2_crep <- rcrep_coefs %*% t(m_crep) # check on whether this coef(survival_mod is a column or a row!!!!!!!!
        predicted_cr_allsites[s,] <- p2_crep
        
        m_vg <- model.matrix(vargrowth_term,data=s_data_for_prediction)
        p2_vg <- rvar_coefs %*% t(m_vg) # 
        predicted_vg_allsites[s,] <- exp(p2_vg)  
        
        m_rec <- model.matrix(rec_term , data=s_data_for_prediction_rec)
        p2_rec <- rrec_coefs %*% t(m_rec) # check on whether this coef(survival_mod is a column or a row!!!!!!!!
        predicted_rec_allsites[s,1] <- p2_rec*log(1.1)            #### account for composite rec variable (rec/log.fr)
        
        
        # for mixed models, the above functions will predict site-specific results (i.e., will incorporate random effects)
        rm(s_data_for_prediction) # this part removes the site-specific dataframe used in this loop, so you do not accidentaly use the previous sites' dataframe instead of the current one-- in case there was an error somehow
      }
      predicted_sur <- apply(predicted_sur_allsites, MARGIN= 2, FUN= mean,na.rm = TRUE) # average size-specific vital rates across sites
      predicted_gr <- apply(predicted_gr_allsites, MARGIN= 2, FUN= mean,na.rm = TRUE)
      predicted_pr <- apply(predicted_pr_allsites, MARGIN= 2, FUN= mean,na.rm = TRUE)
      predicted_cr <- apply(predicted_cr_allsites, MARGIN= 2, FUN= mean,na.rm = TRUE)
      predicted_vg <- apply(predicted_vg_allsites, MARGIN= 2, FUN= mean,na.rm = TRUE)
      predicted_vg[which(predicted_vg<0)] <- 0                 #### to prevent negative variance
      predicted_rec <- mean(predicted_rec_allsites,na.rm = TRUE)
      
      gmx <- matrix(NA, length(bin_mids),length(bin_mids)) ## create empty gmx before entering the no_bins loop
      
      # make the TSF-specific matrix
      for (ss in 1:no_bins) {
        growcdf <- pnorm(bin_edges,predicted_gr[ss],sqrt(predicted_vg[ss])) # Yes. Vargrowth=(pred - obs)^2
        grows <- growcdf[2:length(bin_edges)]-growcdf[1:(length(bin_edges)-1)]
        if(sum(grows)>0){grows <- grows/sum(grows)
        gmx[,ss] <- grows
        } else if (sum(grows)==0 & grow_vals[ss]< min(bin_edges)) { 
          gmx[,ss] <- c(1, rep(0, length(binmids)-1))  
        } else if (sum(grows)==0 & grow_vals[ss]> max(bin_edges)) { 
          gmx[,ss] <- c(rep(0, length(binmids)-1), 1)  } else { # deals with complete eviction-- OUT OF THE BOUNDS  of the kernel
            gmx[,ss] <- NA} 
        # this if statement breaks the code (puts NA's into the matrix) if the sum of the PDF is zero (which happens if all the probability is outside of the size bounds)
      } # end ss loop
      
      # make the surv*growth mx
      survgmx <- gmx*t(matrix( rep(predicted_sur,no_bins),no_bins)) # Yes, this is doing element-wise multiplication 
      reprow <-  predicted_sur*predicted_pr*predicted_cr # prob that you survive, reproduce, then number fruits|reproduce; this multiplcation does it element by element
      
      mx <- matrix(0, no_bins+1, no_bins+1)
      mx[2:(no_bins+1), 2:(no_bins+1)] = survgmx
      mx[1,2:(no_bins+1)] = reprow # FRUIT ROW
      
      sdlng_cdf <- pnorm(bin_edges,mean_seedling_logsize,sqrt(var_seedling_logsize)) 
      sdlng_pdf <- sdlng_cdf[2:length(bin_edges)]-sdlng_cdf[1:(length(bin_edges)-1)]
      if(sum(sdlng_pdf)>0){
        sdlng_pdf <- sdlng_pdf/sum(sdlng_pdf)} else {
          sdlng_pdf <- NA} 
      
      mx[2:(no_bins+1),1]  = predicted_rec *sdlng_pdf 
      
      mx_prod <- mx_prod %*% mx # in the first iteration of the loop, mx just becomes mx_prod
      
    } # closes ii (TSFs) loop
    
    j_pop_gr_rate_Cro[i] <- (popbio::lambda(mx_prod))^(1/length(TSFs)) 
    # multiply them together to get a matrix, etc. 
  }
    j_pop_gr_rate_Cro
}


all_pop_growth_rates_Cro <- cbind(possible_scenarios, all_pop_growth_rates_Cro)
names(all_pop_growth_rates_Cro) <- c(names(possible_scenarios), paste("rep", 1:no_reps, sep = ""))
rep_cols <- grep("^rep", names(all_pop_growth_rates_Cro), value = TRUE)

mean_lambda_across_scenarios_Cro <- all_pop_growth_rates_Cro %>%
  rowwise() %>%
  mutate(
    mean_lambda = mean(c_across(all_of(rep_cols)), na.rm = TRUE),
    lower_bound = quantile(c_across(all_of(rep_cols)), 0.025, na.rm = TRUE),
    upper_bound = quantile(c_across(all_of(rep_cols)), 0.975, na.rm = TRUE)
  ) %>%
  dplyr::select(FRI,s.TBF,mean_lambda,lower_bound, upper_bound) %>%
  mutate(TBF=NA) %>%
  ungroup()

mean_lambda_across_scenarios_Cro$TBF[1:11] <- "Short TBF"
mean_lambda_across_scenarios_Cro$TBF[12:22] <- "Long TBF"

##########. graph 
##########

plt2 <- ggplot(data= mean_lambda_across_scenarios_Cro, aes(x= FRI, y= mean_lambda)) +
  geom_line(aes(color= TBF)) + 
  geom_ribbon(aes(ymin = mean_lambda_across_scenarios_Cro$lower_bound, ymax = mean_lambda_across_scenarios_Cro$upper_bound, fill= TBF), alpha = 0.1) + 
  #ylim(0,1) +
  #xlim(c(0, 12)) +
  labs(y= "Growth Rate", x = "Fire Return Interval",color = NULL, fill = NULL) +
  theme_bw() + theme(legend.position='none') +
  theme(legend.position = "right") +
  scale_x_continuous(breaks = 0:10) +
  theme(text = element_text(size = 16)) 
plt2


```

#### Green Swamp (no history of fire suppression)

```{r}
all_pop_growth_rates_GSP <- matrix(data= NA, nrow= dim(possible_scenarios)[1], ncol= no_reps)
n_cores <- detectCores()
cl <- makeCluster(n_cores - 1)
registerDoParallel(cl)
j_pop_gr_rate_GSP <- rep(NA, dim(possible_scenarios)[1]) ### changed to no_scenarios

all_pop_growth_rates_GSP <- foreach(
  j = 1:no_reps,
  .combine = cbind,
  .packages = c("MASS", "lme4", "popbio", "dplyr")
) %dopar% {
  rsur_coefs <- coefficients(survival_mod) # first, assign all "randomy selected" coefficients to the mean estimated coefficeints
  rsur_coefs[!names(coefficients(survival_mod)) %in% paste("site", my_sites, sep="")] <- MASS::mvrnorm(mu= coefficients(survival_mod), Sigma = vcov(survival_mod))[!names(coefficients(survival_mod)) %in% paste("site", my_sites, sep="")]
  #rsur_coefs <- lme4::fixef(survival_mod)
  #rsur_coefs <-  MASS::mvrnorm(mu= lme4::fixef(survival_mod) , Sigma = vcov(survival_mod))
  rgr_coefs <- lme4::fixef(growth_mod)
  rgr_coefs <-  MASS::mvrnorm(mu= lme4::fixef(growth_mod) , Sigma = vcov(growth_mod)) # no site effect in this model
  rprep_coefs <- coefficients(prep_mod)
  rprep_coefs[!names(coefficients(prep_mod)) %in% paste("site", my_sites, sep="")] <- MASS::mvrnorm(mu= coefficients(prep_mod), Sigma = vcov(prep_mod))[!names(coefficients(prep_mod)) %in% paste("site", my_sites, sep="")]
  rcrep_coefs <- lme4::fixef(crep_mod)
  rcrep_coefs <-  MASS::mvrnorm(mu= lme4::fixef(crep_mod) , Sigma = vcov(crep_mod)) # no site effect in this model
  rvar_coefs <- lme4::fixef(vargrowth_mod)
  rvar_coefs <-  MASS::mvrnorm(mu= lme4::fixef(vargrowth_mod) , Sigma = vcov(vargrowth_mod)) # no site effect in this model
  rrec_coefs <- lme4::fixef(recruit_mod)
  rrec_coefs <-  MASS::mvrnorm(mu= lme4::fixef(recruit_mod) , Sigma = vcov(recruit_mod)) # no site effect in this model
  #rrec_coefs <-  MASS::mvrnorm(mu= lme4::fixef(recruit_mod) , Sigma = vcov(recruit_mod)) 
  
  for (i in 1: dim(possible_scenarios)[1]) {
    
    i_scenario <- possible_scenarios[i,] # which combination of FRI, TBF, and climate do I want to use
    
    TSFs <- 0:(i_scenario$FRI-1) # TSF range from 0 to (FRI-1) such that an FRI of 1 will always have TSF= 0
    i_scenario <- i_scenario[rep(seq_len(nrow(i_scenario)), length(TSFs)), ]
    i_scenario$TSF <- TSFs
    i_scenario$s.TSF <- scale(TSFs, center= attr(TBF_long$s.TSF,"scaled:center"), 
                              scale= attr(TBF_long$s.TSF,"scaled:scale"))
    i_scenario$s.sq.TSF <- scale(TSFs^2, center= attr(TBF_long$s.sq.TSF,"scaled:center"), 
                                 scale= attr(TBF_long$s.sq.TSF,"scaled:scale"))
    
    mx_prod <- diag(no_bins+1) # setting up an identity matrix before you enter the loop
    
    for (ii in TSFs){
      i_ii_scenario <- i_scenario[which(i_scenario$TSF==ii),]
      data_for_prediction <- i_ii_scenario[rep(seq_len(nrow(i_ii_scenario)), length(bin_mids)), ]
      data_for_prediction$s.logsize0 <-   scale(bin_mids, center= attr(TBF_long$s.logsize0,"scaled:center"), 
                                                scale= attr(TBF_long$s.logsize0,"scaled:scale"))
      data_for_prediction_rec <- cbind(i_ii_scenario)
      
      predicted_sur <- predicted_gr <-predicted_pr <- predicted_cr <- predicted_vg <- matrix(data=NA, nrow= length(my_sites), ncol= length(bin_mids))
      predicted_rec <- rep(NA, length(my_sites))
      
      ### make empty mxes before entering the site loop
      
      predicted_sur_allsites <-predicted_gr_allsites<-predicted_pr_allsites <-predicted_cr_allsites<-predicted_vg_allsites <-matrix(NA, nrow = length(my_sites), ncol = no_bins, dimnames = list(my_sites, NULL))
      predicted_rec_allsites <- matrix(NA, nrow = length(my_sites), ncol = 1, dimnames = list(my_sites, NULL))
      
      
      for (s in my_sites[5:6]){ #looping through the sites
        s_data_for_prediction <- data_for_prediction
        s_data_for_prediction_rec <- data_for_prediction_rec
        s_data_for_prediction$site <- s
        s_data_for_prediction_rec$site <- s
        
        s_data_for_prediction$site     <- factor(s_data_for_prediction$site,  levels = my_sites)   ## it is important to add the levels=my_sites here to ensure that the model matrix is built correctly
        s_data_for_prediction_rec$site <- factor(s_data_for_prediction_rec$site, levels = my_sites)
        
        s_data_for_prediction <- s_data_for_prediction %>% mutate(across(c("s.logsize0", "s.sq.TSF", "s.TSF", "s.TBF"), as.numeric))   ### change col type to numeric to avoid matrix col type
        # Zheng-- you will have to change the predict() call in the below to be a custom predict function that uses the random coefficients generated earlier in the code (e.g., rsur_coefs)
        
        ### build model matrices and use mx multiplication to obtain predicted values. Don't forget to transform if needed!!!!!!!!!
        
        m_sur <- model.matrix(sur_term ,data=s_data_for_prediction)
        p2_sur <- rsur_coefs %*% t(m_sur) # check on whether this coef(survival_mod is a column or a row!!!!!!!!
        p2_sur <- plogis(p2_sur)   ## transform to binomial 
        predicted_sur_allsites[s,] <- p2_sur
        
        m_gr <- model.matrix(gr_term, data=s_data_for_prediction)
        p2_gr <- rgr_coefs %*% t(m_gr) # check on whether this coef(survival_mod is a column or a row!!!!!!!!
        predicted_gr_allsites[s,] <- p2_gr
        
        m_prep <- model.matrix(prep_term,data=s_data_for_prediction)
        p2_prep <- rprep_coefs %*% t(m_prep) # check on whether this coef(survival_mod is a column or a row!!!!!!!!
        p2_prep <- plogis(p2_prep)   ## transform to binomial 
        predicted_pr_allsites[s,] <- p2_prep
        
        m_crep <- model.matrix(crep_term  ,data=s_data_for_prediction)
        p2_crep <- rcrep_coefs %*% t(m_crep) # check on whether this coef(survival_mod is a column or a row!!!!!!!!
        predicted_cr_allsites[s,] <- p2_crep
        
        m_vg <- model.matrix(vargrowth_term,data=s_data_for_prediction)
        p2_vg <- rvar_coefs %*% t(m_vg) # 
        predicted_vg_allsites[s,] <- exp(p2_vg)  
        
        m_rec <- model.matrix(rec_term , data=s_data_for_prediction_rec)
        p2_rec <- rrec_coefs %*% t(m_rec) # check on whether this coef(survival_mod is a column or a row!!!!!!!!
        predicted_rec_allsites[s,1] <- p2_rec*log(1.1)            #### account for composite rec variable (rec/log.fr)
        
        
        # for mixed models, the above functions will predict site-specific results (i.e., will incorporate random effects)
        rm(s_data_for_prediction) # this part removes the site-specific dataframe used in this loop, so you do not accidentaly use the previous sites' dataframe instead of the current one-- in case there was an error somehow
      }
      predicted_sur <- apply(predicted_sur_allsites, MARGIN= 2, FUN= mean,na.rm = TRUE) # average size-specific vital rates across sites
      predicted_gr <- apply(predicted_gr_allsites, MARGIN= 2, FUN= mean,na.rm = TRUE)
      predicted_pr <- apply(predicted_pr_allsites, MARGIN= 2, FUN= mean,na.rm = TRUE)
      predicted_cr <- apply(predicted_cr_allsites, MARGIN= 2, FUN= mean,na.rm = TRUE)
      predicted_vg <- apply(predicted_vg_allsites, MARGIN= 2, FUN= mean,na.rm = TRUE)
      predicted_vg[which(predicted_vg<0)] <- 0                 #### to prevent negative variance
      predicted_rec <- mean(predicted_rec_allsites,na.rm = TRUE)
      
      gmx <- matrix(NA, length(bin_mids),length(bin_mids)) ## create empty gmx before entering the no_bins loop
      
      # make the TSF-specific matrix
      for (ss in 1:no_bins) {
        growcdf <- pnorm(bin_edges,predicted_gr[ss],sqrt(predicted_vg[ss])) # Yes. Vargrowth=(pred - obs)^2
        grows <- growcdf[2:length(bin_edges)]-growcdf[1:(length(bin_edges)-1)]
        if(sum(grows)>0){grows <- grows/sum(grows)
        gmx[,ss] <- grows
        } else if (sum(grows)==0 & grow_vals[ss]< min(bin_edges)) { 
          gmx[,ss] <- c(1, rep(0, length(binmids)-1))  
        } else if (sum(grows)==0 & grow_vals[ss]> max(bin_edges)) { 
          gmx[,ss] <- c(rep(0, length(binmids)-1), 1)  } else { # deals with complete eviction-- OUT OF THE BOUNDS  of the kernel
            gmx[,ss] <- NA} 
        # this if statement breaks the code (puts NA's into the matrix) if the sum of the PDF is zero (which happens if all the probability is outside of the size bounds)
      } # end ss loop
      
      # make the surv*growth mx
      survgmx <- gmx*t(matrix( rep(predicted_sur,no_bins),no_bins)) # Yes, this is doing element-wise multiplication 
      reprow <-  predicted_sur*predicted_pr*predicted_cr # prob that you survive, reproduce, then number fruits|reproduce; this multiplcation does it element by element
      
      mx <- matrix(0, no_bins+1, no_bins+1)
      mx[2:(no_bins+1), 2:(no_bins+1)] = survgmx
      mx[1,2:(no_bins+1)] = reprow # FRUIT ROW
      
      sdlng_cdf <- pnorm(bin_edges,mean_seedling_logsize,sqrt(var_seedling_logsize)) 
      sdlng_pdf <- sdlng_cdf[2:length(bin_edges)]-sdlng_cdf[1:(length(bin_edges)-1)]
      if(sum(sdlng_pdf)>0){
        sdlng_pdf <- sdlng_pdf/sum(sdlng_pdf)} else {
          sdlng_pdf <- NA} 
      
      mx[2:(no_bins+1),1]  = predicted_rec *sdlng_pdf 
      
      mx_prod <- mx_prod %*% mx # in the first iteration of the loop, mx just becomes mx_prod
      
    } # closes ii (TSFs) loop
    
    j_pop_gr_rate_GSP[i] <- (popbio::lambda(mx_prod))^(1/length(TSFs)) 
    # multiply them together to get a matrix, etc. 
  }
    j_pop_gr_rate_GSP
}


all_pop_growth_rates_GSP <- cbind(possible_scenarios, all_pop_growth_rates_GSP)
names(all_pop_growth_rates_GSP) <- c(names(possible_scenarios), paste("rep", 1:no_reps, sep = ""))
rep_cols <- grep("^rep", names(all_pop_growth_rates_GSP), value = TRUE)

mean_lambda_across_scenarios_GSP <- all_pop_growth_rates_GSP %>%
  rowwise() %>%
  mutate(
    mean_lambda = mean(c_across(all_of(rep_cols)), na.rm = TRUE),
    lower_bound = quantile(c_across(all_of(rep_cols)), 0.025, na.rm = TRUE),
    upper_bound = quantile(c_across(all_of(rep_cols)), 0.975, na.rm = TRUE)
  ) %>%
  dplyr::select(FRI,s.TBF,mean_lambda,lower_bound, upper_bound) %>%
  mutate(TBF=NA) %>%
  ungroup()

mean_lambda_across_scenarios_GSP$TBF[1:11] <- "Short TBF"
mean_lambda_across_scenarios_GSP$TBF[12:22] <- "Long TBF"

##########. graph 
##########

plt3 <- ggplot(data= mean_lambda_across_scenarios_GSP, aes(x= FRI, y= mean_lambda)) +
  geom_line(aes(color= TBF)) + 
  geom_ribbon(aes(ymin = mean_lambda_across_scenarios_Cro$lower_bound, ymax = mean_lambda_across_scenarios_Cro$upper_bound, fill= TBF), alpha = 0.1) + 
  #ylim(0,1) +
  #xlim(c(0, 12)) +
  labs(y= "Growth Rate", x = "Fire Return Interval",color = NULL, fill = NULL) +
  theme_bw() + theme(legend.position='none') +
  theme(legend.position = "right") +
  scale_x_continuous(breaks = 0:10) +
  theme(text = element_text(size = 16)) 
plt3
```

## graph

```{r}
lambda_globalIPM <- bind_rows(
  mean_lambda_across_scenarios %>% mutate(dataset = "global"),
  mean_lambda_across_scenarios_Cro %>% mutate(dataset = "Croatan"),
  mean_lambda_across_scenarios_GSP %>% mutate(dataset = "Green Swamp")
)

plt4 <- ggplot(lambda_globalIPM, aes(x = FRI, y = mean_lambda)) +
  
  # ribbons
  geom_ribbon(
    aes(ymin = lower_bound, ymax = upper_bound,
        fill = dataset, group = interaction(dataset, TBF)),
    alpha = 0.1) +
  geom_line(
    aes(color = dataset, linetype = TBF,
        group = interaction(dataset, TBF)),
    size = 1) +
  theme_bw() +
  labs(
    y = "Growth Rate",
    x = "Fire Return Interval",
    color = "Dataset",
    fill = "Dataset",
    linetype = "TBF") +
  scale_x_continuous(breaks = 0:10) +
  theme(text = element_text(size = 16))

plt4


```

Here, the differences in growth rate among regions are driven solely by the site fixed effect in two of the vital rate functions (survival and prob of reproduction).



# build region specific IPM (refit vital rate functions using subsets of data)




##Croatan

```{r}
TBF_long_Cro <- TBF_long %>%
  filter(site=="CH"|site=="CM")
```

###refit vital rate functions 

####survival

```{r}
sur_subset_Cro <- 
  TBF_long_Cro %>% 
  filter_at(vars(sur0_1, s.logsize0, s.TSF,s.TBF, site), all_vars(!is.na(.)))

GM_R_sur <- glm(
   sur0_1 ~ s.logsize0 + s.TSF * s.TBF +
    s.T_LA + s.T_LH + s.sq.T_LH + s.T_LD +s.P_RH +
    s.P_RH:s.T_LH + s.P_RD + 
    site,
  data = sur_subset_Cro,
  family = "binomial", 
  na.action = "na.fail"
)

n_cores <- detectCores()
n_cores
cluster <- makeCluster(n_cores - 1)
registerDoParallel(cluster)
clusterExport(cluster, c("sur_subset_Cro"))   ### replace with different subset (different global models)
clusterEvalQ(cluster, {library(lme4); library(MuMIn)})

sur_dredge_Cro <- MuMIn::dredge(
  GM_R_sur,
  cluster = cluster,
  trace   = 2
)

sur_mod_Cro <- get.models(sur_dredge_Cro, 1)[[1]]
summary(sur_mod_Cro)

stopCluster(cluster)
```

####growth

```{r}
gr_subset_Cro <- 
  TBF_long_Cro[which(TBF_long_Cro$sur0_1 == 1),] %>%      ## make sure to subset data to only alive plants
  filter_at(vars(logsize1, s.logsize0, s.TSF,s.sq.TSF, s.TBF, site, s.sq.T_RA,s.T_RA, s.P_RA, s.T_RC, 
                 s.T_RD, s.P_RH), all_vars(!is.na(.)))
GM_R_gr <- lmer(
  logsize1 ~ s.logsize0 + s.TSF * s.TBF  + s.T_LA + s.sq.T_LA + s.P_RA + s.T_LA:s.P_RA +
    s.T_LC + s.T_LD + s.P_RH + (1|site),
  data = gr_subset_Cro,
  na.action = "na.fail"
)

n_cores <- detectCores()
n_cores
cluster <- makeCluster(n_cores - 1)
registerDoParallel(cluster)
clusterExport(cluster, c("gr_subset_Cro"))   ### replace with different subset (different global models)
clusterEvalQ(cluster, {library(lme4); library(MuMIn)})

gr_dredge_Cro <- MuMIn::dredge(
  GM_R_gr,
  cluster = cluster,
  trace   = 2
)
growth_mod_Cro<- get.models(gr_dredge_Cro, 1)[[1]]
summary(growth_mod_Cro)
stopCluster(cluster)


```

####prep

```{r}
prep_subset_Cro <- 
  TBF_long_Cro %>% 
  filter_at(vars(prep1, s.logsize0, s.TSF, s.TBF,s.sq.TSF, site, s.T_RA,s.P_RA, s.T_RH, s.T_RC, 
                 s.P_RH, ), all_vars(!is.na(.)))

GM_R_prep <- glm(
  prep1 ~ s.logsize0 + s.TSF * s.TBF +
    s.T_LA + s.sq.T_LA +s.P_RA+ s.sq.P_RA + s.T_LA:s.P_RA+
    s.T_LH + s.P_RH + s.T_LH:s.P_RH + s.T_LC +
    site,
  data = prep_subset_Cro,
  family = "binomial", 
  na.action = "na.fail"
)

n_cores <- detectCores()
n_cores
cluster <- makeCluster(n_cores - 1)
registerDoParallel(cluster)
clusterExport(cluster, c("prep_subset_Cro"))   ### replace with different subset (different global models)
clusterEvalQ(cluster, {library(lme4); library(MuMIn)})

prep_dredge_Cro <- MuMIn::dredge(
  GM_R_prep,
  cluster = cluster,
  trace   = 2
)
prep_mod_Cro <- get.models(prep_dredge_Cro, 1)[[1]]

stopCluster(cluster)
```

####crep

```{r}
crep_subset_Cro <- 
  TBF_long_Cro %>% 
  filter_at(vars(logcrep1, s.logsize0, TSF, TBF,s.sq.TSF, site, s.T_RA, s.T_RH), all_vars(!is.na(.)))

GM_R_crep <- lmer(
  logcrep1 ~ s.logsize0 + s.TSF * s.TBF  + s.T_LA + s.T_LH +
    (1 | site),
  data = crep_subset_Cro,
  na.action = "na.fail"
)

n_cores <- detectCores()
n_cores
cluster <- makeCluster(n_cores - 1)
registerDoParallel(cluster)
clusterExport(cluster, c("crep_subset_Cro"))   ### replace with different subset (different global models)
clusterEvalQ(cluster, {library(lme4); library(MuMIn)})

crep_dredge_Cro <- MuMIn::dredge(
  GM_R_crep,
  cluster = cluster,
  trace   = 2
)
crep_mod_Cro <- get.models(crep_dredge_Cro, 1)[[1]]

stopCluster(cluster)
```

#### recruitment

since recruitment sample size is very low and subsetting would result in inadequate sample sizes, we use the recruitment model from the global IPM

```{r}
recruit_mod_Cro <- recruit_mod
```

#### variance in growth

```{r}
growth_terms <- attr(terms(growth_mod_Cro), "term.labels")
int_terms <- grep(":", growth_terms, value = TRUE)
int_terms_broken <- unique(unlist(strsplit(int_terms, ":", fixed = TRUE)))
main_terms <- growth_terms[!grepl(":", growth_terms)]
logsize1 <- c("logsize1")
gr_all_terms <- unique(c(main_terms, int_terms_broken, logsize1))

Vargrowth_subset_Cro <- 
  TBF_long_Cro[which(TBF_long_Cro$sur0_1 == 1),] %>%      ## make sure to subset data to only alive plants
  filter_at(vars( s.logsize0, TSF, TBF,logsize1, site), all_vars(!is.na(.)))

Vargrowth_subset_Cro$vargrowth <- NA
Vargrowth_subset_Cro$predgrowth <- NA

Vargrowth_subset_Cro$predgrowth <- predict(growth_mod_Cro,newdata=Vargrowth_subset_Cro)

Vargrowth_subset_Cro$vargrowth <-
  (Vargrowth_subset_Cro$predgrowth - # variance is equivalent to (predicted-expected)^2
     Vargrowth_subset_Cro$logsize1)^2

Vargrowth_subset_Cro$log.vargrowth <- log(Vargrowth_subset_Cro$vargrowth)

#***** to prevent negative numbers later

GM_vargrowth_R <- lmer(log.vargrowth ~ s.logsize0 + s.T_LA+ s.TSF * s.TBF  +
                         s.P_RA  + s.sq.P_RA  + s.T_LC +s.T_LD +
                         s.P_RD + s.P_RW
                       + (1|site), #  
                       data = Vargrowth_subset_Cro, na.action = "na.fail") 


n_cores <- detectCores()
cluster <- makeCluster(n_cores - 1)
registerDoParallel(cluster)
clusterExport(cluster, c("Vargrowth_subset_Cro"))   ### replace with different subset (different global models)
clusterEvalQ(cluster, {library(lme4); library(MuMIn)})


vargrowth_dredge_Cro <- MuMIn::dredge(
  GM_vargrowth_R,
  cluster = cluster,
  trace   = 2
)

vargrowth_mod_Cro <- get.models(vargrowth_dredge_Cro, 1)[[1]]

stopCluster(cluster)
```

### Croatan IPM

keep the same IPM structure as the global IPM (for example, size bound is the same as the global IPM)

```{r}


TBF_long$logsize0 <- (log(TBF_long$size0+0.1))
TBF_long$s.logsize0 <- scale(TBF_long$logsize0) 
TBF_long$s.TBF <- scale(TBF_long$TBF) 
TBF_long$s.TSF <- scale(TBF_long$TSF) 
TBF_long$sq.TSF <- (TBF_long$TSF)^2
TBF_long$s.sq.TSF <- scale(TBF_long$sq.TSF) 
IPM_bounds <- range(TBF_long$logsize0, na.rm = TRUE)  # our IPM will be based on logsize as a size metric. 
no_bins <- 100
bin_edges <- seq(IPM_bounds[1], IPM_bounds[2], length.out= no_bins+1)
bin_mids <- zoo::rollmean(bin_edges,2)

# define mean and sd of seedling sizes
rows_with_new <- TBF_long[grepl("\\bnew", TBF_long$comm1), ]
rows_with_new <- rows_with_new[!grepl("no new", rows_with_new$comm1), ] # delete "no new"
rows_with_new <- rows_with_new[!grepl("no news", rows_with_new$comm1), ] # delete "no new" 
rows_with_new <- rows_with_new[!grepl("not possible to see new plants", rows_with_new$comm1), ]
rows_with_new <- rows_with_new[!grepl("No news", rows_with_new$comm1), ]
rows_with_new <- rows_with_new[!grepl("new lvs", rows_with_new$comm1), ]
rows_with_new <- rows_with_new[!grepl("new nail", rows_with_new$comm1), ]

mean_seedling_logsize <- mean(rows_with_new$logsize1, na.rm=T) # get the mean logsize of seedings   ### the mean seedling size is very large
var_seedling_logsize <- var(rows_with_new$logsize1,na.rm=T)# get the var in logsize of seedings  

### build empty matrices to store IPM output. Here, I'm interested in the effect of fire on lambda, so I'm setting average conditions for all other climate variables
possible_scenarios <- 
  as.data.frame(matrix(NA, nrow= 22, ncol = 28))
names(possible_scenarios) <- c("FRI","s.TBF","s.T_LA", "s.T_LH",
  "s.sq.T_LH",
  "s.sq.P_LA",
  "s.T_LC",
  "s.T_LD",
  "s.P_LD",
  "s.sq.T_LA",
  "s.P_LA",
  "s.P_LW",
  "s.P_LH",
  "s.T_RA",
  "s.sq.T_RA",
  "s.P_RA",
  "s.T_RH",
  "s.P_RH",
  "s.T_RC",
  "s.T_RW",
  "s.T_RD",
  "s.sq.T_RC",
  "s.P_RW",
  "s.P_RD",
  "s.sq.T_RH",
  "s.sq.P_RA",
  "s.sq.P_RW",
  "s.sq.P_RD")


possible_scenarios$FRI <- rep(1:11, 2) # we want FRI to range from 1-11
possible_scenarios$s.TBF <- scale(rep(c(0, 10), each= 11) ,# raw TBF values
                                  center= attr(TBF_long$s.TBF,"scaled:center"), 
                                  scale= attr(TBF_long$s.TBF,"scaled:scale")) # for each set of FRI's, we want to set a short TBF (of 3) and a long TBF (of 11)

possible_scenarios[, c(  "s.T_LA",  "s.T_LH",  "s.sq.T_LH",  "s.sq.P_LA",  "s.T_LC",  "s.T_LD",  "s.P_LD",  "s.sq.T_LA",  "s.P_LA",  
                         "s.P_LW",  "s.P_LH",  "s.T_RA",  "s.sq.T_RA",  "s.P_RA",  "s.T_RH",  "s.P_RH",  "s.T_RC",  "s.T_RW",  "s.T_RD",
                         "s.sq.T_RC","s.P_RW","s.P_RD", "s.sq.T_RH",  "s.sq.P_RA",  "s.sq.P_RW",  "s.sq.P_RD")] <- 0 # setting all climate variable to their mean

my_sites <- unique(TBF_long_Cro$site) 
no_reps <- 1000 # or somethign similar

########### extract model components for later predict() ###############
sur_model_terms <- terms(sur_mod_Cro)
sur_term <- attr(sur_model_terms, "term.labels")
sur_term <- reformulate(sur_term)

gr_model_terms <- terms(growth_mod_Cro)
gr_term <- attr(gr_model_terms, "term.labels")
gr_term <- reformulate(gr_term)

prep_model_terms <- terms(prep_mod_Cro)
prep_term <- attr(prep_model_terms, "term.labels")
prep_term <- reformulate(prep_term)

crep_model_terms <- terms(crep_mod_Cro)
crep_term <- attr(crep_model_terms, "term.labels")
crep_term <- reformulate(crep_term)

vargrowth_model_terms <- terms(vargrowth_mod_Cro)
vargrowth_term <- attr(vargrowth_model_terms, "term.labels")
vargrowth_term <- reformulate(vargrowth_term)

rec_model_terms <- terms(recruit_mod_Cro)
rec_term <- attr(rec_model_terms, "term.labels")
rec_term <- reformulate(rec_term)
```

```{r}
all_pop_growth_rates_Cro_reg <- matrix(data= NA, nrow= dim(possible_scenarios)[1], ncol= no_reps)
n_cores <- detectCores()
cl <- makeCluster(n_cores - 1)
registerDoParallel(cl)
j_pop_gr_rate_Cro_reg <- rep(NA, dim(possible_scenarios)[1]) ### changed to no_scenarios

all_pop_growth_rates_Cro_reg <- foreach(
  j = 1:no_reps,
  .combine = cbind,
  .packages = c("MASS", "lme4", "popbio", "dplyr")
) %dopar% {
  rsur_coefs <- coefficients(sur_mod_Cro) # first, assign all "randomy selected" coefficients to the mean estimated coefficeints
  rsur_coefs[!names(coefficients(sur_mod_Cro)) %in% paste("site", my_sites, sep="")] <- MASS::mvrnorm(mu= coefficients(sur_mod_Cro), Sigma = vcov(sur_mod_Cro))[!names(coefficients(sur_mod_Cro)) %in% paste("site", my_sites, sep="")]
  #rsur_coefs <- lme4::fixef(survival_mod)
  #rsur_coefs <-  MASS::mvrnorm(mu= lme4::fixef(survival_mod) , Sigma = vcov(survival_mod))
  rgr_coefs <- lme4::fixef(growth_mod_Cro)
  rgr_coefs <-  MASS::mvrnorm(mu= lme4::fixef(growth_mod_Cro) , Sigma = vcov(growth_mod_Cro)) # no site effect in this model
  rprep_coefs <- coefficients(prep_mod_Cro)
  rprep_coefs[!names(coefficients(prep_mod_Cro)) %in% paste("site", my_sites, sep="")] <- MASS::mvrnorm(mu= coefficients(prep_mod_Cro), Sigma = vcov(prep_mod_Cro))[!names(coefficients(prep_mod_Cro)) %in% paste("site", my_sites, sep="")]
  rcrep_coefs <- lme4::fixef(crep_mod_Cro)
  rcrep_coefs <-  MASS::mvrnorm(mu= lme4::fixef(crep_mod_Cro) , Sigma = vcov(crep_mod_Cro)) # no site effect in this model
  rvar_coefs <- lme4::fixef(vargrowth_mod_Cro)
  rvar_coefs <-  MASS::mvrnorm(mu= lme4::fixef(vargrowth_mod_Cro) , Sigma = vcov(vargrowth_mod_Cro)) # no site effect in this model
  rrec_coefs <- lme4::fixef(recruit_mod_Cro)
  rrec_coefs <-  MASS::mvrnorm(mu= lme4::fixef(recruit_mod_Cro) , Sigma = vcov(recruit_mod_Cro)) # no site effect in this model
  #rrec_coefs <-  MASS::mvrnorm(mu= lme4::fixef(recruit_mod) , Sigma = vcov(recruit_mod)) 
  
  for (i in 1: dim(possible_scenarios)[1]) {
    
    i_scenario <- possible_scenarios[i,] # which combination of FRI, TBF, and climate do I want to use
    
    TSFs <- 0:(i_scenario$FRI-1) # TSF range from 0 to (FRI-1) such that an FRI of 1 will always have TSF= 0
    i_scenario <- i_scenario[rep(seq_len(nrow(i_scenario)), length(TSFs)), ]
    i_scenario$TSF <- TSFs
    i_scenario$s.TSF <- scale(TSFs, center= attr(TBF_long$s.TSF,"scaled:center"), 
                              scale= attr(TBF_long$s.TSF,"scaled:scale"))
    i_scenario$s.sq.TSF <- scale(TSFs^2, center= attr(TBF_long$s.sq.TSF,"scaled:center"), 
                                 scale= attr(TBF_long$s.sq.TSF,"scaled:scale"))
    
    mx_prod <- diag(no_bins+1) # setting up an identity matrix before you enter the loop
    
    for (ii in TSFs){
      i_ii_scenario <- i_scenario[which(i_scenario$TSF==ii),]
      data_for_prediction <- i_ii_scenario[rep(seq_len(nrow(i_ii_scenario)), length(bin_mids)), ]
      data_for_prediction$s.logsize0 <-   scale(bin_mids, center= attr(TBF_long$s.logsize0,"scaled:center"), 
                                                scale= attr(TBF_long$s.logsize0,"scaled:scale"))
      data_for_prediction_rec <- cbind(i_ii_scenario)
      
      predicted_sur <- predicted_gr <-predicted_pr <- predicted_cr <- predicted_vg <- matrix(data=NA, nrow= length(my_sites), ncol= length(bin_mids))
      predicted_rec <- rep(NA, length(my_sites))
      
      ### make empty mxes before entering the site loop
      
      predicted_sur_allsites <-predicted_gr_allsites<-predicted_pr_allsites <-predicted_cr_allsites<-predicted_vg_allsites <-matrix(NA, nrow = length(my_sites), ncol = no_bins, dimnames = list(my_sites, NULL))
      predicted_rec_allsites <- matrix(NA, nrow = length(my_sites), ncol = 1, dimnames = list(my_sites, NULL))
      
      
      for (s in my_sites){ #looping through the sites
        s_data_for_prediction <- data_for_prediction
        s_data_for_prediction_rec <- data_for_prediction_rec
        s_data_for_prediction$site <- s
        s_data_for_prediction_rec$site <- s
        
        s_data_for_prediction$site     <- factor(s_data_for_prediction$site,  levels = my_sites)   ## it is important to add the levels=my_sites here to ensure that the model matrix is built correctly
        s_data_for_prediction_rec$site <- factor(s_data_for_prediction_rec$site, levels = my_sites)
        
        s_data_for_prediction <- s_data_for_prediction %>% mutate(across(c("s.logsize0", "s.sq.TSF", "s.TSF", "s.TBF"), as.numeric))   
        
        m_sur <- model.matrix(sur_term ,data=s_data_for_prediction)
        p2_sur <- rsur_coefs %*% t(m_sur) # check on whether this coef(survival_mod is a column or a row!!!!!!!!
        p2_sur <- plogis(p2_sur)   ## transform to binomial 
        predicted_sur_allsites[s,] <- p2_sur
        
        m_gr <- model.matrix(gr_term, data=s_data_for_prediction)
        p2_gr <- rgr_coefs %*% t(m_gr) # check on whether this coef(survival_mod is a column or a row!!!!!!!!
        predicted_gr_allsites[s,] <- p2_gr
        
        m_prep <- model.matrix(prep_term,data=s_data_for_prediction)
        p2_prep <- rprep_coefs %*% t(m_prep) # check on whether this coef(survival_mod is a column or a row!!!!!!!!
        p2_prep <- plogis(p2_prep)   ## transform to binomial 
        predicted_pr_allsites[s,] <- p2_prep
        
        m_crep <- model.matrix(crep_term  ,data=s_data_for_prediction)
        p2_crep <- rcrep_coefs %*% t(m_crep) # check on whether this coef(survival_mod is a column or a row!!!!!!!!
        predicted_cr_allsites[s,] <- p2_crep
        
        m_vg <- model.matrix(vargrowth_term,data=s_data_for_prediction)
        p2_vg <- rvar_coefs %*% t(m_vg) # 
        predicted_vg_allsites[s,] <- exp(p2_vg)  
        
        m_rec <- model.matrix(rec_term , data=s_data_for_prediction_rec)
        p2_rec <- rrec_coefs %*% t(m_rec) # check on whether this coef(survival_mod is a column or a row!!!!!!!!
        predicted_rec_allsites[s,1] <- p2_rec*log(1.1)            #### account for composite rec variable (rec/log.fr)
        
        
        # for mixed models, the above functions will predict site-specific results (i.e., will incorporate random effects)
        rm(s_data_for_prediction) # this part removes the site-specific dataframe used in this loop, so you do not accidentaly use the previous sites' dataframe instead of the current one-- in case there was an error somehow
      }
      predicted_sur <- apply(predicted_sur_allsites, MARGIN= 2, FUN= mean, na.rm=T) # average size-specific vital rates across sites
      predicted_gr <- apply(predicted_gr_allsites, MARGIN= 2, FUN= mean, na.rm=T)
      predicted_pr <- apply(predicted_pr_allsites, MARGIN= 2, FUN= mean, na.rm=T)
      predicted_cr <- apply(predicted_cr_allsites, MARGIN= 2, FUN= mean, na.rm=T)
      predicted_vg <- apply(predicted_vg_allsites, MARGIN= 2, FUN= mean, na.rm=T)
      predicted_vg[which(predicted_vg<0)] <- 0                 #### to prevent negative variance
      predicted_rec <- mean(predicted_rec_allsites, na.rm=T)
      
      gmx <- matrix(NA, length(bin_mids),length(bin_mids)) ## create empty gmx before entering the no_bins loop
      
      # make the TSF-specific matrix
      for (ss in 1:no_bins) {
        growcdf <- pnorm(bin_edges,predicted_gr[ss],sqrt(predicted_vg[ss])) # Yes. Vargrowth=(pred - obs)^2
        grows <- growcdf[2:length(bin_edges)]-growcdf[1:(length(bin_edges)-1)]
        if(sum(grows)>0){grows <- grows/sum(grows)
        gmx[,ss] <- grows
        } else if (sum(grows)==0 & grow_vals[ss]< min(bin_edges)) { 
          gmx[,ss] <- c(1, rep(0, length(binmids)-1))  
        } else if (sum(grows)==0 & grow_vals[ss]> max(bin_edges)) { 
          gmx[,ss] <- c(rep(0, length(binmids)-1), 1)  } else { # deals with complete eviction-- OUT OF THE BOUNDS  of the kernel
            gmx[,ss] <- NA} 
        # this if statement breaks the code (puts NA's into the matrix) if the sum of the PDF is zero (which happens if all the probability is outside of the size bounds)
      } # end ss loop
      
      # make the surv*growth mx
      survgmx <- gmx*t(matrix( rep(predicted_sur,no_bins),no_bins)) # Yes, this is doing element-wise multiplication 
      reprow <-  predicted_sur*predicted_pr*predicted_cr # prob that you survive, reproduce, then number fruits|reproduce; this multiplcation does it element by element
      
      mx <- matrix(0, no_bins+1, no_bins+1)
      mx[2:(no_bins+1), 2:(no_bins+1)] = survgmx
      mx[1,2:(no_bins+1)] = reprow # FRUIT ROW
      
      sdlng_cdf <- pnorm(bin_edges,mean_seedling_logsize,sqrt(var_seedling_logsize)) 
      sdlng_pdf <- sdlng_cdf[2:length(bin_edges)]-sdlng_cdf[1:(length(bin_edges)-1)]
      if(sum(sdlng_pdf)>0){
        sdlng_pdf <- sdlng_pdf/sum(sdlng_pdf)} else {
          sdlng_pdf <- NA} 
      
      mx[2:(no_bins+1),1]  = predicted_rec *sdlng_pdf 
      
      mx_prod <- mx_prod %*% mx # in the first iteration of the loop, mx just becomes mx_prod
      
    } # closes ii (TSFs) loop
    
    j_pop_gr_rate_Cro_reg[i] <- (popbio::lambda(mx_prod))^(1/length(TSFs)) 
    # multiply them together to get a matrix, etc. 
  }
    j_pop_gr_rate_Cro_reg
}


all_pop_growth_rates_Cro_reg <- cbind(possible_scenarios, all_pop_growth_rates_Cro_reg)
names(all_pop_growth_rates_Cro_reg) <- c(names(possible_scenarios), paste("rep", 1:no_reps, sep = ""))
rep_cols <- grep("^rep", names(all_pop_growth_rates_Cro_reg), value = TRUE)

mean_lambda_across_scenarios_Cro_reg <- all_pop_growth_rates_Cro_reg %>%
  rowwise() %>%
  mutate(
    mean_lambda = mean(c_across(all_of(rep_cols)), na.rm = TRUE),
    lower_bound = quantile(c_across(all_of(rep_cols)), 0.025, na.rm = TRUE),
    upper_bound = quantile(c_across(all_of(rep_cols)), 0.975, na.rm = TRUE)
  ) %>%
  dplyr::select(FRI,s.TBF,mean_lambda,lower_bound, upper_bound) %>%
  mutate(TBF=NA) %>%
  ungroup()

mean_lambda_across_scenarios_Cro_reg$TBF[1:11] <- "Short TBF"
mean_lambda_across_scenarios_Cro_reg$TBF[12:22] <- "Long TBF"

##########. graph 
##########

plt5<- ggplot(data= mean_lambda_across_scenarios_Cro_reg, aes(x= FRI, y= mean_lambda)) +
  geom_line(aes(color= TBF)) + 
  geom_ribbon(aes(ymin = mean_lambda_across_scenarios_Cro_reg$lower_bound, ymax = mean_lambda_across_scenarios_Cro_reg$upper_bound, fill= TBF), alpha = 0.1) + 
  #ylim(0,1) +
  #xlim(c(0, 12)) +
  labs(y= "Growth Rate", x = "Fire Return Interval",color = NULL, fill = NULL) +
  theme_bw() + theme(legend.position='none') +
  theme(legend.position = "right") +
  scale_x_continuous(breaks = 0:10) +
  theme(text = element_text(size = 16)) 
plt5
```

## Green Swamp

```{r}
TBF_long_GSP <- TBF_long %>%
  filter(site=="GSP-LI"|site=="GSP-BI")
```


###refit vital rate 

####survival

```{r}
sur_subset_GSP <- 
  TBF_long_GSP %>% 
  filter_at(vars(sur0_1, s.logsize0, s.TSF,s.TBF, site), all_vars(!is.na(.)))

GM_R_sur <- glm(
   sur0_1 ~ s.logsize0 + s.TSF * s.TBF +
    s.T_LA + s.T_LH + s.sq.T_LH + s.T_LD +s.P_RH +
    s.P_RH:s.T_LH + s.P_RD + 
    site,
  data = sur_subset_GSP,
  family = "binomial", 
  na.action = "na.fail"
)

n_cores <- detectCores()
n_cores
cluster <- makeCluster(n_cores - 1)
registerDoParallel(cluster)
clusterExport(cluster, c("sur_subset_GSP"))   ### replace with different subset (different global models)
clusterEvalQ(cluster, {library(lme4); library(MuMIn)})

sur_dredge_GSP <- MuMIn::dredge(
  GM_R_sur,
  cluster = cluster,
  trace   = 2
)

sur_mod_GSP <- get.models(sur_dredge_GSP, 1)[[1]]
summary(sur_mod_GSP)

stopCluster(cluster)
```

####growth

```{r}
gr_subset_GSP <- 
  TBF_long_GSP[which(TBF_long_GSP$sur0_1 == 1),] %>%      ## make sure to subset data to only alive plants
  filter_at(vars(logsize1, s.logsize0, s.TSF,s.sq.TSF, s.TBF, site, s.sq.T_RA,s.T_RA, s.P_RA, s.T_RC, 
                 s.T_RD, s.P_RH), all_vars(!is.na(.)))
GM_R_gr <- lmer(
  logsize1 ~ s.logsize0 + s.TSF * s.TBF  + s.T_LA + s.sq.T_LA + s.P_RA + s.T_LA:s.P_RA +
    s.T_LC + s.T_LD + s.P_RH + (1|site),
  data = gr_subset_GSP,
  na.action = "na.fail"
)

n_cores <- detectCores()
n_cores
cluster <- makeCluster(n_cores - 1)
registerDoParallel(cluster)
clusterExport(cluster, c("gr_subset_GSP"))   ### replace with different subset (different global models)
clusterEvalQ(cluster, {library(lme4); library(MuMIn)})

gr_dredge_GSP <- MuMIn::dredge(
  GM_R_gr,
  cluster = cluster,
  trace   = 2
)
growth_mod_GSP<- get.models(gr_dredge_GSP, 1)[[1]]
summary(growth_mod_GSP)
stopCluster(cluster)
```
####prep

```{r}
prep_subset_GSP <- 
  TBF_long_GSP %>% 
  filter_at(vars(prep1, s.logsize0, s.TSF, s.TBF,s.sq.TSF, site, s.T_RA,s.P_RA, s.T_RH, s.T_RC, 
                 s.P_RH, ), all_vars(!is.na(.)))

GM_R_prep <- glm(
  prep1 ~ s.logsize0 + s.TSF * s.TBF +
    s.T_LA + s.sq.T_LA +s.P_RA+ s.sq.P_RA + s.T_LA:s.P_RA+
    s.T_LH + s.P_RH + s.T_LH:s.P_RH + s.T_LC +
    site,
  data = prep_subset_GSP,
  family = "binomial", 
  na.action = "na.fail"
)

n_cores <- detectCores()
n_cores
cluster <- makeCluster(n_cores - 1)
registerDoParallel(cluster)
clusterExport(cluster, c("prep_subset_GSP"))   ### replace with different subset (different global models)
clusterEvalQ(cluster, {library(lme4); library(MuMIn)})

prep_dredge_GSP <- MuMIn::dredge(
  GM_R_prep,
  cluster = cluster,
  trace   = 2
)
prep_mod_GSP <- get.models(prep_dredge_GSP, 1)[[1]]
summary(prep_mod_GSP)
stopCluster(cluster)
```

####crep

```{r}

crep_subset_GSP <- 
  TBF_long_GSP %>% 
  filter_at(vars(logcrep1, s.logsize0, TSF, TBF,s.sq.TSF, site, s.T_RA, s.T_RH), all_vars(!is.na(.)))

GM_R_crep <- lmer(
  logcrep1 ~ s.logsize0 + s.TSF * s.TBF  + s.T_LA + s.T_LH +
    (1 | site),
  data = crep_subset_GSP,
  na.action = "na.fail"
)

n_cores <- detectCores()
n_cores
cluster <- makeCluster(n_cores - 1)
registerDoParallel(cluster)
clusterExport(cluster, c("crep_subset_GSP"))   ### replace with different subset (different global models)
clusterEvalQ(cluster, {library(lme4); library(MuMIn)})

crep_dredge_GSP <- MuMIn::dredge(
  GM_R_crep,
  cluster = cluster,
  trace   = 2
)
crep_mod_GSP <- get.models(crep_dredge_GSP, 1)[[1]]

stopCluster(cluster)
```
####recruitment
```{r}
recruit_mod_GSP <- recruit_mod
```
####variance in growth

```{r}
growth_terms <- attr(terms(growth_mod_GSP), "term.labels")
int_terms <- grep(":", growth_terms, value = TRUE)
int_terms_broken <- unique(unlist(strsplit(int_terms, ":", fixed = TRUE)))
main_terms <- growth_terms[!grepl(":", growth_terms)]
logsize1 <- c("logsize1")
gr_all_terms <- unique(c(main_terms, int_terms_broken, logsize1))

Vargrowth_subset_GSP <- 
  TBF_long_GSP[which(TBF_long_GSP$sur0_1 == 1),] %>%      ## make sure to subset data to only alive plants
  filter_at(vars( s.logsize0, TSF, TBF,logsize1, site), all_vars(!is.na(.)))

Vargrowth_subset_GSP$vargrowth <- NA
Vargrowth_subset_GSP$predgrowth <- NA

Vargrowth_subset_GSP$predgrowth <- predict(growth_mod_GSP,newdata=Vargrowth_subset_GSP)

Vargrowth_subset_GSP$vargrowth <-
  (Vargrowth_subset_GSP$predgrowth - # variance is equivalent to (predicted-expected)^2
     Vargrowth_subset_GSP$logsize1)^2

Vargrowth_subset_GSP$log.vargrowth <- log(Vargrowth_subset_GSP$vargrowth)

#***** to prevent negative numbers later

GM_vargrowth_R <- lmer(log.vargrowth ~ s.logsize0 + s.T_LA+ s.TSF * s.TBF  +
                         s.P_RA  + s.sq.P_RA  + s.T_LC +s.T_LD +
                         s.P_RD + s.P_RW
                       + (1|site), #  
                       data = Vargrowth_subset_GSP, na.action = "na.fail") 


n_cores <- detectCores()
cluster <- makeCluster(n_cores - 1)
registerDoParallel(cluster)
clusterExport(cluster, c("Vargrowth_subset_GSP"))   ### replace with different subset (different global models)
clusterEvalQ(cluster, {library(lme4); library(MuMIn)})


vargrowth_dredge_GSP <- MuMIn::dredge(
  GM_vargrowth_R,
  cluster = cluster,
  trace   = 2
)

vargrowth_mod_GSP <- get.models(vargrowth_dredge_GSP, 1)[[1]]

stopCluster(cluster)
```
###Green Swamp IPM

```{r}
TBF_long$logsize0 <- (log(TBF_long$size0+0.1))
TBF_long$s.logsize0 <- scale(TBF_long$logsize0) 
TBF_long$s.TBF <- scale(TBF_long$TBF) 
TBF_long$s.TSF <- scale(TBF_long$TSF) 
TBF_long$sq.TSF <- (TBF_long$TSF)^2
TBF_long$s.sq.TSF <- scale(TBF_long$sq.TSF) 
IPM_bounds <- range(TBF_long$logsize0, na.rm = TRUE)  # our IPM will be based on logsize as a size metric. 
no_bins <- 100
bin_edges <- seq(IPM_bounds[1], IPM_bounds[2], length.out= no_bins+1)
bin_mids <- zoo::rollmean(bin_edges,2)

# define mean and sd of seedling sizes
rows_with_new <- TBF_long[grepl("\\bnew", TBF_long$comm1), ]
rows_with_new <- rows_with_new[!grepl("no new", rows_with_new$comm1), ] # delete "no new"
rows_with_new <- rows_with_new[!grepl("no news", rows_with_new$comm1), ] # delete "no new" 
rows_with_new <- rows_with_new[!grepl("not possible to see new plants", rows_with_new$comm1), ]
rows_with_new <- rows_with_new[!grepl("No news", rows_with_new$comm1), ]
rows_with_new <- rows_with_new[!grepl("new lvs", rows_with_new$comm1), ]
rows_with_new <- rows_with_new[!grepl("new nail", rows_with_new$comm1), ]

mean_seedling_logsize <- mean(rows_with_new$logsize1, na.rm=T) # get the mean logsize of seedings   ### the mean seedling size is very large
var_seedling_logsize <- var(rows_with_new$logsize1,na.rm=T)# get the var in logsize of seedings  

### build empty matrices to store IPM output. Here, I'm interested in the effect of fire on lambda, so I'm setting average conditions for all other climate variables
possible_scenarios <- 
  as.data.frame(matrix(NA, nrow= 22, ncol = 28))
names(possible_scenarios) <- c("FRI","s.TBF","s.T_LA", "s.T_LH",
  "s.sq.T_LH",
  "s.sq.P_LA",
  "s.T_LC",
  "s.T_LD",
  "s.P_LD",
  "s.sq.T_LA",
  "s.P_LA",
  "s.P_LW",
  "s.P_LH",
  "s.T_RA",
  "s.sq.T_RA",
  "s.P_RA",
  "s.T_RH",
  "s.P_RH",
  "s.T_RC",
  "s.T_RW",
  "s.T_RD",
  "s.sq.T_RC",
  "s.P_RW",
  "s.P_RD",
  "s.sq.T_RH",
  "s.sq.P_RA",
  "s.sq.P_RW",
  "s.sq.P_RD")


possible_scenarios$FRI <- rep(1:11, 2) # we want FRI to range from 1-11
possible_scenarios$s.TBF <- scale(rep(c(0, 10), each= 11) ,# raw TBF values
                                  center= attr(TBF_long$s.TBF,"scaled:center"), 
                                  scale= attr(TBF_long$s.TBF,"scaled:scale")) # for each set of FRI's, we want to set a short TBF (of 3) and a long TBF (of 11)

possible_scenarios[, c(  "s.T_LA",  "s.T_LH",  "s.sq.T_LH",  "s.sq.P_LA",  "s.T_LC",  "s.T_LD",  "s.P_LD",  "s.sq.T_LA",  "s.P_LA",  
                         "s.P_LW",  "s.P_LH",  "s.T_RA",  "s.sq.T_RA",  "s.P_RA",  "s.T_RH",  "s.P_RH",  "s.T_RC",  "s.T_RW",  "s.T_RD",
                         "s.sq.T_RC","s.P_RW","s.P_RD", "s.sq.T_RH",  "s.sq.P_RA",  "s.sq.P_RW",  "s.sq.P_RD")] <- 0 # setting all climate variable to their mean

my_sites <- unique(TBF_long_GSP$site) 
no_reps <- 1000 # or somethign similar

########### extract model components for later predict() ###############
sur_model_terms <- terms(sur_mod_GSP)
sur_term <- attr(sur_model_terms, "term.labels")
sur_term <- reformulate(sur_term)

gr_model_terms <- terms(growth_mod_GSP)
gr_term <- attr(gr_model_terms, "term.labels")
gr_term <- reformulate(gr_term)

prep_model_terms <- terms(prep_mod_GSP)
prep_term <- attr(prep_model_terms, "term.labels")
prep_term <- reformulate(prep_term)

crep_model_terms <- terms(crep_mod_GSP)
crep_term <- attr(crep_model_terms, "term.labels")
crep_term <- reformulate(crep_term)

vargrowth_model_terms <- terms(vargrowth_mod_GSP)
vargrowth_term <- attr(vargrowth_model_terms, "term.labels")
vargrowth_term <- reformulate(vargrowth_term)

rec_model_terms <- terms(recruit_mod_GSP)
rec_term <- attr(rec_model_terms, "term.labels")
rec_term <- reformulate(rec_term)
```

```{r}
all_pop_growth_rates_GSP_reg <- matrix(data= NA, nrow= dim(possible_scenarios)[1], ncol= no_reps)
n_cores <- detectCores()
cl <- makeCluster(n_cores - 1)
registerDoParallel(cl)
j_pop_gr_rate_GSP_reg <- rep(NA, dim(possible_scenarios)[1]) ### changed to no_scenarios

all_pop_growth_rates_GSP_reg <- foreach(
  j = 1:no_reps,
  .combine = cbind,
  .packages = c("MASS", "lme4", "popbio", "dplyr")
) %dopar% {
  rsur_coefs <- coefficients(sur_mod_GSP) # first, assign all "randomy selected" coefficients to the mean estimated coefficeints
  rsur_coefs[!names(coefficients(sur_mod_GSP)) %in% paste("site", my_sites, sep="")] <- MASS::mvrnorm(mu= coefficients(sur_mod_GSP), Sigma = vcov(sur_mod_GSP))[!names(coefficients(sur_mod_GSP)) %in% paste("site", my_sites, sep="")]
  #rsur_coefs <- lme4::fixef(survival_mod)
  #rsur_coefs <-  MASS::mvrnorm(mu= lme4::fixef(survival_mod) , Sigma = vcov(survival_mod))
  rgr_coefs <- lme4::fixef(growth_mod_GSP)
  rgr_coefs <-  MASS::mvrnorm(mu= lme4::fixef(growth_mod_GSP) , Sigma = vcov(growth_mod_GSP)) # no site effect in this model
  rprep_coefs <- coefficients(prep_mod_GSP)
  rprep_coefs[!names(coefficients(prep_mod_GSP)) %in% paste("site", my_sites, sep="")] <- MASS::mvrnorm(mu= coefficients(prep_mod_GSP), Sigma = vcov(prep_mod_GSP))[!names(coefficients(prep_mod_GSP)) %in% paste("site", my_sites, sep="")]
  rcrep_coefs <- lme4::fixef(crep_mod_GSP)
  rcrep_coefs <-  MASS::mvrnorm(mu= lme4::fixef(crep_mod_GSP) , Sigma = vcov(crep_mod_GSP)) # no site effect in this model
  rvar_coefs <- lme4::fixef(vargrowth_mod_GSP)
  rvar_coefs <-  MASS::mvrnorm(mu= lme4::fixef(vargrowth_mod_GSP) , Sigma = vcov(vargrowth_mod_GSP)) # no site effect in this model
  rrec_coefs <- lme4::fixef(recruit_mod_GSP)
  rrec_coefs <-  MASS::mvrnorm(mu= lme4::fixef(recruit_mod_GSP) , Sigma = vcov(recruit_mod_GSP)) # no site effect in this model
  #rrec_coefs <-  MASS::mvrnorm(mu= lme4::fixef(recruit_mod) , Sigma = vcov(recruit_mod)) 
  
  for (i in 1: dim(possible_scenarios)[1]) {
    
    i_scenario <- possible_scenarios[i,] # which combination of FRI, TBF, and climate do I want to use
    
    TSFs <- 0:(i_scenario$FRI-1) # TSF range from 0 to (FRI-1) such that an FRI of 1 will always have TSF= 0
    i_scenario <- i_scenario[rep(seq_len(nrow(i_scenario)), length(TSFs)), ]
    i_scenario$TSF <- TSFs
    i_scenario$s.TSF <- scale(TSFs, center= attr(TBF_long$s.TSF,"scaled:center"), 
                              scale= attr(TBF_long$s.TSF,"scaled:scale"))
    i_scenario$s.sq.TSF <- scale(TSFs^2, center= attr(TBF_long$s.sq.TSF,"scaled:center"), 
                                 scale= attr(TBF_long$s.sq.TSF,"scaled:scale"))
    
    mx_prod <- diag(no_bins+1) # setting up an identity matrix before you enter the loop
    
    for (ii in TSFs){
      i_ii_scenario <- i_scenario[which(i_scenario$TSF==ii),]
      data_for_prediction <- i_ii_scenario[rep(seq_len(nrow(i_ii_scenario)), length(bin_mids)), ]
      data_for_prediction$s.logsize0 <-   scale(bin_mids, center= attr(TBF_long$s.logsize0,"scaled:center"), 
                                                scale= attr(TBF_long$s.logsize0,"scaled:scale"))
      data_for_prediction_rec <- cbind(i_ii_scenario)
      
      predicted_sur <- predicted_gr <-predicted_pr <- predicted_cr <- predicted_vg <- matrix(data=NA, nrow= length(my_sites), ncol= length(bin_mids))
      predicted_rec <- rep(NA, length(my_sites))
      
      ### make empty mxes before entering the site loop
      
      predicted_sur_allsites <-predicted_gr_allsites<-predicted_pr_allsites <-predicted_cr_allsites<-predicted_vg_allsites <-matrix(NA, nrow = length(my_sites), ncol = no_bins, dimnames = list(my_sites, NULL))
      predicted_rec_allsites <- matrix(NA, nrow = length(my_sites), ncol = 1, dimnames = list(my_sites, NULL))
      
      
      for (s in my_sites){ #looping through the sites
        s_data_for_prediction <- data_for_prediction
        s_data_for_prediction_rec <- data_for_prediction_rec
        s_data_for_prediction$site <- s
        s_data_for_prediction_rec$site <- s
        
        s_data_for_prediction$site     <- factor(s_data_for_prediction$site,  levels = my_sites)   ## it is important to add the levels=my_sites here to ensure that the model matrix is built correctly
        s_data_for_prediction_rec$site <- factor(s_data_for_prediction_rec$site, levels = my_sites)
        
        s_data_for_prediction <- s_data_for_prediction %>% mutate(across(c("s.logsize0", "s.sq.TSF", "s.TSF", "s.TBF"), as.numeric))   
        
        m_sur <- model.matrix(sur_term ,data=s_data_for_prediction)
        p2_sur <- rsur_coefs %*% t(m_sur) # check on whether this coef(survival_mod is a column or a row!!!!!!!!
        p2_sur <- plogis(p2_sur)   ## transform to binomial 
        predicted_sur_allsites[s,] <- p2_sur
        
        m_gr <- model.matrix(gr_term, data=s_data_for_prediction)
        p2_gr <- rgr_coefs %*% t(m_gr) # check on whether this coef(survival_mod is a column or a row!!!!!!!!
        predicted_gr_allsites[s,] <- p2_gr
        
        m_prep <- model.matrix(prep_term,data=s_data_for_prediction)
        p2_prep <- rprep_coefs %*% t(m_prep) # check on whether this coef(survival_mod is a column or a row!!!!!!!!
        p2_prep <- plogis(p2_prep)   ## transform to binomial 
        predicted_pr_allsites[s,] <- p2_prep
        
        m_crep <- model.matrix(crep_term  ,data=s_data_for_prediction)
        p2_crep <- rcrep_coefs %*% t(m_crep) # check on whether this coef(survival_mod is a column or a row!!!!!!!!
        predicted_cr_allsites[s,] <- p2_crep
        
        m_vg <- model.matrix(vargrowth_term,data=s_data_for_prediction)
        p2_vg <- rvar_coefs %*% t(m_vg) # 
        predicted_vg_allsites[s,] <- exp(p2_vg)  
        
        m_rec <- model.matrix(rec_term , data=s_data_for_prediction_rec)
        p2_rec <- rrec_coefs %*% t(m_rec) # check on whether this coef(survival_mod is a column or a row!!!!!!!!
        predicted_rec_allsites[s,1] <- p2_rec*log(1.1)            #### account for composite rec variable (rec/log.fr)
        
        
        # for mixed models, the above functions will predict site-specific results (i.e., will incorporate random effects)
        rm(s_data_for_prediction) # this part removes the site-specific dataframe used in this loop, so you do not accidentaly use the previous sites' dataframe instead of the current one-- in case there was an error somehow
      }
      predicted_sur <- apply(predicted_sur_allsites, MARGIN= 2, FUN= mean, na.rm=T) # average size-specific vital rates across sites
      predicted_gr <- apply(predicted_gr_allsites, MARGIN= 2, FUN= mean, na.rm=T)
      predicted_pr <- apply(predicted_pr_allsites, MARGIN= 2, FUN= mean, na.rm=T)
      predicted_cr <- apply(predicted_cr_allsites, MARGIN= 2, FUN= mean, na.rm=T)
      predicted_vg <- apply(predicted_vg_allsites, MARGIN= 2, FUN= mean, na.rm=T)
      predicted_vg[which(predicted_vg<0)] <- 0                 #### to prevent negative variance
      predicted_rec <- mean(predicted_rec_allsites, na.rm=T)
      
      gmx <- matrix(NA, length(bin_mids),length(bin_mids)) ## create empty gmx before entering the no_bins loop
      
      # make the TSF-specific matrix
      for (ss in 1:no_bins) {
        growcdf <- pnorm(bin_edges,predicted_gr[ss],sqrt(predicted_vg[ss])) # Yes. Vargrowth=(pred - obs)^2
        grows <- growcdf[2:length(bin_edges)]-growcdf[1:(length(bin_edges)-1)]
        if(sum(grows)>0){grows <- grows/sum(grows)
        gmx[,ss] <- grows
        } else if (sum(grows)==0 & predicted_gr[ss]< min(bin_edges)) { 
          gmx[,ss] <- c(1, rep(0, length(bin_mids)-1))  
        } else if (sum(grows)==0 & predicted_gr[ss]> max(bin_edges)) { 
          gmx[,ss] <- c(rep(0, length(bin_mids)-1), 1)  } else { # deals with complete eviction-- OUT OF THE BOUNDS  of the kernel
            gmx[,ss] <- NA} 
        # this if statement breaks the code (puts NA's into the matrix) if the sum of the PDF is zero (which happens if all the probability is outside of the size bounds)
      } # end ss loop
      
      # make the surv*growth mx
      survgmx <- gmx*t(matrix( rep(predicted_sur,no_bins),no_bins)) # Yes, this is doing element-wise multiplication 
      reprow <-  predicted_sur*predicted_pr*predicted_cr # prob that you survive, reproduce, then number fruits|reproduce; this multiplcation does it element by element
      
      mx <- matrix(0, no_bins+1, no_bins+1)
      mx[2:(no_bins+1), 2:(no_bins+1)] = survgmx
      mx[1,2:(no_bins+1)] = reprow # FRUIT ROW
      
      sdlng_cdf <- pnorm(bin_edges,mean_seedling_logsize,sqrt(var_seedling_logsize)) 
      sdlng_pdf <- sdlng_cdf[2:length(bin_edges)]-sdlng_cdf[1:(length(bin_edges)-1)]
      if(sum(sdlng_pdf)>0){
        sdlng_pdf <- sdlng_pdf/sum(sdlng_pdf)} else {
          sdlng_pdf <- NA} 
      
      mx[2:(no_bins+1),1]  = predicted_rec *sdlng_pdf 
      
      mx_prod <- mx_prod %*% mx # in the first iteration of the loop, mx just becomes mx_prod
      
    } # closes ii (TSFs) loop
    
    j_pop_gr_rate_GSP_reg[i] <- (popbio::lambda(mx_prod))^(1/length(TSFs)) 
    # multiply them together to get a matrix, etc. 
  }
    j_pop_gr_rate_GSP_reg
}


all_pop_growth_rates_GSP_reg <- cbind(possible_scenarios, all_pop_growth_rates_GSP_reg)
names(all_pop_growth_rates_GSP_reg) <- c(names(possible_scenarios), paste("rep", 1:no_reps, sep = ""))
rep_cols <- grep("^rep", names(all_pop_growth_rates_GSP_reg), value = TRUE)

mean_lambda_across_scenarios_GSP_reg <- all_pop_growth_rates_GSP_reg %>%
  rowwise() %>%
  mutate(
    mean_lambda = mean(c_across(all_of(rep_cols)), na.rm = TRUE),
    lower_bound = quantile(c_across(all_of(rep_cols)), 0.025, na.rm = TRUE),
    upper_bound = quantile(c_across(all_of(rep_cols)), 0.975, na.rm = TRUE)
  ) %>%
  dplyr::select(FRI,s.TBF,mean_lambda,lower_bound, upper_bound) %>%
  mutate(TBF=NA) %>%
  ungroup()

mean_lambda_across_scenarios_GSP_reg$TBF[1:11] <- "Short TBF"
mean_lambda_across_scenarios_GSP_reg$TBF[12:22] <- "Long TBF"

##########. graph 
##########

plt6<- ggplot(data= mean_lambda_across_scenarios_GSP_reg, aes(x= FRI, y= mean_lambda)) +
  geom_line(aes(color= TBF)) + 
  geom_ribbon(aes(ymin = mean_lambda_across_scenarios_GSP_reg$lower_bound, ymax = mean_lambda_across_scenarios_GSP_reg$upper_bound, fill= TBF), alpha = 0.1) + 
  #ylim(0,1) +
  #xlim(c(0, 12)) +
  labs(y= "Growth Rate", x = "Fire Return Interval",color = NULL, fill = NULL) +
  theme_bw() + theme(legend.position='none') +
  theme(legend.position = "right") +
  scale_x_continuous(breaks = 0:10) +
  theme(text = element_text(size = 16)) 
plt6
```
##graph

```{r}
lambda_regionalIPM <- bind_rows(
  mean_lambda_across_scenarios %>% mutate(dataset = "global IPM"),
  mean_lambda_across_scenarios_Cro_reg %>% mutate(dataset = "Croatan"),
  mean_lambda_across_scenarios_GSP_reg %>% mutate(dataset = "Green Swamp"))

plt7 <- ggplot(lambda_regionalIPM, aes(x = FRI, y = mean_lambda)) +
    geom_ribbon(
    aes(ymin = lower_bound, ymax = upper_bound,
        fill = dataset, group = interaction(dataset, TBF)),
    alpha = 0.1) +
  geom_line(
    aes(color = dataset, linetype = TBF,
        group = interaction(dataset, TBF)),
    size = 1) +
  theme_bw() +
  labs(
    y = "Growth Rate",
    x = "Fire Return Interval",
    color = "Dataset",
    fill = "Dataset",
    linetype = "TBF") +
  scale_x_continuous(breaks = 0:10) +
  theme(text = element_text(size = 16))

plt7
```



