---
output:
  html_document: default
  pdf_document: default
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# load packages

```{r}
library(AICcmodavg)
library(lme4)
library(tidyverse)
library(dplyr)
library(glue)
library(zoo)
library(MASS)
library(foreach)
library(doParallel)
library(MuMIn)
```

# load necessary data

Here I'm loading a list of best-fit viral rate models as well as a master long-formatted demographic dataset containing all years and sites (produced by scripts in a previous project).

```{r}
load("model assessment/data/model_export_mar_2026.RData")    ## this environment contains vital rate functions plus the raw data used to build them.

```

# 1. global IPM

First, we want to get population growth rate aggregating data from all sites, which assume the assumptions of SFTS (space for time substitution).

First, using the models we just loaded, we'll first build an fire (TSF, TBF) and climate driven, deterministic IPM predicting long term population growth rate (lambda) against FRI (fire return interval) scenarios. This IPM will have 100 bins. Upper and lower bounds of the IPM are the upper and lower sizes observed across all years. Seedlings that enter the kernel all have the same distribution. In the code chunck below, I first set up the parameters of the IPM as mentioned above, then create a dataframe representing hypothetical FRI scenarios with or without a history of fire suppression; all other covariates were held constant at the means. We'll resample model parameter estimates (n=1000) to account for parameter uncertainty and ease the calculate of CI around lambda.

## 1.1 initiate IPM

```{r}
#size bounds for IPM
TBF_long$logsize0 <- (log(TBF_long$size0+0.1))
TBF_long$s.logsize0 <- scale(TBF_long$logsize0) 
TBF_long$s.TBF <- scale(TBF_long$TBF) 
TBF_long$s.TSF <- scale(TBF_long$TSF) 
TBF_long$sq.TSF <- (TBF_long$TSF)^2
TBF_long$s.sq.TSF <- scale(TBF_long$sq.TSF) 
IPM_bounds <- range(TBF_long$logsize0, na.rm = TRUE)  # our IPM will be based on logsize as a size metric. 
no_bins <- 100
bin_edges <- seq(IPM_bounds[1], IPM_bounds[2], length.out= no_bins+1)
bin_mids <- zoo::rollmean(bin_edges,2)

log.fr <- log(1.1)
# define mean and sd of seedling sizes
rows_with_new <- TBF_long[grepl("\\bnew", TBF_long$comm1), ]
rows_with_new <- rows_with_new[!grepl("no new", rows_with_new$comm1), ] # delete "no new"
rows_with_new <- rows_with_new[!grepl("no news", rows_with_new$comm1), ] # delete "no new" 
rows_with_new <- rows_with_new[!grepl("not possible to see new plants", rows_with_new$comm1), ]
rows_with_new <- rows_with_new[!grepl("No news", rows_with_new$comm1), ]
rows_with_new <- rows_with_new[!grepl("new lvs", rows_with_new$comm1), ]
rows_with_new <- rows_with_new[!grepl("new nail", rows_with_new$comm1), ]

mean_seedling_logsize <- mean(rows_with_new$logsize1, na.rm=T) # get the mean logsize of seedings   ### the mean seedling size is very large
var_seedling_logsize <- var(rows_with_new$logsize1,na.rm=T)# get the var in logsize of seedings  

### build empty matrices to store IPM output. Here, I'm interested in the effect of fire on lambda, so I'm setting average conditions for all other climate variables
possible_scenarios <- 
  as.data.frame(matrix(NA, nrow= 22, ncol = 28))
names(possible_scenarios) <- c("FRI","s.TBF","s.T_LA", "s.T_LH",
  "s.sq.T_LH",
  "s.sq.P_LA",
  "s.T_LC",
  "s.T_LD",
  "s.P_LD",
  "s.sq.T_LA",
  "s.P_LA",
  "s.P_LW",
  "s.P_LH",
  "s.T_RA",
  "s.sq.T_RA",
  "s.P_RA",
  "s.T_RH",
  "s.P_RH",
  "s.T_RC",
  "s.T_RW",
  "s.T_RD",
  "s.sq.T_RC",
  "s.P_RW",
  "s.P_RD",
  "s.sq.T_RH",
  "s.sq.P_RA",
  "s.sq.P_RW",
  "s.sq.P_RD")


possible_scenarios$FRI <- rep(1:11, 2) # we want FRI to range from 1-11
possible_scenarios$s.TBF <- scale(rep(c(0, 10), each= 11) ,# raw TBF values
                                  center= attr(TBF_long$s.TBF,"scaled:center"), 
                                  scale= attr(TBF_long$s.TBF,"scaled:scale")) # for each set of FRI's, we want to set a short TBF (of 3) and a long TBF (of 11)

possible_scenarios[, c(  "s.T_LA",  "s.T_LH",  "s.sq.T_LH",  "s.sq.P_LA",  "s.T_LC",  "s.T_LD",  "s.P_LD",  "s.sq.T_LA",  "s.P_LA",  
                         "s.P_LW",  "s.P_LH",  "s.T_RA",  "s.sq.T_RA",  "s.P_RA",  "s.T_RH",  "s.P_RH",  "s.T_RC",  "s.T_RW",  "s.T_RD",
                         "s.sq.T_RC","s.P_RW","s.P_RD", "s.sq.T_RH",  "s.sq.P_RA",  "s.sq.P_RW",  "s.sq.P_RD")] <- 0 # setting all climate variable to their mean

my_sites <- unique(TBF_long$site) 
no_reps <- 1000 # or somethign similar

########### extract model components for later predict() ###############
sur_model_terms <- terms(survival_mod)
sur_term <- attr(sur_model_terms, "term.labels")
sur_term <- reformulate(sur_term)

gr_model_terms <- terms(growth_mod)
gr_term <- attr(gr_model_terms, "term.labels")
gr_term <- reformulate(gr_term)

prep_model_terms <- terms(prep_mod)
prep_term <- attr(prep_model_terms, "term.labels")
prep_term <- reformulate(prep_term)

crep_model_terms <- terms(crep_mod)
crep_term <- attr(crep_model_terms, "term.labels")
crep_term <- reformulate(crep_term)

vargrowth_model_terms <- terms(vargrowth_mod)
vargrowth_term <- attr(vargrowth_model_terms, "term.labels")
vargrowth_term <- reformulate(vargrowth_term)

rec_model_terms <- terms(recruit_mod)
rec_term <- attr(rec_model_terms, "term.labels")
rec_term <- reformulate(rec_term)

##################################################################
###################### initiate IPM  ##############################
##################################################################
all_pop_growth_rates <- matrix(data= NA, nrow= dim(possible_scenarios)[1], ncol= no_reps)
n_cores <- detectCores()
cl <- makeCluster(n_cores - 1)
registerDoParallel(cl)
j_pop_gr_rate <- rep(NA, dim(possible_scenarios)[1]) ### changed to no_scenarios

all_pop_growth_rates <- foreach(
  j = 1:no_reps,
  .combine = cbind,
  .packages = c("MASS", "lme4", "popbio", "dplyr")
) %dopar% {
  rsur_coefs <- coefficients(survival_mod) # first, assign all "randomy selected" coefficients to the mean estimated coefficeints
  rsur_coefs[!names(coefficients(survival_mod)) %in% paste("site", my_sites, sep="")] <- MASS::mvrnorm(mu= coefficients(survival_mod), Sigma = vcov(survival_mod))[!names(coefficients(survival_mod)) %in% paste("site", my_sites, sep="")]
  #rsur_coefs <- lme4::fixef(survival_mod)
  #rsur_coefs <-  MASS::mvrnorm(mu= lme4::fixef(survival_mod) , Sigma = vcov(survival_mod))
  rgr_coefs <- lme4::fixef(growth_mod)
  rgr_coefs <-  MASS::mvrnorm(mu= lme4::fixef(growth_mod) , Sigma = vcov(growth_mod)) # no site effect in this model
  rprep_coefs <- coefficients(prep_mod)
  rprep_coefs[!names(coefficients(prep_mod)) %in% paste("site", my_sites, sep="")] <- MASS::mvrnorm(mu= coefficients(prep_mod), Sigma = vcov(prep_mod))[!names(coefficients(prep_mod)) %in% paste("site", my_sites, sep="")]
  rcrep_coefs <- lme4::fixef(crep_mod)
  rcrep_coefs <-  MASS::mvrnorm(mu= lme4::fixef(crep_mod) , Sigma = vcov(crep_mod)) # no site effect in this model
  rvar_coefs <- lme4::fixef(vargrowth_mod)
  rvar_coefs <-  MASS::mvrnorm(mu= lme4::fixef(vargrowth_mod) , Sigma = vcov(vargrowth_mod)) # no site effect in this model
  rrec_coefs <- lme4::fixef(recruit_mod)
  rrec_coefs <-  MASS::mvrnorm(mu= lme4::fixef(recruit_mod) , Sigma = vcov(recruit_mod)) # no site effect in this model
  #rrec_coefs <-  MASS::mvrnorm(mu= lme4::fixef(recruit_mod) , Sigma = vcov(recruit_mod)) 
  
  for (i in 1: dim(possible_scenarios)[1]) {
    
    i_scenario <- possible_scenarios[i,] # which combination of FRI, TBF, and climate do I want to use
    
    TSFs <- 0:(i_scenario$FRI-1) # TSF range from 0 to (FRI-1) such that an FRI of 1 will always have TSF= 0
    i_scenario <- i_scenario[rep(seq_len(nrow(i_scenario)), length(TSFs)), ]
    i_scenario$TSF <- TSFs
    i_scenario$s.TSF <- scale(TSFs, center= attr(TBF_long$s.TSF,"scaled:center"), 
                              scale= attr(TBF_long$s.TSF,"scaled:scale"))
    i_scenario$s.sq.TSF <- scale(TSFs^2, center= attr(TBF_long$s.sq.TSF,"scaled:center"), 
                                 scale= attr(TBF_long$s.sq.TSF,"scaled:scale"))
    
    mx_prod <- diag(no_bins+1) # setting up an identity matrix before you enter the loop
    
    for (ii in TSFs){
      i_ii_scenario <- i_scenario[which(i_scenario$TSF==ii),]
      data_for_prediction <- i_ii_scenario[rep(seq_len(nrow(i_ii_scenario)), length(bin_mids)), ]
      data_for_prediction$s.logsize0 <-   scale(bin_mids, center= attr(TBF_long$s.logsize0,"scaled:center"), 
                                                scale= attr(TBF_long$s.logsize0,"scaled:scale"))
      data_for_prediction_rec <- cbind(i_ii_scenario)
      
      predicted_sur <- predicted_gr <-predicted_pr <- predicted_cr <- predicted_vg <- matrix(data=NA, nrow= length(my_sites), ncol= length(bin_mids))
      predicted_rec <- rep(NA, length(my_sites))
      
      ### make empty mxes before entering the site loop
      
      predicted_sur_allsites <-predicted_gr_allsites<-predicted_pr_allsites <-predicted_cr_allsites<-predicted_vg_allsites <-matrix(NA, nrow = length(my_sites), ncol = no_bins, dimnames = list(my_sites, NULL))
      predicted_rec_allsites <- matrix(NA, nrow = length(my_sites), ncol = 1, dimnames = list(my_sites, NULL))
      
      
      for (s in my_sites){ #looping through the sites
        s_data_for_prediction <- data_for_prediction
        s_data_for_prediction_rec <- data_for_prediction_rec
        s_data_for_prediction$site <- s
        s_data_for_prediction_rec$site <- s
        
        s_data_for_prediction$site     <- factor(s_data_for_prediction$site,  levels = my_sites)   ## it is important to add the levels=my_sites here to ensure that the model matrix is built correctly
        s_data_for_prediction_rec$site <- factor(s_data_for_prediction_rec$site, levels = my_sites)
        
        s_data_for_prediction <- s_data_for_prediction %>% mutate(across(c("s.logsize0", "s.sq.TSF", "s.TSF", "s.TBF"), as.numeric))   ### change col type to numeric to avoid matrix col type
        # Zheng-- you will have to change the predict() call in the below to be a custom predict function that uses the random coefficients generated earlier in the code (e.g., rsur_coefs)
        
        ### build model matrices and use mx multiplication to obtain predicted values. Don't forget to transform if needed!!!!!!!!!
        
        m_sur <- model.matrix(sur_term ,data=s_data_for_prediction)
        p2_sur <- rsur_coefs %*% t(m_sur) # check on whether this coef(survival_mod is a column or a row!!!!!!!!
        p2_sur <- plogis(p2_sur)   ## transform to binomial 
        predicted_sur_allsites[s,] <- p2_sur
        
        m_gr <- model.matrix(gr_term, data=s_data_for_prediction)
        p2_gr <- rgr_coefs %*% t(m_gr) # check on whether this coef(survival_mod is a column or a row!!!!!!!!
        predicted_gr_allsites[s,] <- p2_gr
        
        m_prep <- model.matrix(prep_term,data=s_data_for_prediction)
        p2_prep <- rprep_coefs %*% t(m_prep) # check on whether this coef(survival_mod is a column or a row!!!!!!!!
        p2_prep <- plogis(p2_prep)   ## transform to binomial 
        predicted_pr_allsites[s,] <- p2_prep
        
        m_crep <- model.matrix(crep_term  ,data=s_data_for_prediction)
        p2_crep <- rcrep_coefs %*% t(m_crep) # check on whether this coef(survival_mod is a column or a row!!!!!!!!
        predicted_cr_allsites[s,] <- p2_crep
        
        m_vg <- model.matrix(vargrowth_term,data=s_data_for_prediction)
        p2_vg <- rvar_coefs %*% t(m_vg) # 
        predicted_vg_allsites[s,] <- exp(p2_vg)  
        
        m_rec <- model.matrix(rec_term , data=s_data_for_prediction_rec)
        p2_rec <- rrec_coefs %*% t(m_rec) # check on whether this coef(survival_mod is a column or a row!!!!!!!!
        predicted_rec_allsites[s,1] <- p2_rec*log(1.1)            #### account for composite rec variable (rec/log.fr)
        
        
        # for mixed models, the above functions will predict site-specific results (i.e., will incorporate random effects)
        rm(s_data_for_prediction) # this part removes the site-specific dataframe used in this loop, so you do not accidentaly use the previous sites' dataframe instead of the current one-- in case there was an error somehow
      }
      predicted_sur <- apply(predicted_sur_allsites, MARGIN= 2, FUN= mean) # average size-specific vital rates across sites
      predicted_gr <- apply(predicted_gr_allsites, MARGIN= 2, FUN= mean)
      predicted_pr <- apply(predicted_pr_allsites, MARGIN= 2, FUN= mean)
      predicted_cr <- apply(predicted_cr_allsites, MARGIN= 2, FUN= mean)
      predicted_vg <- apply(predicted_vg_allsites, MARGIN= 2, FUN= mean)
      predicted_vg[which(predicted_vg<0)] <- 0                 #### to prevent negative variance
      predicted_rec <- mean(predicted_rec_allsites)
      
      gmx <- matrix(NA, length(bin_mids),length(bin_mids)) ## create empty gmx before entering the no_bins loop
      
      # make the TSF-specific matrix
      for (ss in 1:no_bins) {
        growcdf <- pnorm(bin_edges,predicted_gr[ss],sqrt(predicted_vg[ss])) # Yes. Vargrowth=(pred - obs)^2
        grows <- growcdf[2:length(bin_edges)]-growcdf[1:(length(bin_edges)-1)]
        if(sum(grows)>0){grows <- grows/sum(grows)
        gmx[,ss] <- grows
        } else if (sum(grows)==0 & grow_vals[ss]< min(bin_edges)) { 
          gmx[,ss] <- c(1, rep(0, length(binmids)-1))  
        } else if (sum(grows)==0 & grow_vals[ss]> max(bin_edges)) { 
          gmx[,ss] <- c(rep(0, length(binmids)-1), 1)  } else { # deals with complete eviction-- OUT OF THE BOUNDS  of the kernel
            gmx[,ss] <- NA} 
        # this if statement breaks the code (puts NA's into the matrix) if the sum of the PDF is zero (which happens if all the probability is outside of the size bounds)
      } # end ss loop
      
      # make the surv*growth mx
      survgmx <- gmx*t(matrix( rep(predicted_sur,no_bins),no_bins)) # Yes, this is doing element-wise multiplication 
      reprow <-  predicted_sur*predicted_pr*predicted_cr # prob that you survive, reproduce, then number fruits|reproduce; this multiplcation does it element by element
      
      mx <- matrix(0, no_bins+1, no_bins+1)
      mx[2:(no_bins+1), 2:(no_bins+1)] = survgmx
      mx[1,2:(no_bins+1)] = reprow # FRUIT ROW
      
      sdlng_cdf <- pnorm(bin_edges,mean_seedling_logsize,sqrt(var_seedling_logsize)) 
      sdlng_pdf <- sdlng_cdf[2:length(bin_edges)]-sdlng_cdf[1:(length(bin_edges)-1)]
      if(sum(sdlng_pdf)>0){
        sdlng_pdf <- sdlng_pdf/sum(sdlng_pdf)} else {
          sdlng_pdf <- NA} 
      
      mx[2:(no_bins+1),1]  = predicted_rec *sdlng_pdf 
      
      mx_prod <- mx_prod %*% mx # in the first iteration of the loop, mx just becomes mx_prod
      
    } # closes ii (TSFs) loop
    
    j_pop_gr_rate[i] <- (popbio::lambda(mx_prod))^(1/length(TSFs)) 
    # multiply them together to get a matrix, etc. 
  }
    j_pop_gr_rate
  
}

```

## 1.2 graph lambda \~ FRI, colored by TBF

This is the relationship observed using aggregated data (Space for time substitution)

```{r}
######################################
###################################### global IPM results, rep=1000 
######################################

all_pop_growth_rates <- cbind(possible_scenarios, all_pop_growth_rates)
names(all_pop_growth_rates) <- c(names(possible_scenarios), paste("rep", 1:no_reps, sep = ""))

rep_cols <- grep("^rep", names(all_pop_growth_rates), value = TRUE)

mean_lambda_across_scenarios <- all_pop_growth_rates %>%
  rowwise() %>%
  mutate(
    mean_lambda = mean(c_across(all_of(rep_cols)), na.rm = TRUE),
    lower_bound = quantile(c_across(all_of(rep_cols)), 0.025, na.rm = TRUE),
    upper_bound = quantile(c_across(all_of(rep_cols)), 0.975, na.rm = TRUE)
  ) %>%
  dplyr::select(FRI,s.TBF,mean_lambda,lower_bound, upper_bound) %>%
  mutate(TBF=NA) %>%
  ungroup()

mean_lambda_across_scenarios$TBF[1:11] <- "Short TBF"
mean_lambda_across_scenarios$TBF[12:22] <- "Long TBF"

##########. graph 
##########

plt1 <- ggplot(data= mean_lambda_across_scenarios, aes(x= FRI, y= mean_lambda)) +
  geom_line(aes(color= TBF)) + 
  geom_ribbon(aes(ymin = mean_lambda_across_scenarios$lower_bound, ymax = mean_lambda_across_scenarios$upper_bound, fill= TBF), alpha = 0.1) + 
  #ylim(0,1) +
  #xlim(c(0, 12)) +
  labs(y= "Growth Rate", x = "Fire Return Interval",color = NULL, fill = NULL) +
  theme_bw() + theme(legend.position='none') +
  theme(legend.position = "right") +
  scale_x_continuous(breaks = 0:10) +
  theme(text = element_text(size = 16)) 
plt1
```

## 1.3 Use the global IPM to predict region level responses

we can use the global IPM to make region specific inference, though variations among site are driven exclusively by fixed site intercepts.

#### 1.3.1 Croatan

The Croatan region contains CH and CM. Both sites had history of fire suppression

```{r}

all_pop_growth_rates_Cro <- matrix(data= NA, nrow= dim(possible_scenarios)[1], ncol= no_reps)
n_cores <- detectCores()
cl <- makeCluster(n_cores - 1)
registerDoParallel(cl)
j_pop_gr_rate_Cro <- rep(NA, dim(possible_scenarios)[1]) ### changed to no_scenarios

all_pop_growth_rates_Cro <- foreach(
  j = 1:no_reps,
  .combine = cbind,
  .packages = c("MASS", "lme4", "popbio", "dplyr")
) %dopar% {
  rsur_coefs <- coefficients(survival_mod) # first, assign all "randomy selected" coefficients to the mean estimated coefficeints
  rsur_coefs[!names(coefficients(survival_mod)) %in% paste("site", my_sites, sep="")] <- MASS::mvrnorm(mu= coefficients(survival_mod), Sigma = vcov(survival_mod))[!names(coefficients(survival_mod)) %in% paste("site", my_sites, sep="")]
  #rsur_coefs <- lme4::fixef(survival_mod)
  #rsur_coefs <-  MASS::mvrnorm(mu= lme4::fixef(survival_mod) , Sigma = vcov(survival_mod))
  rgr_coefs <- lme4::fixef(growth_mod)
  rgr_coefs <-  MASS::mvrnorm(mu= lme4::fixef(growth_mod) , Sigma = vcov(growth_mod)) # no site effect in this model
  rprep_coefs <- coefficients(prep_mod)
  rprep_coefs[!names(coefficients(prep_mod)) %in% paste("site", my_sites, sep="")] <- MASS::mvrnorm(mu= coefficients(prep_mod), Sigma = vcov(prep_mod))[!names(coefficients(prep_mod)) %in% paste("site", my_sites, sep="")]
  rcrep_coefs <- lme4::fixef(crep_mod)
  rcrep_coefs <-  MASS::mvrnorm(mu= lme4::fixef(crep_mod) , Sigma = vcov(crep_mod)) # no site effect in this model
  rvar_coefs <- lme4::fixef(vargrowth_mod)
  rvar_coefs <-  MASS::mvrnorm(mu= lme4::fixef(vargrowth_mod) , Sigma = vcov(vargrowth_mod)) # no site effect in this model
  rrec_coefs <- lme4::fixef(recruit_mod)
  rrec_coefs <-  MASS::mvrnorm(mu= lme4::fixef(recruit_mod) , Sigma = vcov(recruit_mod)) # no site effect in this model
  #rrec_coefs <-  MASS::mvrnorm(mu= lme4::fixef(recruit_mod) , Sigma = vcov(recruit_mod)) 
  
  for (i in 1: dim(possible_scenarios)[1]) {
    
    i_scenario <- possible_scenarios[i,] # which combination of FRI, TBF, and climate do I want to use
    
    TSFs <- 0:(i_scenario$FRI-1) # TSF range from 0 to (FRI-1) such that an FRI of 1 will always have TSF= 0
    i_scenario <- i_scenario[rep(seq_len(nrow(i_scenario)), length(TSFs)), ]
    i_scenario$TSF <- TSFs
    i_scenario$s.TSF <- scale(TSFs, center= attr(TBF_long$s.TSF,"scaled:center"), 
                              scale= attr(TBF_long$s.TSF,"scaled:scale"))
    i_scenario$s.sq.TSF <- scale(TSFs^2, center= attr(TBF_long$s.sq.TSF,"scaled:center"), 
                                 scale= attr(TBF_long$s.sq.TSF,"scaled:scale"))
    
    mx_prod <- diag(no_bins+1) # setting up an identity matrix before you enter the loop
    
    for (ii in TSFs){
      i_ii_scenario <- i_scenario[which(i_scenario$TSF==ii),]
      data_for_prediction <- i_ii_scenario[rep(seq_len(nrow(i_ii_scenario)), length(bin_mids)), ]
      data_for_prediction$s.logsize0 <-   scale(bin_mids, center= attr(TBF_long$s.logsize0,"scaled:center"), 
                                                scale= attr(TBF_long$s.logsize0,"scaled:scale"))
      data_for_prediction_rec <- cbind(i_ii_scenario)
      
      predicted_sur <- predicted_gr <-predicted_pr <- predicted_cr <- predicted_vg <- matrix(data=NA, nrow= length(my_sites), ncol= length(bin_mids))
      predicted_rec <- rep(NA, length(my_sites))
      
      ### make empty mxes before entering the site loop
      
      predicted_sur_allsites <-predicted_gr_allsites<-predicted_pr_allsites <-predicted_cr_allsites<-predicted_vg_allsites <-matrix(NA, nrow = length(my_sites), ncol = no_bins, dimnames = list(my_sites, NULL))
      predicted_rec_allsites <- matrix(NA, nrow = length(my_sites), ncol = 1, dimnames = list(my_sites, NULL))
      
      
      for (s in my_sites[3:4]){ #looping through the sites
        s_data_for_prediction <- data_for_prediction
        s_data_for_prediction_rec <- data_for_prediction_rec
        s_data_for_prediction$site <- s
        s_data_for_prediction_rec$site <- s
        
        s_data_for_prediction$site     <- factor(s_data_for_prediction$site,  levels = my_sites)   ## it is important to add the levels=my_sites here to ensure that the model matrix is built correctly
        s_data_for_prediction_rec$site <- factor(s_data_for_prediction_rec$site, levels = my_sites)
        
        s_data_for_prediction <- s_data_for_prediction %>% mutate(across(c("s.logsize0", "s.sq.TSF", "s.TSF", "s.TBF"), as.numeric))   ### change col type to numeric to avoid matrix col type
        # Zheng-- you will have to change the predict() call in the below to be a custom predict function that uses the random coefficients generated earlier in the code (e.g., rsur_coefs)
        
        ### build model matrices and use mx multiplication to obtain predicted values. Don't forget to transform if needed!!!!!!!!!
        
        m_sur <- model.matrix(sur_term ,data=s_data_for_prediction)
        p2_sur <- rsur_coefs %*% t(m_sur) # check on whether this coef(survival_mod is a column or a row!!!!!!!!
        p2_sur <- plogis(p2_sur)   ## transform to binomial 
        predicted_sur_allsites[s,] <- p2_sur
        
        m_gr <- model.matrix(gr_term, data=s_data_for_prediction)
        p2_gr <- rgr_coefs %*% t(m_gr) # check on whether this coef(survival_mod is a column or a row!!!!!!!!
        predicted_gr_allsites[s,] <- p2_gr
        
        m_prep <- model.matrix(prep_term,data=s_data_for_prediction)
        p2_prep <- rprep_coefs %*% t(m_prep) # check on whether this coef(survival_mod is a column or a row!!!!!!!!
        p2_prep <- plogis(p2_prep)   ## transform to binomial 
        predicted_pr_allsites[s,] <- p2_prep
        
        m_crep <- model.matrix(crep_term  ,data=s_data_for_prediction)
        p2_crep <- rcrep_coefs %*% t(m_crep) # check on whether this coef(survival_mod is a column or a row!!!!!!!!
        predicted_cr_allsites[s,] <- p2_crep
        
        m_vg <- model.matrix(vargrowth_term,data=s_data_for_prediction)
        p2_vg <- rvar_coefs %*% t(m_vg) # 
        predicted_vg_allsites[s,] <- exp(p2_vg)  
        
        m_rec <- model.matrix(rec_term , data=s_data_for_prediction_rec)
        p2_rec <- rrec_coefs %*% t(m_rec) # check on whether this coef(survival_mod is a column or a row!!!!!!!!
        predicted_rec_allsites[s,1] <- p2_rec*log(1.1)            #### account for composite rec variable (rec/log.fr)
        
        
        # for mixed models, the above functions will predict site-specific results (i.e., will incorporate random effects)
        rm(s_data_for_prediction) # this part removes the site-specific dataframe used in this loop, so you do not accidentaly use the previous sites' dataframe instead of the current one-- in case there was an error somehow
      }
      predicted_sur <- apply(predicted_sur_allsites, MARGIN= 2, FUN= mean,na.rm = TRUE) # average size-specific vital rates across sites
      predicted_gr <- apply(predicted_gr_allsites, MARGIN= 2, FUN= mean,na.rm = TRUE)
      predicted_pr <- apply(predicted_pr_allsites, MARGIN= 2, FUN= mean,na.rm = TRUE)
      predicted_cr <- apply(predicted_cr_allsites, MARGIN= 2, FUN= mean,na.rm = TRUE)
      predicted_vg <- apply(predicted_vg_allsites, MARGIN= 2, FUN= mean,na.rm = TRUE)
      predicted_vg[which(predicted_vg<0)] <- 0                 #### to prevent negative variance
      predicted_rec <- mean(predicted_rec_allsites,na.rm = TRUE)
      
      gmx <- matrix(NA, length(bin_mids),length(bin_mids)) ## create empty gmx before entering the no_bins loop
      
      # make the TSF-specific matrix
      for (ss in 1:no_bins) {
        growcdf <- pnorm(bin_edges,predicted_gr[ss],sqrt(predicted_vg[ss])) # Yes. Vargrowth=(pred - obs)^2
        grows <- growcdf[2:length(bin_edges)]-growcdf[1:(length(bin_edges)-1)]
        if(sum(grows)>0){grows <- grows/sum(grows)
        gmx[,ss] <- grows
        } else if (sum(grows)==0 & grow_vals[ss]< min(bin_edges)) { 
          gmx[,ss] <- c(1, rep(0, length(binmids)-1))  
        } else if (sum(grows)==0 & grow_vals[ss]> max(bin_edges)) { 
          gmx[,ss] <- c(rep(0, length(binmids)-1), 1)  } else { # deals with complete eviction-- OUT OF THE BOUNDS  of the kernel
            gmx[,ss] <- NA} 
        # this if statement breaks the code (puts NA's into the matrix) if the sum of the PDF is zero (which happens if all the probability is outside of the size bounds)
      } # end ss loop
      
      # make the surv*growth mx
      survgmx <- gmx*t(matrix( rep(predicted_sur,no_bins),no_bins)) # Yes, this is doing element-wise multiplication 
      reprow <-  predicted_sur*predicted_pr*predicted_cr # prob that you survive, reproduce, then number fruits|reproduce; this multiplcation does it element by element
      
      mx <- matrix(0, no_bins+1, no_bins+1)
      mx[2:(no_bins+1), 2:(no_bins+1)] = survgmx
      mx[1,2:(no_bins+1)] = reprow # FRUIT ROW
      
      sdlng_cdf <- pnorm(bin_edges,mean_seedling_logsize,sqrt(var_seedling_logsize)) 
      sdlng_pdf <- sdlng_cdf[2:length(bin_edges)]-sdlng_cdf[1:(length(bin_edges)-1)]
      if(sum(sdlng_pdf)>0){
        sdlng_pdf <- sdlng_pdf/sum(sdlng_pdf)} else {
          sdlng_pdf <- NA} 
      
      mx[2:(no_bins+1),1]  = predicted_rec *sdlng_pdf 
      
      mx_prod <- mx_prod %*% mx # in the first iteration of the loop, mx just becomes mx_prod
      
    } # closes ii (TSFs) loop
    
    j_pop_gr_rate_Cro[i] <- (popbio::lambda(mx_prod))^(1/length(TSFs)) 
    # multiply them together to get a matrix, etc. 
  }
    j_pop_gr_rate_Cro
}


all_pop_growth_rates_Cro <- cbind(possible_scenarios, all_pop_growth_rates_Cro)
names(all_pop_growth_rates_Cro) <- c(names(possible_scenarios), paste("rep", 1:no_reps, sep = ""))
rep_cols <- grep("^rep", names(all_pop_growth_rates_Cro), value = TRUE)

mean_lambda_across_scenarios_Cro <- all_pop_growth_rates_Cro %>%
  rowwise() %>%
  mutate(
    mean_lambda = mean(c_across(all_of(rep_cols)), na.rm = TRUE),
    lower_bound = quantile(c_across(all_of(rep_cols)), 0.025, na.rm = TRUE),
    upper_bound = quantile(c_across(all_of(rep_cols)), 0.975, na.rm = TRUE)
  ) %>%
  dplyr::select(FRI,s.TBF,mean_lambda,lower_bound, upper_bound) %>%
  mutate(TBF=NA) %>%
  ungroup()

mean_lambda_across_scenarios_Cro$TBF[1:11] <- "Short TBF"
mean_lambda_across_scenarios_Cro$TBF[12:22] <- "Long TBF"

##########. graph 
##########

plt2 <- ggplot(data= mean_lambda_across_scenarios_Cro, aes(x= FRI, y= mean_lambda)) +
  geom_line(aes(color= TBF)) + 
  geom_ribbon(aes(ymin = mean_lambda_across_scenarios_Cro$lower_bound, ymax = mean_lambda_across_scenarios_Cro$upper_bound, fill= TBF), alpha = 0.1) + 
  #ylim(0,1) +
  #xlim(c(0, 12)) +
  labs(y= "Growth Rate", x = "Fire Return Interval",color = NULL, fill = NULL) +
  theme_bw() + theme(legend.position='none') +
  theme(legend.position = "right") +
  scale_x_continuous(breaks = 0:10) +
  theme(text = element_text(size = 16)) 
plt2


```

#### 1.3.2 Green Swamps (no history of fire suppression)

The GSP (Green Swamp) region contains GSP-BI and GSP-LI, both sites were subject to fire management and didnot have histories of fire suppression.

```{r}
all_pop_growth_rates_GSP <- matrix(data= NA, nrow= dim(possible_scenarios)[1], ncol= no_reps)
n_cores <- detectCores()
cl <- makeCluster(n_cores - 1)
registerDoParallel(cl)
j_pop_gr_rate_GSP <- rep(NA, dim(possible_scenarios)[1]) ### changed to no_scenarios

all_pop_growth_rates_GSP <- foreach(
  j = 1:no_reps,
  .combine = cbind,
  .packages = c("MASS", "lme4", "popbio", "dplyr")
) %dopar% {
  rsur_coefs <- coefficients(survival_mod) # first, assign all "randomy selected" coefficients to the mean estimated coefficeints
  rsur_coefs[!names(coefficients(survival_mod)) %in% paste("site", my_sites, sep="")] <- MASS::mvrnorm(mu= coefficients(survival_mod), Sigma = vcov(survival_mod))[!names(coefficients(survival_mod)) %in% paste("site", my_sites, sep="")]
  #rsur_coefs <- lme4::fixef(survival_mod)
  #rsur_coefs <-  MASS::mvrnorm(mu= lme4::fixef(survival_mod) , Sigma = vcov(survival_mod))
  rgr_coefs <- lme4::fixef(growth_mod)
  rgr_coefs <-  MASS::mvrnorm(mu= lme4::fixef(growth_mod) , Sigma = vcov(growth_mod)) # no site effect in this model
  rprep_coefs <- coefficients(prep_mod)
  rprep_coefs[!names(coefficients(prep_mod)) %in% paste("site", my_sites, sep="")] <- MASS::mvrnorm(mu= coefficients(prep_mod), Sigma = vcov(prep_mod))[!names(coefficients(prep_mod)) %in% paste("site", my_sites, sep="")]
  rcrep_coefs <- lme4::fixef(crep_mod)
  rcrep_coefs <-  MASS::mvrnorm(mu= lme4::fixef(crep_mod) , Sigma = vcov(crep_mod)) # no site effect in this model
  rvar_coefs <- lme4::fixef(vargrowth_mod)
  rvar_coefs <-  MASS::mvrnorm(mu= lme4::fixef(vargrowth_mod) , Sigma = vcov(vargrowth_mod)) # no site effect in this model
  rrec_coefs <- lme4::fixef(recruit_mod)
  rrec_coefs <-  MASS::mvrnorm(mu= lme4::fixef(recruit_mod) , Sigma = vcov(recruit_mod)) # no site effect in this model
  #rrec_coefs <-  MASS::mvrnorm(mu= lme4::fixef(recruit_mod) , Sigma = vcov(recruit_mod)) 
  
  for (i in 1: dim(possible_scenarios)[1]) {
    
    i_scenario <- possible_scenarios[i,] # which combination of FRI, TBF, and climate do I want to use
    
    TSFs <- 0:(i_scenario$FRI-1) # TSF range from 0 to (FRI-1) such that an FRI of 1 will always have TSF= 0
    i_scenario <- i_scenario[rep(seq_len(nrow(i_scenario)), length(TSFs)), ]
    i_scenario$TSF <- TSFs
    i_scenario$s.TSF <- scale(TSFs, center= attr(TBF_long$s.TSF,"scaled:center"), 
                              scale= attr(TBF_long$s.TSF,"scaled:scale"))
    i_scenario$s.sq.TSF <- scale(TSFs^2, center= attr(TBF_long$s.sq.TSF,"scaled:center"), 
                                 scale= attr(TBF_long$s.sq.TSF,"scaled:scale"))
    
    mx_prod <- diag(no_bins+1) # setting up an identity matrix before you enter the loop
    
    for (ii in TSFs){
      i_ii_scenario <- i_scenario[which(i_scenario$TSF==ii),]
      data_for_prediction <- i_ii_scenario[rep(seq_len(nrow(i_ii_scenario)), length(bin_mids)), ]
      data_for_prediction$s.logsize0 <-   scale(bin_mids, center= attr(TBF_long$s.logsize0,"scaled:center"), 
                                                scale= attr(TBF_long$s.logsize0,"scaled:scale"))
      data_for_prediction_rec <- cbind(i_ii_scenario)
      
      predicted_sur <- predicted_gr <-predicted_pr <- predicted_cr <- predicted_vg <- matrix(data=NA, nrow= length(my_sites), ncol= length(bin_mids))
      predicted_rec <- rep(NA, length(my_sites))
      
      ### make empty mxes before entering the site loop
      
      predicted_sur_allsites <-predicted_gr_allsites<-predicted_pr_allsites <-predicted_cr_allsites<-predicted_vg_allsites <-matrix(NA, nrow = length(my_sites), ncol = no_bins, dimnames = list(my_sites, NULL))
      predicted_rec_allsites <- matrix(NA, nrow = length(my_sites), ncol = 1, dimnames = list(my_sites, NULL))
      
      
      for (s in my_sites[5:6]){ #looping through the sites
        s_data_for_prediction <- data_for_prediction
        s_data_for_prediction_rec <- data_for_prediction_rec
        s_data_for_prediction$site <- s
        s_data_for_prediction_rec$site <- s
        
        s_data_for_prediction$site     <- factor(s_data_for_prediction$site,  levels = my_sites)   ## it is important to add the levels=my_sites here to ensure that the model matrix is built correctly
        s_data_for_prediction_rec$site <- factor(s_data_for_prediction_rec$site, levels = my_sites)
        
        s_data_for_prediction <- s_data_for_prediction %>% mutate(across(c("s.logsize0", "s.sq.TSF", "s.TSF", "s.TBF"), as.numeric))   ### change col type to numeric to avoid matrix col type
        # Zheng-- you will have to change the predict() call in the below to be a custom predict function that uses the random coefficients generated earlier in the code (e.g., rsur_coefs)
        
        ### build model matrices and use mx multiplication to obtain predicted values. Don't forget to transform if needed!!!!!!!!!
        
        m_sur <- model.matrix(sur_term ,data=s_data_for_prediction)
        p2_sur <- rsur_coefs %*% t(m_sur) # check on whether this coef(survival_mod is a column or a row!!!!!!!!
        p2_sur <- plogis(p2_sur)   ## transform to binomial 
        predicted_sur_allsites[s,] <- p2_sur
        
        m_gr <- model.matrix(gr_term, data=s_data_for_prediction)
        p2_gr <- rgr_coefs %*% t(m_gr) # check on whether this coef(survival_mod is a column or a row!!!!!!!!
        predicted_gr_allsites[s,] <- p2_gr
        
        m_prep <- model.matrix(prep_term,data=s_data_for_prediction)
        p2_prep <- rprep_coefs %*% t(m_prep) # check on whether this coef(survival_mod is a column or a row!!!!!!!!
        p2_prep <- plogis(p2_prep)   ## transform to binomial 
        predicted_pr_allsites[s,] <- p2_prep
        
        m_crep <- model.matrix(crep_term  ,data=s_data_for_prediction)
        p2_crep <- rcrep_coefs %*% t(m_crep) # check on whether this coef(survival_mod is a column or a row!!!!!!!!
        predicted_cr_allsites[s,] <- p2_crep
        
        m_vg <- model.matrix(vargrowth_term,data=s_data_for_prediction)
        p2_vg <- rvar_coefs %*% t(m_vg) # 
        predicted_vg_allsites[s,] <- exp(p2_vg)  
        
        m_rec <- model.matrix(rec_term , data=s_data_for_prediction_rec)
        p2_rec <- rrec_coefs %*% t(m_rec) # check on whether this coef(survival_mod is a column or a row!!!!!!!!
        predicted_rec_allsites[s,1] <- p2_rec*log(1.1)            #### account for composite rec variable (rec/log.fr)
        
        
        # for mixed models, the above functions will predict site-specific results (i.e., will incorporate random effects)
        rm(s_data_for_prediction) # this part removes the site-specific dataframe used in this loop, so you do not accidentaly use the previous sites' dataframe instead of the current one-- in case there was an error somehow
      }
      predicted_sur <- apply(predicted_sur_allsites, MARGIN= 2, FUN= mean,na.rm = TRUE) # average size-specific vital rates across sites
      predicted_gr <- apply(predicted_gr_allsites, MARGIN= 2, FUN= mean,na.rm = TRUE)
      predicted_pr <- apply(predicted_pr_allsites, MARGIN= 2, FUN= mean,na.rm = TRUE)
      predicted_cr <- apply(predicted_cr_allsites, MARGIN= 2, FUN= mean,na.rm = TRUE)
      predicted_vg <- apply(predicted_vg_allsites, MARGIN= 2, FUN= mean,na.rm = TRUE)
      predicted_vg[which(predicted_vg<0)] <- 0                 #### to prevent negative variance
      predicted_rec <- mean(predicted_rec_allsites,na.rm = TRUE)
      
      gmx <- matrix(NA, length(bin_mids),length(bin_mids)) ## create empty gmx before entering the no_bins loop
      
      # make the TSF-specific matrix
      for (ss in 1:no_bins) {
        growcdf <- pnorm(bin_edges,predicted_gr[ss],sqrt(predicted_vg[ss])) # Yes. Vargrowth=(pred - obs)^2
        grows <- growcdf[2:length(bin_edges)]-growcdf[1:(length(bin_edges)-1)]
        if(sum(grows)>0){grows <- grows/sum(grows)
        gmx[,ss] <- grows
        } else if (sum(grows)==0 & grow_vals[ss]< min(bin_edges)) { 
          gmx[,ss] <- c(1, rep(0, length(binmids)-1))  
        } else if (sum(grows)==0 & grow_vals[ss]> max(bin_edges)) { 
          gmx[,ss] <- c(rep(0, length(binmids)-1), 1)  } else { # deals with complete eviction-- OUT OF THE BOUNDS  of the kernel
            gmx[,ss] <- NA} 
        # this if statement breaks the code (puts NA's into the matrix) if the sum of the PDF is zero (which happens if all the probability is outside of the size bounds)
      } # end ss loop
      
      # make the surv*growth mx
      survgmx <- gmx*t(matrix( rep(predicted_sur,no_bins),no_bins)) # Yes, this is doing element-wise multiplication 
      reprow <-  predicted_sur*predicted_pr*predicted_cr # prob that you survive, reproduce, then number fruits|reproduce; this multiplcation does it element by element
      
      mx <- matrix(0, no_bins+1, no_bins+1)
      mx[2:(no_bins+1), 2:(no_bins+1)] = survgmx
      mx[1,2:(no_bins+1)] = reprow # FRUIT ROW
      
      sdlng_cdf <- pnorm(bin_edges,mean_seedling_logsize,sqrt(var_seedling_logsize)) 
      sdlng_pdf <- sdlng_cdf[2:length(bin_edges)]-sdlng_cdf[1:(length(bin_edges)-1)]
      if(sum(sdlng_pdf)>0){
        sdlng_pdf <- sdlng_pdf/sum(sdlng_pdf)} else {
          sdlng_pdf <- NA} 
      
      mx[2:(no_bins+1),1]  = predicted_rec *sdlng_pdf 
      
      mx_prod <- mx_prod %*% mx # in the first iteration of the loop, mx just becomes mx_prod
      
    } # closes ii (TSFs) loop
    
    j_pop_gr_rate_GSP[i] <- (popbio::lambda(mx_prod))^(1/length(TSFs)) 
    # multiply them together to get a matrix, etc. 
  }
    j_pop_gr_rate_GSP
}


all_pop_growth_rates_GSP <- cbind(possible_scenarios, all_pop_growth_rates_GSP)
names(all_pop_growth_rates_GSP) <- c(names(possible_scenarios), paste("rep", 1:no_reps, sep = ""))
rep_cols <- grep("^rep", names(all_pop_growth_rates_GSP), value = TRUE)

mean_lambda_across_scenarios_GSP <- all_pop_growth_rates_GSP %>%
  rowwise() %>%
  mutate(
    mean_lambda = mean(c_across(all_of(rep_cols)), na.rm = TRUE),
    lower_bound = quantile(c_across(all_of(rep_cols)), 0.025, na.rm = TRUE),
    upper_bound = quantile(c_across(all_of(rep_cols)), 0.975, na.rm = TRUE)
  ) %>%
  dplyr::select(FRI,s.TBF,mean_lambda,lower_bound, upper_bound) %>%
  mutate(TBF=NA) %>%
  ungroup()

mean_lambda_across_scenarios_GSP$TBF[1:11] <- "Short TBF"
mean_lambda_across_scenarios_GSP$TBF[12:22] <- "Long TBF"

##########. graph 
##########

plt3 <- ggplot(data= mean_lambda_across_scenarios_GSP, aes(x= FRI, y= mean_lambda)) +
  geom_line(aes(color= TBF)) + 
  geom_ribbon(aes(ymin = mean_lambda_across_scenarios_Cro$lower_bound, ymax = mean_lambda_across_scenarios_Cro$upper_bound, fill= TBF), alpha = 0.1) + 
  #ylim(0,1) +
  #xlim(c(0, 12)) +
  labs(y= "Growth Rate", x = "Fire Return Interval",color = NULL, fill = NULL) +
  theme_bw() + theme(legend.position='none') +
  theme(legend.position = "right") +
  scale_x_continuous(breaks = 0:10) +
  theme(text = element_text(size = 16)) 
plt3
```

#### 1.3.3 graph region specific predictions together with the aggregated prediction

```{r}
lambda_globalIPM <- bind_rows(
  mean_lambda_across_scenarios %>% mutate(dataset = "global"),
  mean_lambda_across_scenarios_Cro %>% mutate(dataset = "Croatan"),
  mean_lambda_across_scenarios_GSP %>% mutate(dataset = "Green Swamp")
)

plt4 <- ggplot(lambda_globalIPM, aes(x = FRI, y = mean_lambda)) +
  
  # ribbons
  geom_ribbon(
    aes(ymin = lower_bound, ymax = upper_bound,
        fill = dataset, group = interaction(dataset, TBF)),
    alpha = 0.1) +
  geom_line(
    aes(color = dataset, linetype = TBF,
        group = interaction(dataset, TBF)),
    size = 1) +
  theme_bw() +
  labs(
    y = "Growth Rate",
    x = "Fire Return Interval",
    color = "Dataset",
    fill = "Dataset",
    linetype = "TBF") +
  scale_x_continuous(breaks = 0:10) +
  theme(text = element_text(size = 16))

plt4


```

Again, the differences in growth rate among regions are driven solely by the site fixed effect in two of the vital rate functions (survival and prob of reproduction).

# 2. build region specific IPM (refit vital rate functions using subsets of data)

Now we want to build region specific IPMs with refitted vital rate functions using subsets of the master demographic data. This ensure that the vital rate functions and the IPM will capture region specific responses of plants to environmental drivers.

## 2.1 Croatan

```{r}
TBF_long_Cro <- TBF_long %>%
  filter(site=="CH"|site=="CM")
```

#### 2.1.1 refit vital rate functions

Refit vital rate functions. Here we start with the same global model and then perform model selection using AICc to determine the best fit model for each vital rate.

###### survival

```{r}
sur_subset_Cro <- 
  TBF_long_Cro %>% 
  filter_at(vars(sur0_1, s.logsize0, s.TSF,s.TBF, site), all_vars(!is.na(.)))

GM_R_sur <- glm(
   sur0_1 ~ s.logsize0 + s.TSF * s.TBF +
    s.T_LA + s.T_LH + s.sq.T_LH + s.T_LD +s.P_RH +
    s.P_RH:s.T_LH + s.P_RD + 
    site,
  data = sur_subset_Cro,
  family = "binomial", 
  na.action = "na.fail"
)

n_cores <- detectCores()
n_cores
cluster <- makeCluster(n_cores - 1)
registerDoParallel(cluster)
clusterExport(cluster, c("sur_subset_Cro"))   ### replace with different subset (different global models)
clusterEvalQ(cluster, {library(lme4); library(MuMIn)})

sur_dredge_Cro <- MuMIn::dredge(
  GM_R_sur,
  cluster = cluster,
  trace   = 2
)

sur_mod_Cro <- get.models(sur_dredge_Cro, 1)[[1]]
summary(sur_mod_Cro)

stopCluster(cluster)
```

###### growth

```{r}
gr_subset_Cro <- 
  TBF_long_Cro[which(TBF_long_Cro$sur0_1 == 1),] %>%      ## make sure to subset data to only alive plants
  filter_at(vars(logsize1, s.logsize0, s.TSF,s.sq.TSF, s.TBF, site, s.sq.T_RA,s.T_RA, s.P_RA, s.T_RC, 
                 s.T_RD, s.P_RH), all_vars(!is.na(.)))
GM_R_gr <- lmer(
  logsize1 ~ s.logsize0 + s.TSF * s.TBF  + s.T_LA + s.sq.T_LA + s.P_RA + s.T_LA:s.P_RA +
    s.T_LC + s.T_LD + s.P_RH + (1|site),
  data = gr_subset_Cro,
  na.action = "na.fail"
)

n_cores <- detectCores()
n_cores
cluster <- makeCluster(n_cores - 1)
registerDoParallel(cluster)
clusterExport(cluster, c("gr_subset_Cro"))   ### replace with different subset (different global models)
clusterEvalQ(cluster, {library(lme4); library(MuMIn)})

gr_dredge_Cro <- MuMIn::dredge(
  GM_R_gr,
  cluster = cluster,
  trace   = 2
)
growth_mod_Cro<- get.models(gr_dredge_Cro, 1)[[1]]
summary(growth_mod_Cro)
stopCluster(cluster)


```

###### prep

```{r}
prep_subset_Cro <- 
  TBF_long_Cro %>% 
  filter_at(vars(prep1, s.logsize0, s.TSF, s.TBF,s.sq.TSF, site, s.T_RA,s.P_RA, s.T_RH, s.T_RC, 
                 s.P_RH, ), all_vars(!is.na(.)))

GM_R_prep <- glm(
  prep1 ~ s.logsize0 + s.TSF * s.TBF +
    s.T_LA + s.sq.T_LA +s.P_RA+ s.sq.P_RA + s.T_LA:s.P_RA+
    s.T_LH + s.P_RH + s.T_LH:s.P_RH + s.T_LC +
    site,
  data = prep_subset_Cro,
  family = "binomial", 
  na.action = "na.fail"
)

n_cores <- detectCores()
n_cores
cluster <- makeCluster(n_cores - 1)
registerDoParallel(cluster)
clusterExport(cluster, c("prep_subset_Cro"))   ### replace with different subset (different global models)
clusterEvalQ(cluster, {library(lme4); library(MuMIn)})

prep_dredge_Cro <- MuMIn::dredge(
  GM_R_prep,
  cluster = cluster,
  trace   = 2
)
prep_mod_Cro <- get.models(prep_dredge_Cro, 1)[[1]]

stopCluster(cluster)
```

###### crep

```{r}
crep_subset_Cro <- 
  TBF_long_Cro %>% 
  filter_at(vars(logcrep1, s.logsize0, TSF, TBF,s.sq.TSF, site, s.T_RA, s.T_RH), all_vars(!is.na(.)))

GM_R_crep <- lmer(
  logcrep1 ~ s.logsize0 + s.TSF * s.TBF  + s.T_LA + s.T_LH +
    (1 | site),
  data = crep_subset_Cro,
  na.action = "na.fail"
)

n_cores <- detectCores()
n_cores
cluster <- makeCluster(n_cores - 1)
registerDoParallel(cluster)
clusterExport(cluster, c("crep_subset_Cro"))   ### replace with different subset (different global models)
clusterEvalQ(cluster, {library(lme4); library(MuMIn)})

crep_dredge_Cro <- MuMIn::dredge(
  GM_R_crep,
  cluster = cluster,
  trace   = 2
)
crep_mod_Cro <- get.models(crep_dredge_Cro, 1)[[1]]

stopCluster(cluster)
```

###### recruitment

since recruitment sample size is very low and subsetting would result in inadequate sample sizes, we use the recruitment model from the global IPM

```{r}
recruit_mod_Cro <- recruit_mod
```

###### variance in growth

```{r}
growth_terms <- attr(terms(growth_mod_Cro), "term.labels")
int_terms <- grep(":", growth_terms, value = TRUE)
int_terms_broken <- unique(unlist(strsplit(int_terms, ":", fixed = TRUE)))
main_terms <- growth_terms[!grepl(":", growth_terms)]
logsize1 <- c("logsize1")
gr_all_terms <- unique(c(main_terms, int_terms_broken, logsize1))

Vargrowth_subset_Cro <- 
  TBF_long_Cro[which(TBF_long_Cro$sur0_1 == 1),] %>%      ## make sure to subset data to only alive plants
  filter_at(vars( s.logsize0, TSF, TBF,logsize1, site), all_vars(!is.na(.)))

Vargrowth_subset_Cro$vargrowth <- NA
Vargrowth_subset_Cro$predgrowth <- NA

Vargrowth_subset_Cro$predgrowth <- predict(growth_mod_Cro,newdata=Vargrowth_subset_Cro)

Vargrowth_subset_Cro$vargrowth <-
  (Vargrowth_subset_Cro$predgrowth - # variance is equivalent to (predicted-expected)^2
     Vargrowth_subset_Cro$logsize1)^2

Vargrowth_subset_Cro$log.vargrowth <- log(Vargrowth_subset_Cro$vargrowth)

#***** to prevent negative numbers later

GM_vargrowth_R <- lmer(log.vargrowth ~ s.logsize0 + s.T_LA+ s.TSF * s.TBF  +
                         s.P_RA  + s.sq.P_RA  + s.T_LC +s.T_LD +
                         s.P_RD + s.P_RW
                       + (1|site), #  
                       data = Vargrowth_subset_Cro, na.action = "na.fail") 


n_cores <- detectCores()
cluster <- makeCluster(n_cores - 1)
registerDoParallel(cluster)
clusterExport(cluster, c("Vargrowth_subset_Cro"))   ### replace with different subset (different global models)
clusterEvalQ(cluster, {library(lme4); library(MuMIn)})


vargrowth_dredge_Cro <- MuMIn::dredge(
  GM_vargrowth_R,
  cluster = cluster,
  trace   = 2
)

vargrowth_mod_Cro <- get.models(vargrowth_dredge_Cro, 1)[[1]]

stopCluster(cluster)
```

#### 2.1.2 Croatan IPM

keep the same IPM structure as the global IPM (for example, size bound is the same as the global IPM)

```{r}


TBF_long$logsize0 <- (log(TBF_long$size0+0.1))
TBF_long$s.logsize0 <- scale(TBF_long$logsize0) 
TBF_long$s.TBF <- scale(TBF_long$TBF) 
TBF_long$s.TSF <- scale(TBF_long$TSF) 
TBF_long$sq.TSF <- (TBF_long$TSF)^2
TBF_long$s.sq.TSF <- scale(TBF_long$sq.TSF) 
IPM_bounds <- range(TBF_long$logsize0, na.rm = TRUE)  # our IPM will be based on logsize as a size metric. 
no_bins <- 100
bin_edges <- seq(IPM_bounds[1], IPM_bounds[2], length.out= no_bins+1)
bin_mids <- zoo::rollmean(bin_edges,2)

# define mean and sd of seedling sizes
rows_with_new <- TBF_long[grepl("\\bnew", TBF_long$comm1), ]
rows_with_new <- rows_with_new[!grepl("no new", rows_with_new$comm1), ] # delete "no new"
rows_with_new <- rows_with_new[!grepl("no news", rows_with_new$comm1), ] # delete "no new" 
rows_with_new <- rows_with_new[!grepl("not possible to see new plants", rows_with_new$comm1), ]
rows_with_new <- rows_with_new[!grepl("No news", rows_with_new$comm1), ]
rows_with_new <- rows_with_new[!grepl("new lvs", rows_with_new$comm1), ]
rows_with_new <- rows_with_new[!grepl("new nail", rows_with_new$comm1), ]

mean_seedling_logsize <- mean(rows_with_new$logsize1, na.rm=T) # get the mean logsize of seedings   ### the mean seedling size is very large
var_seedling_logsize <- var(rows_with_new$logsize1,na.rm=T)# get the var in logsize of seedings  

### build empty matrices to store IPM output. Here, I'm interested in the effect of fire on lambda, so I'm setting average conditions for all other climate variables
possible_scenarios <- 
  as.data.frame(matrix(NA, nrow= 22, ncol = 28))
names(possible_scenarios) <- c("FRI","s.TBF","s.T_LA", "s.T_LH",
  "s.sq.T_LH",
  "s.sq.P_LA",
  "s.T_LC",
  "s.T_LD",
  "s.P_LD",
  "s.sq.T_LA",
  "s.P_LA",
  "s.P_LW",
  "s.P_LH",
  "s.T_RA",
  "s.sq.T_RA",
  "s.P_RA",
  "s.T_RH",
  "s.P_RH",
  "s.T_RC",
  "s.T_RW",
  "s.T_RD",
  "s.sq.T_RC",
  "s.P_RW",
  "s.P_RD",
  "s.sq.T_RH",
  "s.sq.P_RA",
  "s.sq.P_RW",
  "s.sq.P_RD")


possible_scenarios$FRI <- rep(1:11, 2) # we want FRI to range from 1-11
possible_scenarios$s.TBF <- scale(rep(c(0, 10), each= 11) ,# raw TBF values
                                  center= attr(TBF_long$s.TBF,"scaled:center"), 
                                  scale= attr(TBF_long$s.TBF,"scaled:scale")) # for each set of FRI's, we want to set a short TBF (of 3) and a long TBF (of 11)

possible_scenarios[, c(  "s.T_LA",  "s.T_LH",  "s.sq.T_LH",  "s.sq.P_LA",  "s.T_LC",  "s.T_LD",  "s.P_LD",  "s.sq.T_LA",  "s.P_LA",  
                         "s.P_LW",  "s.P_LH",  "s.T_RA",  "s.sq.T_RA",  "s.P_RA",  "s.T_RH",  "s.P_RH",  "s.T_RC",  "s.T_RW",  "s.T_RD",
                         "s.sq.T_RC","s.P_RW","s.P_RD", "s.sq.T_RH",  "s.sq.P_RA",  "s.sq.P_RW",  "s.sq.P_RD")] <- 0 # setting all climate variable to their mean

my_sites <- unique(TBF_long_Cro$site) 
no_reps <- 1000 # or somethign similar

########### extract model components for later predict() ###############
sur_model_terms <- terms(sur_mod_Cro)
sur_term <- attr(sur_model_terms, "term.labels")
sur_term <- reformulate(sur_term)

gr_model_terms <- terms(growth_mod_Cro)
gr_term <- attr(gr_model_terms, "term.labels")
gr_term <- reformulate(gr_term)

prep_model_terms <- terms(prep_mod_Cro)
prep_term <- attr(prep_model_terms, "term.labels")
prep_term <- reformulate(prep_term)

crep_model_terms <- terms(crep_mod_Cro)
crep_term <- attr(crep_model_terms, "term.labels")
crep_term <- reformulate(crep_term)

vargrowth_model_terms <- terms(vargrowth_mod_Cro)
vargrowth_term <- attr(vargrowth_model_terms, "term.labels")
vargrowth_term <- reformulate(vargrowth_term)

rec_model_terms <- terms(recruit_mod_Cro)
rec_term <- attr(rec_model_terms, "term.labels")
rec_term <- reformulate(rec_term)
```

```{r}
all_pop_growth_rates_Cro_reg <- matrix(data= NA, nrow= dim(possible_scenarios)[1], ncol= no_reps)
n_cores <- detectCores()
cl <- makeCluster(n_cores - 1)
registerDoParallel(cl)
j_pop_gr_rate_Cro_reg <- rep(NA, dim(possible_scenarios)[1]) ### changed to no_scenarios

all_pop_growth_rates_Cro_reg <- foreach(
  j = 1:no_reps,
  .combine = cbind,
  .packages = c("MASS", "lme4", "popbio", "dplyr")
) %dopar% {
  rsur_coefs <- coefficients(sur_mod_Cro) # first, assign all "randomy selected" coefficients to the mean estimated coefficeints
  rsur_coefs[!names(coefficients(sur_mod_Cro)) %in% paste("site", my_sites, sep="")] <- MASS::mvrnorm(mu= coefficients(sur_mod_Cro), Sigma = vcov(sur_mod_Cro))[!names(coefficients(sur_mod_Cro)) %in% paste("site", my_sites, sep="")]
  #rsur_coefs <- lme4::fixef(survival_mod)
  #rsur_coefs <-  MASS::mvrnorm(mu= lme4::fixef(survival_mod) , Sigma = vcov(survival_mod))
  rgr_coefs <- lme4::fixef(growth_mod_Cro)
  rgr_coefs <-  MASS::mvrnorm(mu= lme4::fixef(growth_mod_Cro) , Sigma = vcov(growth_mod_Cro)) # no site effect in this model
  rprep_coefs <- coefficients(prep_mod_Cro)
  rprep_coefs[!names(coefficients(prep_mod_Cro)) %in% paste("site", my_sites, sep="")] <- MASS::mvrnorm(mu= coefficients(prep_mod_Cro), Sigma = vcov(prep_mod_Cro))[!names(coefficients(prep_mod_Cro)) %in% paste("site", my_sites, sep="")]
  rcrep_coefs <- lme4::fixef(crep_mod_Cro)
  rcrep_coefs <-  MASS::mvrnorm(mu= lme4::fixef(crep_mod_Cro) , Sigma = vcov(crep_mod_Cro)) # no site effect in this model
  rvar_coefs <- lme4::fixef(vargrowth_mod_Cro)
  rvar_coefs <-  MASS::mvrnorm(mu= lme4::fixef(vargrowth_mod_Cro) , Sigma = vcov(vargrowth_mod_Cro)) # no site effect in this model
  rrec_coefs <- lme4::fixef(recruit_mod_Cro)
  rrec_coefs <-  MASS::mvrnorm(mu= lme4::fixef(recruit_mod_Cro) , Sigma = vcov(recruit_mod_Cro)) # no site effect in this model
  #rrec_coefs <-  MASS::mvrnorm(mu= lme4::fixef(recruit_mod) , Sigma = vcov(recruit_mod)) 
  
  for (i in 1: dim(possible_scenarios)[1]) {
    
    i_scenario <- possible_scenarios[i,] # which combination of FRI, TBF, and climate do I want to use
    
    TSFs <- 0:(i_scenario$FRI-1) # TSF range from 0 to (FRI-1) such that an FRI of 1 will always have TSF= 0
    i_scenario <- i_scenario[rep(seq_len(nrow(i_scenario)), length(TSFs)), ]
    i_scenario$TSF <- TSFs
    i_scenario$s.TSF <- scale(TSFs, center= attr(TBF_long$s.TSF,"scaled:center"), 
                              scale= attr(TBF_long$s.TSF,"scaled:scale"))
    i_scenario$s.sq.TSF <- scale(TSFs^2, center= attr(TBF_long$s.sq.TSF,"scaled:center"), 
                                 scale= attr(TBF_long$s.sq.TSF,"scaled:scale"))
    
    mx_prod <- diag(no_bins+1) # setting up an identity matrix before you enter the loop
    
    for (ii in TSFs){
      i_ii_scenario <- i_scenario[which(i_scenario$TSF==ii),]
      data_for_prediction <- i_ii_scenario[rep(seq_len(nrow(i_ii_scenario)), length(bin_mids)), ]
      data_for_prediction$s.logsize0 <-   scale(bin_mids, center= attr(TBF_long$s.logsize0,"scaled:center"), 
                                                scale= attr(TBF_long$s.logsize0,"scaled:scale"))
      data_for_prediction_rec <- cbind(i_ii_scenario)
      
      predicted_sur <- predicted_gr <-predicted_pr <- predicted_cr <- predicted_vg <- matrix(data=NA, nrow= length(my_sites), ncol= length(bin_mids))
      predicted_rec <- rep(NA, length(my_sites))
      
      ### make empty mxes before entering the site loop
      
      predicted_sur_allsites <-predicted_gr_allsites<-predicted_pr_allsites <-predicted_cr_allsites<-predicted_vg_allsites <-matrix(NA, nrow = length(my_sites), ncol = no_bins, dimnames = list(my_sites, NULL))
      predicted_rec_allsites <- matrix(NA, nrow = length(my_sites), ncol = 1, dimnames = list(my_sites, NULL))
      
      
      for (s in my_sites){ #looping through the sites
        s_data_for_prediction <- data_for_prediction
        s_data_for_prediction_rec <- data_for_prediction_rec
        s_data_for_prediction$site <- s
        s_data_for_prediction_rec$site <- s
        
        s_data_for_prediction$site     <- factor(s_data_for_prediction$site,  levels = my_sites)   ## it is important to add the levels=my_sites here to ensure that the model matrix is built correctly
        s_data_for_prediction_rec$site <- factor(s_data_for_prediction_rec$site, levels = my_sites)
        
        s_data_for_prediction <- s_data_for_prediction %>% mutate(across(c("s.logsize0", "s.sq.TSF", "s.TSF", "s.TBF"), as.numeric))   
        
        m_sur <- model.matrix(sur_term ,data=s_data_for_prediction)
        p2_sur <- rsur_coefs %*% t(m_sur) # check on whether this coef(survival_mod is a column or a row!!!!!!!!
        p2_sur <- plogis(p2_sur)   ## transform to binomial 
        predicted_sur_allsites[s,] <- p2_sur
        
        m_gr <- model.matrix(gr_term, data=s_data_for_prediction)
        p2_gr <- rgr_coefs %*% t(m_gr) # check on whether this coef(survival_mod is a column or a row!!!!!!!!
        predicted_gr_allsites[s,] <- p2_gr
        
        m_prep <- model.matrix(prep_term,data=s_data_for_prediction)
        p2_prep <- rprep_coefs %*% t(m_prep) # check on whether this coef(survival_mod is a column or a row!!!!!!!!
        p2_prep <- plogis(p2_prep)   ## transform to binomial 
        predicted_pr_allsites[s,] <- p2_prep
        
        m_crep <- model.matrix(crep_term  ,data=s_data_for_prediction)
        p2_crep <- rcrep_coefs %*% t(m_crep) # check on whether this coef(survival_mod is a column or a row!!!!!!!!
        predicted_cr_allsites[s,] <- p2_crep
        
        m_vg <- model.matrix(vargrowth_term,data=s_data_for_prediction)
        p2_vg <- rvar_coefs %*% t(m_vg) # 
        predicted_vg_allsites[s,] <- exp(p2_vg)  
        
        m_rec <- model.matrix(rec_term , data=s_data_for_prediction_rec)
        p2_rec <- rrec_coefs %*% t(m_rec) # check on whether this coef(survival_mod is a column or a row!!!!!!!!
        predicted_rec_allsites[s,1] <- p2_rec*log(1.1)            #### account for composite rec variable (rec/log.fr)
        
        
        # for mixed models, the above functions will predict site-specific results (i.e., will incorporate random effects)
        rm(s_data_for_prediction) # this part removes the site-specific dataframe used in this loop, so you do not accidentaly use the previous sites' dataframe instead of the current one-- in case there was an error somehow
      }
      predicted_sur <- apply(predicted_sur_allsites, MARGIN= 2, FUN= mean, na.rm=T) # average size-specific vital rates across sites
      predicted_gr <- apply(predicted_gr_allsites, MARGIN= 2, FUN= mean, na.rm=T)
      predicted_pr <- apply(predicted_pr_allsites, MARGIN= 2, FUN= mean, na.rm=T)
      predicted_cr <- apply(predicted_cr_allsites, MARGIN= 2, FUN= mean, na.rm=T)
      predicted_vg <- apply(predicted_vg_allsites, MARGIN= 2, FUN= mean, na.rm=T)
      predicted_vg[which(predicted_vg<0)] <- 0                 #### to prevent negative variance
      predicted_rec <- mean(predicted_rec_allsites, na.rm=T)
      
      gmx <- matrix(NA, length(bin_mids),length(bin_mids)) ## create empty gmx before entering the no_bins loop
      
      # make the TSF-specific matrix
      for (ss in 1:no_bins) {
        growcdf <- pnorm(bin_edges,predicted_gr[ss],sqrt(predicted_vg[ss])) # Yes. Vargrowth=(pred - obs)^2
        grows <- growcdf[2:length(bin_edges)]-growcdf[1:(length(bin_edges)-1)]
        if(sum(grows)>0){grows <- grows/sum(grows)
        gmx[,ss] <- grows
        } else if (sum(grows)==0 & grow_vals[ss]< min(bin_edges)) { 
          gmx[,ss] <- c(1, rep(0, length(binmids)-1))  
        } else if (sum(grows)==0 & grow_vals[ss]> max(bin_edges)) { 
          gmx[,ss] <- c(rep(0, length(binmids)-1), 1)  } else { # deals with complete eviction-- OUT OF THE BOUNDS  of the kernel
            gmx[,ss] <- NA} 
        # this if statement breaks the code (puts NA's into the matrix) if the sum of the PDF is zero (which happens if all the probability is outside of the size bounds)
      } # end ss loop
      
      # make the surv*growth mx
      survgmx <- gmx*t(matrix( rep(predicted_sur,no_bins),no_bins)) # Yes, this is doing element-wise multiplication 
      reprow <-  predicted_sur*predicted_pr*predicted_cr # prob that you survive, reproduce, then number fruits|reproduce; this multiplcation does it element by element
      
      mx <- matrix(0, no_bins+1, no_bins+1)
      mx[2:(no_bins+1), 2:(no_bins+1)] = survgmx
      mx[1,2:(no_bins+1)] = reprow # FRUIT ROW
      
      sdlng_cdf <- pnorm(bin_edges,mean_seedling_logsize,sqrt(var_seedling_logsize)) 
      sdlng_pdf <- sdlng_cdf[2:length(bin_edges)]-sdlng_cdf[1:(length(bin_edges)-1)]
      if(sum(sdlng_pdf)>0){
        sdlng_pdf <- sdlng_pdf/sum(sdlng_pdf)} else {
          sdlng_pdf <- NA} 
      
      mx[2:(no_bins+1),1]  = predicted_rec *sdlng_pdf 
      
      mx_prod <- mx_prod %*% mx # in the first iteration of the loop, mx just becomes mx_prod
      
    } # closes ii (TSFs) loop
    
    j_pop_gr_rate_Cro_reg[i] <- (popbio::lambda(mx_prod))^(1/length(TSFs)) 
    # multiply them together to get a matrix, etc. 
  }
    j_pop_gr_rate_Cro_reg
}

```

#### 2.1.3 graph

```{r}
all_pop_growth_rates_Cro_reg <- cbind(possible_scenarios, all_pop_growth_rates_Cro_reg)
names(all_pop_growth_rates_Cro_reg) <- c(names(possible_scenarios), paste("rep", 1:no_reps, sep = ""))
rep_cols <- grep("^rep", names(all_pop_growth_rates_Cro_reg), value = TRUE)

mean_lambda_across_scenarios_Cro_reg <- all_pop_growth_rates_Cro_reg %>%
  rowwise() %>%
  mutate(
    mean_lambda = mean(c_across(all_of(rep_cols)), na.rm = TRUE),
    lower_bound = quantile(c_across(all_of(rep_cols)), 0.025, na.rm = TRUE),
    upper_bound = quantile(c_across(all_of(rep_cols)), 0.975, na.rm = TRUE)
  ) %>%
  dplyr::select(FRI,s.TBF,mean_lambda,lower_bound, upper_bound) %>%
  mutate(TBF=NA) %>%
  ungroup()

mean_lambda_across_scenarios_Cro_reg$TBF[1:11] <- "Short TBF"
mean_lambda_across_scenarios_Cro_reg$TBF[12:22] <- "Long TBF"

##########. graph 
##########

plt5<- ggplot(data= mean_lambda_across_scenarios_Cro_reg, aes(x= FRI, y= mean_lambda)) +
  geom_line(aes(color= TBF)) + 
  geom_ribbon(aes(ymin = mean_lambda_across_scenarios_Cro_reg$lower_bound, ymax = mean_lambda_across_scenarios_Cro_reg$upper_bound, fill= TBF), alpha = 0.1) + 
  #ylim(0,1) +
  #xlim(c(0, 12)) +
  labs(y= "Growth Rate", x = "Fire Return Interval",color = NULL, fill = NULL) +
  theme_bw() + theme(legend.position='none') +
  theme(legend.position = "right") +
  scale_x_continuous(breaks = 0:10) +
  theme(text = element_text(size = 16)) 
plt5
```

## 2.2 Green Swamp

Same procedure but for the GSP region.

```{r}
TBF_long_GSP <- TBF_long %>%
  filter(site=="GSP-LI"|site=="GSP-BI")
```

#### 2.2.1 refit vital rate

###### survival

```{r}
sur_subset_GSP <- 
  TBF_long_GSP %>% 
  filter_at(vars(sur0_1, s.logsize0, s.TSF,s.TBF, site), all_vars(!is.na(.)))

GM_R_sur <- glm(
   sur0_1 ~ s.logsize0 + s.TSF * s.TBF +
    s.T_LA + s.T_LH + s.sq.T_LH + s.T_LD +s.P_RH +
    s.P_RH:s.T_LH + s.P_RD + 
    site,
  data = sur_subset_GSP,
  family = "binomial", 
  na.action = "na.fail"
)

n_cores <- detectCores()
n_cores
cluster <- makeCluster(n_cores - 1)
registerDoParallel(cluster)
clusterExport(cluster, c("sur_subset_GSP"))   ### replace with different subset (different global models)
clusterEvalQ(cluster, {library(lme4); library(MuMIn)})

sur_dredge_GSP <- MuMIn::dredge(
  GM_R_sur,
  cluster = cluster,
  trace   = 2
)

sur_mod_GSP <- get.models(sur_dredge_GSP, 1)[[1]]
summary(sur_mod_GSP)

stopCluster(cluster)
```

###### growth

```{r}
gr_subset_GSP <- 
  TBF_long_GSP[which(TBF_long_GSP$sur0_1 == 1),] %>%      ## make sure to subset data to only alive plants
  filter_at(vars(logsize1, s.logsize0, s.TSF,s.sq.TSF, s.TBF, site, s.sq.T_RA,s.T_RA, s.P_RA, s.T_RC, 
                 s.T_RD, s.P_RH), all_vars(!is.na(.)))
GM_R_gr <- lmer(
  logsize1 ~ s.logsize0 + s.TSF * s.TBF  + s.T_LA + s.sq.T_LA + s.P_RA + s.T_LA:s.P_RA +
    s.T_LC + s.T_LD + s.P_RH + (1|site),
  data = gr_subset_GSP,
  na.action = "na.fail"
)

n_cores <- detectCores()
n_cores
cluster <- makeCluster(n_cores - 1)
registerDoParallel(cluster)
clusterExport(cluster, c("gr_subset_GSP"))   ### replace with different subset (different global models)
clusterEvalQ(cluster, {library(lme4); library(MuMIn)})

gr_dredge_GSP <- MuMIn::dredge(
  GM_R_gr,
  cluster = cluster,
  trace   = 2
)
growth_mod_GSP<- get.models(gr_dredge_GSP, 1)[[1]]
summary(growth_mod_GSP)
stopCluster(cluster)
```

###### prep

```{r}
prep_subset_GSP <- 
  TBF_long_GSP %>% 
  filter_at(vars(prep1, s.logsize0, s.TSF, s.TBF,s.sq.TSF, site, s.T_RA,s.P_RA, s.T_RH, s.T_RC, 
                 s.P_RH, ), all_vars(!is.na(.)))

GM_R_prep <- glm(
  prep1 ~ s.logsize0 + s.TSF * s.TBF +
    s.T_LA + s.sq.T_LA +s.P_RA+ s.sq.P_RA + s.T_LA:s.P_RA+
    s.T_LH + s.P_RH + s.T_LH:s.P_RH + s.T_LC +
    site,
  data = prep_subset_GSP,
  family = "binomial", 
  na.action = "na.fail"
)

n_cores <- detectCores()
n_cores
cluster <- makeCluster(n_cores - 1)
registerDoParallel(cluster)
clusterExport(cluster, c("prep_subset_GSP"))   ### replace with different subset (different global models)
clusterEvalQ(cluster, {library(lme4); library(MuMIn)})

prep_dredge_GSP <- MuMIn::dredge(
  GM_R_prep,
  cluster = cluster,
  trace   = 2
)
prep_mod_GSP <- get.models(prep_dredge_GSP, 1)[[1]]
summary(prep_mod_GSP)
stopCluster(cluster)
```

###### crep

```{r}

crep_subset_GSP <- 
  TBF_long_GSP %>% 
  filter_at(vars(logcrep1, s.logsize0, TSF, TBF,s.sq.TSF, site, s.T_RA, s.T_RH), all_vars(!is.na(.)))

GM_R_crep <- lmer(
  logcrep1 ~ s.logsize0 + s.TSF * s.TBF  + s.T_LA + s.T_LH +
    (1 | site),
  data = crep_subset_GSP,
  na.action = "na.fail"
)

n_cores <- detectCores()
n_cores
cluster <- makeCluster(n_cores - 1)
registerDoParallel(cluster)
clusterExport(cluster, c("crep_subset_GSP"))   ### replace with different subset (different global models)
clusterEvalQ(cluster, {library(lme4); library(MuMIn)})

crep_dredge_GSP <- MuMIn::dredge(
  GM_R_crep,
  cluster = cluster,
  trace   = 2
)
crep_mod_GSP <- get.models(crep_dredge_GSP, 1)[[1]]

stopCluster(cluster)
```

###### recruitment

```{r}
recruit_mod_GSP <- recruit_mod
```

###### variance in growth

```{r}
growth_terms <- attr(terms(growth_mod_GSP), "term.labels")
int_terms <- grep(":", growth_terms, value = TRUE)
int_terms_broken <- unique(unlist(strsplit(int_terms, ":", fixed = TRUE)))
main_terms <- growth_terms[!grepl(":", growth_terms)]
logsize1 <- c("logsize1")
gr_all_terms <- unique(c(main_terms, int_terms_broken, logsize1))

Vargrowth_subset_GSP <- 
  TBF_long_GSP[which(TBF_long_GSP$sur0_1 == 1),] %>%      ## make sure to subset data to only alive plants
  filter_at(vars( s.logsize0, TSF, TBF,logsize1, site), all_vars(!is.na(.)))

Vargrowth_subset_GSP$vargrowth <- NA
Vargrowth_subset_GSP$predgrowth <- NA

Vargrowth_subset_GSP$predgrowth <- predict(growth_mod_GSP,newdata=Vargrowth_subset_GSP)

Vargrowth_subset_GSP$vargrowth <-
  (Vargrowth_subset_GSP$predgrowth - # variance is equivalent to (predicted-expected)^2
     Vargrowth_subset_GSP$logsize1)^2

Vargrowth_subset_GSP$log.vargrowth <- log(Vargrowth_subset_GSP$vargrowth)

#***** to prevent negative numbers later

GM_vargrowth_R <- lmer(log.vargrowth ~ s.logsize0 + s.T_LA+ s.TSF * s.TBF  +
                         s.P_RA  + s.sq.P_RA  + s.T_LC +s.T_LD +
                         s.P_RD + s.P_RW
                       + (1|site), #  
                       data = Vargrowth_subset_GSP, na.action = "na.fail") 


n_cores <- detectCores()
cluster <- makeCluster(n_cores - 1)
registerDoParallel(cluster)
clusterExport(cluster, c("Vargrowth_subset_GSP"))   ### replace with different subset (different global models)
clusterEvalQ(cluster, {library(lme4); library(MuMIn)})


vargrowth_dredge_GSP <- MuMIn::dredge(
  GM_vargrowth_R,
  cluster = cluster,
  trace   = 2
)

vargrowth_mod_GSP <- get.models(vargrowth_dredge_GSP, 1)[[1]]

stopCluster(cluster)
```

#### 2.2.2 GSP IPM

```{r}
TBF_long$logsize0 <- (log(TBF_long$size0+0.1))
TBF_long$s.logsize0 <- scale(TBF_long$logsize0) 
TBF_long$s.TBF <- scale(TBF_long$TBF) 
TBF_long$s.TSF <- scale(TBF_long$TSF) 
TBF_long$sq.TSF <- (TBF_long$TSF)^2
TBF_long$s.sq.TSF <- scale(TBF_long$sq.TSF) 
IPM_bounds <- range(TBF_long$logsize0, na.rm = TRUE)  # our IPM will be based on logsize as a size metric. 
no_bins <- 100
bin_edges <- seq(IPM_bounds[1], IPM_bounds[2], length.out= no_bins+1)
bin_mids <- zoo::rollmean(bin_edges,2)

# define mean and sd of seedling sizes
rows_with_new <- TBF_long[grepl("\\bnew", TBF_long$comm1), ]
rows_with_new <- rows_with_new[!grepl("no new", rows_with_new$comm1), ] # delete "no new"
rows_with_new <- rows_with_new[!grepl("no news", rows_with_new$comm1), ] # delete "no new" 
rows_with_new <- rows_with_new[!grepl("not possible to see new plants", rows_with_new$comm1), ]
rows_with_new <- rows_with_new[!grepl("No news", rows_with_new$comm1), ]
rows_with_new <- rows_with_new[!grepl("new lvs", rows_with_new$comm1), ]
rows_with_new <- rows_with_new[!grepl("new nail", rows_with_new$comm1), ]

mean_seedling_logsize <- mean(rows_with_new$logsize1, na.rm=T) # get the mean logsize of seedings   ### the mean seedling size is very large
var_seedling_logsize <- var(rows_with_new$logsize1,na.rm=T)# get the var in logsize of seedings  

### build empty matrices to store IPM output. Here, I'm interested in the effect of fire on lambda, so I'm setting average conditions for all other climate variables
possible_scenarios <- 
  as.data.frame(matrix(NA, nrow= 22, ncol = 28))
names(possible_scenarios) <- c("FRI","s.TBF","s.T_LA", "s.T_LH",
  "s.sq.T_LH",
  "s.sq.P_LA",
  "s.T_LC",
  "s.T_LD",
  "s.P_LD",
  "s.sq.T_LA",
  "s.P_LA",
  "s.P_LW",
  "s.P_LH",
  "s.T_RA",
  "s.sq.T_RA",
  "s.P_RA",
  "s.T_RH",
  "s.P_RH",
  "s.T_RC",
  "s.T_RW",
  "s.T_RD",
  "s.sq.T_RC",
  "s.P_RW",
  "s.P_RD",
  "s.sq.T_RH",
  "s.sq.P_RA",
  "s.sq.P_RW",
  "s.sq.P_RD")


possible_scenarios$FRI <- rep(1:11, 2) # we want FRI to range from 1-11
possible_scenarios$s.TBF <- scale(rep(c(0, 10), each= 11) ,# raw TBF values
                                  center= attr(TBF_long$s.TBF,"scaled:center"), 
                                  scale= attr(TBF_long$s.TBF,"scaled:scale")) # for each set of FRI's, we want to set a short TBF (of 3) and a long TBF (of 11)

possible_scenarios[, c(  "s.T_LA",  "s.T_LH",  "s.sq.T_LH",  "s.sq.P_LA",  "s.T_LC",  "s.T_LD",  "s.P_LD",  "s.sq.T_LA",  "s.P_LA",  
                         "s.P_LW",  "s.P_LH",  "s.T_RA",  "s.sq.T_RA",  "s.P_RA",  "s.T_RH",  "s.P_RH",  "s.T_RC",  "s.T_RW",  "s.T_RD",
                         "s.sq.T_RC","s.P_RW","s.P_RD", "s.sq.T_RH",  "s.sq.P_RA",  "s.sq.P_RW",  "s.sq.P_RD")] <- 0 # setting all climate variable to their mean

my_sites <- unique(TBF_long_GSP$site) 
no_reps <- 1000 # or somethign similar

########### extract model components for later predict() ###############
sur_model_terms <- terms(sur_mod_GSP)
sur_term <- attr(sur_model_terms, "term.labels")
sur_term <- reformulate(sur_term)

gr_model_terms <- terms(growth_mod_GSP)
gr_term <- attr(gr_model_terms, "term.labels")
gr_term <- reformulate(gr_term)

prep_model_terms <- terms(prep_mod_GSP)
prep_term <- attr(prep_model_terms, "term.labels")
prep_term <- reformulate(prep_term)

crep_model_terms <- terms(crep_mod_GSP)
crep_term <- attr(crep_model_terms, "term.labels")
crep_term <- reformulate(crep_term)

vargrowth_model_terms <- terms(vargrowth_mod_GSP)
vargrowth_term <- attr(vargrowth_model_terms, "term.labels")
vargrowth_term <- reformulate(vargrowth_term)

rec_model_terms <- terms(recruit_mod_GSP)
rec_term <- attr(rec_model_terms, "term.labels")
rec_term <- reformulate(rec_term)
```

```{r}
all_pop_growth_rates_GSP_reg <- matrix(data= NA, nrow= dim(possible_scenarios)[1], ncol= no_reps)
n_cores <- detectCores()
cl <- makeCluster(n_cores - 1)
registerDoParallel(cl)
j_pop_gr_rate_GSP_reg <- rep(NA, dim(possible_scenarios)[1]) ### changed to no_scenarios

all_pop_growth_rates_GSP_reg <- foreach(
  j = 1:no_reps,
  .combine = cbind,
  .packages = c("MASS", "lme4", "popbio", "dplyr")
) %dopar% {
  rsur_coefs <- coefficients(sur_mod_GSP) # first, assign all "randomy selected" coefficients to the mean estimated coefficeints
  rsur_coefs[!names(coefficients(sur_mod_GSP)) %in% paste("site", my_sites, sep="")] <- MASS::mvrnorm(mu= coefficients(sur_mod_GSP), Sigma = vcov(sur_mod_GSP))[!names(coefficients(sur_mod_GSP)) %in% paste("site", my_sites, sep="")]
  #rsur_coefs <- lme4::fixef(survival_mod)
  #rsur_coefs <-  MASS::mvrnorm(mu= lme4::fixef(survival_mod) , Sigma = vcov(survival_mod))
  rgr_coefs <- lme4::fixef(growth_mod_GSP)
  rgr_coefs <-  MASS::mvrnorm(mu= lme4::fixef(growth_mod_GSP) , Sigma = vcov(growth_mod_GSP)) # no site effect in this model
  rprep_coefs <- coefficients(prep_mod_GSP)
  rprep_coefs[!names(coefficients(prep_mod_GSP)) %in% paste("site", my_sites, sep="")] <- MASS::mvrnorm(mu= coefficients(prep_mod_GSP), Sigma = vcov(prep_mod_GSP))[!names(coefficients(prep_mod_GSP)) %in% paste("site", my_sites, sep="")]
  rcrep_coefs <- lme4::fixef(crep_mod_GSP)
  rcrep_coefs <-  MASS::mvrnorm(mu= lme4::fixef(crep_mod_GSP) , Sigma = vcov(crep_mod_GSP)) # no site effect in this model
  rvar_coefs <- lme4::fixef(vargrowth_mod_GSP)
  rvar_coefs <-  MASS::mvrnorm(mu= lme4::fixef(vargrowth_mod_GSP) , Sigma = vcov(vargrowth_mod_GSP)) # no site effect in this model
  rrec_coefs <- lme4::fixef(recruit_mod_GSP)
  rrec_coefs <-  MASS::mvrnorm(mu= lme4::fixef(recruit_mod_GSP) , Sigma = vcov(recruit_mod_GSP)) # no site effect in this model
  #rrec_coefs <-  MASS::mvrnorm(mu= lme4::fixef(recruit_mod) , Sigma = vcov(recruit_mod)) 
  
  for (i in 1: dim(possible_scenarios)[1]) {
    
    i_scenario <- possible_scenarios[i,] # which combination of FRI, TBF, and climate do I want to use
    
    TSFs <- 0:(i_scenario$FRI-1) # TSF range from 0 to (FRI-1) such that an FRI of 1 will always have TSF= 0
    i_scenario <- i_scenario[rep(seq_len(nrow(i_scenario)), length(TSFs)), ]
    i_scenario$TSF <- TSFs
    i_scenario$s.TSF <- scale(TSFs, center= attr(TBF_long$s.TSF,"scaled:center"), 
                              scale= attr(TBF_long$s.TSF,"scaled:scale"))
    i_scenario$s.sq.TSF <- scale(TSFs^2, center= attr(TBF_long$s.sq.TSF,"scaled:center"), 
                                 scale= attr(TBF_long$s.sq.TSF,"scaled:scale"))
    
    mx_prod <- diag(no_bins+1) # setting up an identity matrix before you enter the loop
    
    for (ii in TSFs){
      i_ii_scenario <- i_scenario[which(i_scenario$TSF==ii),]
      data_for_prediction <- i_ii_scenario[rep(seq_len(nrow(i_ii_scenario)), length(bin_mids)), ]
      data_for_prediction$s.logsize0 <-   scale(bin_mids, center= attr(TBF_long$s.logsize0,"scaled:center"), 
                                                scale= attr(TBF_long$s.logsize0,"scaled:scale"))
      data_for_prediction_rec <- cbind(i_ii_scenario)
      
      predicted_sur <- predicted_gr <-predicted_pr <- predicted_cr <- predicted_vg <- matrix(data=NA, nrow= length(my_sites), ncol= length(bin_mids))
      predicted_rec <- rep(NA, length(my_sites))
      
      ### make empty mxes before entering the site loop
      
      predicted_sur_allsites <-predicted_gr_allsites<-predicted_pr_allsites <-predicted_cr_allsites<-predicted_vg_allsites <-matrix(NA, nrow = length(my_sites), ncol = no_bins, dimnames = list(my_sites, NULL))
      predicted_rec_allsites <- matrix(NA, nrow = length(my_sites), ncol = 1, dimnames = list(my_sites, NULL))
      
      
      for (s in my_sites){ #looping through the sites
        s_data_for_prediction <- data_for_prediction
        s_data_for_prediction_rec <- data_for_prediction_rec
        s_data_for_prediction$site <- s
        s_data_for_prediction_rec$site <- s
        
        s_data_for_prediction$site     <- factor(s_data_for_prediction$site,  levels = my_sites)   ## it is important to add the levels=my_sites here to ensure that the model matrix is built correctly
        s_data_for_prediction_rec$site <- factor(s_data_for_prediction_rec$site, levels = my_sites)
        
        s_data_for_prediction <- s_data_for_prediction %>% mutate(across(c("s.logsize0", "s.sq.TSF", "s.TSF", "s.TBF"), as.numeric))   
        
        m_sur <- model.matrix(sur_term ,data=s_data_for_prediction)
        p2_sur <- rsur_coefs %*% t(m_sur) # check on whether this coef(survival_mod is a column or a row!!!!!!!!
        p2_sur <- plogis(p2_sur)   ## transform to binomial 
        predicted_sur_allsites[s,] <- p2_sur
        
        m_gr <- model.matrix(gr_term, data=s_data_for_prediction)
        p2_gr <- rgr_coefs %*% t(m_gr) # check on whether this coef(survival_mod is a column or a row!!!!!!!!
        predicted_gr_allsites[s,] <- p2_gr
        
        m_prep <- model.matrix(prep_term,data=s_data_for_prediction)
        p2_prep <- rprep_coefs %*% t(m_prep) # check on whether this coef(survival_mod is a column or a row!!!!!!!!
        p2_prep <- plogis(p2_prep)   ## transform to binomial 
        predicted_pr_allsites[s,] <- p2_prep
        
        m_crep <- model.matrix(crep_term  ,data=s_data_for_prediction)
        p2_crep <- rcrep_coefs %*% t(m_crep) # check on whether this coef(survival_mod is a column or a row!!!!!!!!
        predicted_cr_allsites[s,] <- p2_crep
        
        m_vg <- model.matrix(vargrowth_term,data=s_data_for_prediction)
        p2_vg <- rvar_coefs %*% t(m_vg) # 
        predicted_vg_allsites[s,] <- exp(p2_vg)  
        
        m_rec <- model.matrix(rec_term , data=s_data_for_prediction_rec)
        p2_rec <- rrec_coefs %*% t(m_rec) # check on whether this coef(survival_mod is a column or a row!!!!!!!!
        predicted_rec_allsites[s,1] <- p2_rec*log(1.1)            #### account for composite rec variable (rec/log.fr)
        
        
        # for mixed models, the above functions will predict site-specific results (i.e., will incorporate random effects)
        rm(s_data_for_prediction) # this part removes the site-specific dataframe used in this loop, so you do not accidentaly use the previous sites' dataframe instead of the current one-- in case there was an error somehow
      }
      predicted_sur <- apply(predicted_sur_allsites, MARGIN= 2, FUN= mean, na.rm=T) # average size-specific vital rates across sites
      predicted_gr <- apply(predicted_gr_allsites, MARGIN= 2, FUN= mean, na.rm=T)
      predicted_pr <- apply(predicted_pr_allsites, MARGIN= 2, FUN= mean, na.rm=T)
      predicted_cr <- apply(predicted_cr_allsites, MARGIN= 2, FUN= mean, na.rm=T)
      predicted_vg <- apply(predicted_vg_allsites, MARGIN= 2, FUN= mean, na.rm=T)
      predicted_vg[which(predicted_vg<0)] <- 0                 #### to prevent negative variance
      predicted_rec <- mean(predicted_rec_allsites, na.rm=T)
      
      gmx <- matrix(NA, length(bin_mids),length(bin_mids)) ## create empty gmx before entering the no_bins loop
      
      # make the TSF-specific matrix
      for (ss in 1:no_bins) {
        growcdf <- pnorm(bin_edges,predicted_gr[ss],sqrt(predicted_vg[ss])) # Yes. Vargrowth=(pred - obs)^2
        grows <- growcdf[2:length(bin_edges)]-growcdf[1:(length(bin_edges)-1)]
        if(sum(grows)>0){grows <- grows/sum(grows)
        gmx[,ss] <- grows
        } else if (sum(grows)==0 & predicted_gr[ss]< min(bin_edges)) { 
          gmx[,ss] <- c(1, rep(0, length(bin_mids)-1))  
        } else if (sum(grows)==0 & predicted_gr[ss]> max(bin_edges)) { 
          gmx[,ss] <- c(rep(0, length(bin_mids)-1), 1)  } else { # deals with complete eviction-- OUT OF THE BOUNDS  of the kernel
            gmx[,ss] <- NA} 
        # this if statement breaks the code (puts NA's into the matrix) if the sum of the PDF is zero (which happens if all the probability is outside of the size bounds)
      } # end ss loop
      
      # make the surv*growth mx
      survgmx <- gmx*t(matrix( rep(predicted_sur,no_bins),no_bins)) # Yes, this is doing element-wise multiplication 
      reprow <-  predicted_sur*predicted_pr*predicted_cr # prob that you survive, reproduce, then number fruits|reproduce; this multiplcation does it element by element
      
      mx <- matrix(0, no_bins+1, no_bins+1)
      mx[2:(no_bins+1), 2:(no_bins+1)] = survgmx
      mx[1,2:(no_bins+1)] = reprow # FRUIT ROW
      
      sdlng_cdf <- pnorm(bin_edges,mean_seedling_logsize,sqrt(var_seedling_logsize)) 
      sdlng_pdf <- sdlng_cdf[2:length(bin_edges)]-sdlng_cdf[1:(length(bin_edges)-1)]
      if(sum(sdlng_pdf)>0){
        sdlng_pdf <- sdlng_pdf/sum(sdlng_pdf)} else {
          sdlng_pdf <- NA} 
      
      mx[2:(no_bins+1),1]  = predicted_rec *sdlng_pdf 
      
      mx_prod <- mx_prod %*% mx # in the first iteration of the loop, mx just becomes mx_prod
      
    } # closes ii (TSFs) loop
    
    j_pop_gr_rate_GSP_reg[i] <- (popbio::lambda(mx_prod))^(1/length(TSFs)) 
    # multiply them together to get a matrix, etc. 
  }
    j_pop_gr_rate_GSP_reg
}

```

#### 2.2.3 graph

```{r}
all_pop_growth_rates_GSP_reg <- cbind(possible_scenarios, all_pop_growth_rates_GSP_reg)
names(all_pop_growth_rates_GSP_reg) <- c(names(possible_scenarios), paste("rep", 1:no_reps, sep = ""))
rep_cols <- grep("^rep", names(all_pop_growth_rates_GSP_reg), value = TRUE)

mean_lambda_across_scenarios_GSP_reg <- all_pop_growth_rates_GSP_reg %>%
  rowwise() %>%
  mutate(
    mean_lambda = mean(c_across(all_of(rep_cols)), na.rm = TRUE),
    lower_bound = quantile(c_across(all_of(rep_cols)), 0.025, na.rm = TRUE),
    upper_bound = quantile(c_across(all_of(rep_cols)), 0.975, na.rm = TRUE)
  ) %>%
  dplyr::select(FRI,s.TBF,mean_lambda,lower_bound, upper_bound) %>%
  mutate(TBF=NA) %>%
  ungroup()

mean_lambda_across_scenarios_GSP_reg$TBF[1:11] <- "Short TBF"
mean_lambda_across_scenarios_GSP_reg$TBF[12:22] <- "Long TBF"

##########. graph 
##########

plt6<- ggplot(data= mean_lambda_across_scenarios_GSP_reg, aes(x= FRI, y= mean_lambda)) +
  geom_line(aes(color= TBF)) + 
  geom_ribbon(aes(ymin = mean_lambda_across_scenarios_GSP_reg$lower_bound, ymax = mean_lambda_across_scenarios_GSP_reg$upper_bound, fill= TBF), alpha = 0.1) + 
  #ylim(0,1) +
  #xlim(c(0, 12)) +
  labs(y= "Growth Rate", x = "Fire Return Interval",color = NULL, fill = NULL) +
  theme_bw() + theme(legend.position='none') +
  theme(legend.position = "right") +
  scale_x_continuous(breaks = 0:10) +
  theme(text = element_text(size = 16)) 
plt6
```

## 2.3 Fort Bragg

The Fort Bragg region contains B1 and B2. These two populations are in only inland populations.

```{r}
TBF_long_FB <- TBF_long %>%
  filter(site=="B1"|site=="B2")
```

#### 2.3.1 refit vital rate

###### survival

```{r}
sur_subset_FB <- 
  TBF_long_FB %>% 
  filter_at(vars(sur0_1, s.logsize0, s.TSF,s.TBF, site), all_vars(!is.na(.)))

GM_R_sur <- glm(
   sur0_1 ~ s.logsize0 + s.TSF * s.TBF +
    s.T_LA + s.T_LH + s.sq.T_LH + s.T_LD +s.P_RH +
    s.P_RH:s.T_LH + s.P_RD + 
    site,
  data = sur_subset_FB,
  family = "binomial", 
  na.action = "na.fail"
)

n_cores <- detectCores()
n_cores
cluster <- makeCluster(n_cores - 1)
registerDoParallel(cluster)
clusterExport(cluster, c("sur_subset_FB"))   ### replace with different subset (different global models)
clusterEvalQ(cluster, {library(lme4); library(MuMIn)})

sur_dredge_FB <- MuMIn::dredge(
  GM_R_sur,
  cluster = cluster,
  trace   = 2
)

sur_mod_FB <- get.models(sur_dredge_FB, 1)[[1]]
summary(sur_mod_FB)

stopCluster(cluster)
```

###### growth

```{r}
gr_subset_FB <- 
  TBF_long_FB[which(TBF_long_FB$sur0_1 == 1),] %>%      ## make sure to subset data to only alive plants
  filter_at(vars(logsize1, s.logsize0, s.TSF,s.sq.TSF, s.TBF, site, s.sq.T_RA,s.T_RA, s.P_RA, s.T_RC, 
                 s.T_RD, s.P_RH), all_vars(!is.na(.)))
GM_R_gr <- lmer(
  logsize1 ~ s.logsize0 + s.TSF * s.TBF  + s.T_LA + s.sq.T_LA + s.P_RA + s.T_LA:s.P_RA +
    s.T_LC + s.T_LD + s.P_RH + (1|site),
  data = gr_subset_FB,
  na.action = "na.fail"
)

n_cores <- detectCores()
n_cores
cluster <- makeCluster(n_cores - 1)
registerDoParallel(cluster)
clusterExport(cluster, c("gr_subset_FB"))   ### replace with different subset (different global models)
clusterEvalQ(cluster, {library(lme4); library(MuMIn)})

gr_dredge_FB <- MuMIn::dredge(
  GM_R_gr,
  cluster = cluster,
  trace   = 2
)
growth_mod_FB<- get.models(gr_dredge_FB, 1)[[1]]
summary(growth_mod_FB)
stopCluster(cluster)
```

###### prep

```{r}
prep_subset_FB <- 
  TBF_long_FB %>% 
  filter_at(vars(prep1, s.logsize0, s.TSF, s.TBF,s.sq.TSF, site, s.T_RA,s.P_RA, s.T_RH, s.T_RC, 
                 s.P_RH, ), all_vars(!is.na(.)))

GM_R_prep <- glm(
  prep1 ~ s.logsize0 + s.TSF * s.TBF +
    s.T_LA + s.sq.T_LA +s.P_RA+ s.sq.P_RA + s.T_LA:s.P_RA+
    s.T_LH + s.P_RH + s.T_LH:s.P_RH + s.T_LC +
    site,
  data = prep_subset_FB,
  family = "binomial", 
  na.action = "na.fail"
)

n_cores <- detectCores()
n_cores
cluster <- makeCluster(n_cores - 1)
registerDoParallel(cluster)
clusterExport(cluster, c("prep_subset_FB"))   ### replace with different subset (different global models)
clusterEvalQ(cluster, {library(lme4); library(MuMIn)})

prep_dredge_FB <- MuMIn::dredge(
  GM_R_prep,
  cluster = cluster,
  trace   = 2
)
prep_mod_FB <- get.models(prep_dredge_FB, 1)[[1]]
summary(prep_mod_FB)
stopCluster(cluster)
```

###### crep

```{r}

crep_subset_FB <- 
  TBF_long_FB %>% 
  filter_at(vars(logcrep1, s.logsize0, TSF, TBF,s.sq.TSF, site, s.T_RA, s.T_RH), all_vars(!is.na(.)))

GM_R_crep <- lmer(
  logcrep1 ~ s.logsize0 + s.TSF * s.TBF  + s.T_LA + s.T_LH +
    (1 | site),
  data = crep_subset_FB,
  na.action = "na.fail"
)

n_cores <- detectCores()
n_cores
cluster <- makeCluster(n_cores - 1)
registerDoParallel(cluster)
clusterExport(cluster, c("crep_subset_FB"))   ### replace with different subset (different global models)
clusterEvalQ(cluster, {library(lme4); library(MuMIn)})

crep_dredge_FB <- MuMIn::dredge(
  GM_R_crep,
  cluster = cluster,
  trace   = 2
)
crep_mod_FB <- get.models(crep_dredge_FB, 1)[[1]]

stopCluster(cluster)
```

###### recruitment

```{r}
recruit_mod_FB <- recruit_mod
```

###### variance in growth

```{r}
growth_terms <- attr(terms(growth_mod_FB), "term.labels")
int_terms <- grep(":", growth_terms, value = TRUE)
int_terms_broken <- unique(unlist(strsplit(int_terms, ":", fixed = TRUE)))
main_terms <- growth_terms[!grepl(":", growth_terms)]
logsize1 <- c("logsize1")
gr_all_terms <- unique(c(main_terms, int_terms_broken, logsize1))

Vargrowth_subset_FB <- 
  TBF_long_FB[which(TBF_long_FB$sur0_1 == 1),] %>%      ## make sure to subset data to only alive plants
  filter_at(vars( s.logsize0, TSF, TBF,logsize1, site), all_vars(!is.na(.)))

Vargrowth_subset_FB$vargrowth <- NA
Vargrowth_subset_FB$predgrowth <- NA

Vargrowth_subset_FB$predgrowth <- predict(growth_mod_FB,newdata=Vargrowth_subset_FB)

Vargrowth_subset_FB$vargrowth <-
  (Vargrowth_subset_FB$predgrowth - # variance is equivalent to (predicted-expected)^2
     Vargrowth_subset_FB$logsize1)^2

Vargrowth_subset_FB$log.vargrowth <- log(Vargrowth_subset_FB$vargrowth)

#***** to prevent negative numbers later

GM_vargrowth_R <- lmer(log.vargrowth ~ s.logsize0 + s.T_LA+ s.TSF * s.TBF  +
                         s.P_RA  + s.sq.P_RA  + s.T_LC +s.T_LD +
                         s.P_RD + s.P_RW
                       + (1|site), #  
                       data = Vargrowth_subset_FB, na.action = "na.fail") 


n_cores <- detectCores()
cluster <- makeCluster(n_cores - 1)
registerDoParallel(cluster)
clusterExport(cluster, c("Vargrowth_subset_FB"))   ### replace with different subset (different global models)
clusterEvalQ(cluster, {library(lme4); library(MuMIn)})


vargrowth_dredge_FB <- MuMIn::dredge(
  GM_vargrowth_R,
  cluster = cluster,
  trace   = 2
)

vargrowth_mod_FB <- get.models(vargrowth_dredge_FB, 1)[[1]]

stopCluster(cluster)
```

#### 2.3.2 FB IPM

```{r}
TBF_long$logsize0 <- (log(TBF_long$size0+0.1))
TBF_long$s.logsize0 <- scale(TBF_long$logsize0) 
TBF_long$s.TBF <- scale(TBF_long$TBF) 
TBF_long$s.TSF <- scale(TBF_long$TSF) 
TBF_long$sq.TSF <- (TBF_long$TSF)^2
TBF_long$s.sq.TSF <- scale(TBF_long$sq.TSF) 
IPM_bounds <- range(TBF_long$logsize0, na.rm = TRUE)  # our IPM will be based on logsize as a size metric. 
no_bins <- 100
bin_edges <- seq(IPM_bounds[1], IPM_bounds[2], length.out= no_bins+1)
bin_mids <- zoo::rollmean(bin_edges,2)

# define mean and sd of seedling sizes
rows_with_new <- TBF_long[grepl("\\bnew", TBF_long$comm1), ]
rows_with_new <- rows_with_new[!grepl("no new", rows_with_new$comm1), ] # delete "no new"
rows_with_new <- rows_with_new[!grepl("no news", rows_with_new$comm1), ] # delete "no new" 
rows_with_new <- rows_with_new[!grepl("not possible to see new plants", rows_with_new$comm1), ]
rows_with_new <- rows_with_new[!grepl("No news", rows_with_new$comm1), ]
rows_with_new <- rows_with_new[!grepl("new lvs", rows_with_new$comm1), ]
rows_with_new <- rows_with_new[!grepl("new nail", rows_with_new$comm1), ]

mean_seedling_logsize <- mean(rows_with_new$logsize1, na.rm=T) # get the mean logsize of seedings   ### the mean seedling size is very large
var_seedling_logsize <- var(rows_with_new$logsize1,na.rm=T)# get the var in logsize of seedings  

### build empty matrices to store IPM output. Here, I'm interested in the effect of fire on lambda, so I'm setting average conditions for all other climate variables
possible_scenarios <- 
  as.data.frame(matrix(NA, nrow= 22, ncol = 28))
names(possible_scenarios) <- c("FRI","s.TBF","s.T_LA", "s.T_LH",
  "s.sq.T_LH",
  "s.sq.P_LA",
  "s.T_LC",
  "s.T_LD",
  "s.P_LD",
  "s.sq.T_LA",
  "s.P_LA",
  "s.P_LW",
  "s.P_LH",
  "s.T_RA",
  "s.sq.T_RA",
  "s.P_RA",
  "s.T_RH",
  "s.P_RH",
  "s.T_RC",
  "s.T_RW",
  "s.T_RD",
  "s.sq.T_RC",
  "s.P_RW",
  "s.P_RD",
  "s.sq.T_RH",
  "s.sq.P_RA",
  "s.sq.P_RW",
  "s.sq.P_RD")


possible_scenarios$FRI <- rep(1:11, 2) # we want FRI to range from 1-11
possible_scenarios$s.TBF <- scale(rep(c(0, 10), each= 11) ,# raw TBF values
                                  center= attr(TBF_long$s.TBF,"scaled:center"), 
                                  scale= attr(TBF_long$s.TBF,"scaled:scale")) # for each set of FRI's, we want to set a short TBF (of 3) and a long TBF (of 11)

possible_scenarios[, c(  "s.T_LA",  "s.T_LH",  "s.sq.T_LH",  "s.sq.P_LA",  "s.T_LC",  "s.T_LD",  "s.P_LD",  "s.sq.T_LA",  "s.P_LA",  
                         "s.P_LW",  "s.P_LH",  "s.T_RA",  "s.sq.T_RA",  "s.P_RA",  "s.T_RH",  "s.P_RH",  "s.T_RC",  "s.T_RW",  "s.T_RD",
                         "s.sq.T_RC","s.P_RW","s.P_RD", "s.sq.T_RH",  "s.sq.P_RA",  "s.sq.P_RW",  "s.sq.P_RD")] <- 0 # setting all climate variable to their mean

my_sites <- unique(TBF_long_FB$site) 
no_reps <- 1000 # or somethign similar

########### extract model components for later predict() ###############
sur_model_terms <- terms(sur_mod_FB)
sur_term <- attr(sur_model_terms, "term.labels")
sur_term <- reformulate(sur_term)

gr_model_terms <- terms(growth_mod_FB)
gr_term <- attr(gr_model_terms, "term.labels")
gr_term <- reformulate(gr_term)

prep_model_terms <- terms(prep_mod_FB)
prep_term <- attr(prep_model_terms, "term.labels")
prep_term <- reformulate(prep_term)

crep_model_terms <- terms(crep_mod_FB)
crep_term <- attr(crep_model_terms, "term.labels")
crep_term <- reformulate(crep_term)

vargrowth_model_terms <- terms(vargrowth_mod_FB)
vargrowth_term <- attr(vargrowth_model_terms, "term.labels")
vargrowth_term <- reformulate(vargrowth_term)

rec_model_terms <- terms(recruit_mod_FB)
rec_term <- attr(rec_model_terms, "term.labels")
rec_term <- reformulate(rec_term)
```

```{r}
all_pop_growth_rates_FB_reg <- matrix(data= NA, nrow= dim(possible_scenarios)[1], ncol= no_reps)
n_cores <- detectCores()
cl <- makeCluster(n_cores - 1)
registerDoParallel(cl)
j_pop_gr_rate_FB_reg <- rep(NA, dim(possible_scenarios)[1]) ### changed to no_scenarios

all_pop_growth_rates_FB_reg <- foreach(
  j = 1:no_reps,
  .combine = cbind,
  .packages = c("MASS", "lme4", "popbio", "dplyr")
) %dopar% {
  rsur_coefs <- coefficients(sur_mod_FB) # first, assign all "randomy selected" coefficients to the mean estimated coefficeints
  rsur_coefs[!names(coefficients(sur_mod_FB)) %in% paste("site", my_sites, sep="")] <- MASS::mvrnorm(mu= coefficients(sur_mod_FB), Sigma = vcov(sur_mod_FB))[!names(coefficients(sur_mod_FB)) %in% paste("site", my_sites, sep="")]
  #rsur_coefs <- lme4::fixef(survival_mod)
  #rsur_coefs <-  MASS::mvrnorm(mu= lme4::fixef(survival_mod) , Sigma = vcov(survival_mod))
  rgr_coefs <- lme4::fixef(growth_mod_FB)
  rgr_coefs <-  MASS::mvrnorm(mu= lme4::fixef(growth_mod_FB) , Sigma = vcov(growth_mod_FB)) # no site effect in this model
  rprep_coefs <- coefficients(prep_mod_FB)
  rprep_coefs[!names(coefficients(prep_mod_FB)) %in% paste("site", my_sites, sep="")] <- MASS::mvrnorm(mu= coefficients(prep_mod_FB), Sigma = vcov(prep_mod_FB))[!names(coefficients(prep_mod_FB)) %in% paste("site", my_sites, sep="")]
  rcrep_coefs <- lme4::fixef(crep_mod_FB)
  rcrep_coefs <-  MASS::mvrnorm(mu= lme4::fixef(crep_mod_FB) , Sigma = vcov(crep_mod_FB)) # no site effect in this model
  rvar_coefs <- lme4::fixef(vargrowth_mod_FB)
  rvar_coefs <-  MASS::mvrnorm(mu= lme4::fixef(vargrowth_mod_FB) , Sigma = vcov(vargrowth_mod_FB)) # no site effect in this model
  rrec_coefs <- lme4::fixef(recruit_mod_FB)
  rrec_coefs <-  MASS::mvrnorm(mu= lme4::fixef(recruit_mod_FB) , Sigma = vcov(recruit_mod_FB)) # no site effect in this model
  #rrec_coefs <-  MASS::mvrnorm(mu= lme4::fixef(recruit_mod) , Sigma = vcov(recruit_mod)) 
  
  for (i in 1: dim(possible_scenarios)[1]) {
    
    i_scenario <- possible_scenarios[i,] # which combination of FRI, TBF, and climate do I want to use
    
    TSFs <- 0:(i_scenario$FRI-1) # TSF range from 0 to (FRI-1) such that an FRI of 1 will always have TSF= 0
    i_scenario <- i_scenario[rep(seq_len(nrow(i_scenario)), length(TSFs)), ]
    i_scenario$TSF <- TSFs
    i_scenario$s.TSF <- scale(TSFs, center= attr(TBF_long$s.TSF,"scaled:center"), 
                              scale= attr(TBF_long$s.TSF,"scaled:scale"))
    i_scenario$s.sq.TSF <- scale(TSFs^2, center= attr(TBF_long$s.sq.TSF,"scaled:center"), 
                                 scale= attr(TBF_long$s.sq.TSF,"scaled:scale"))
    
    mx_prod <- diag(no_bins+1) # setting up an identity matrix before you enter the loop
    
    for (ii in TSFs){
      i_ii_scenario <- i_scenario[which(i_scenario$TSF==ii),]
      data_for_prediction <- i_ii_scenario[rep(seq_len(nrow(i_ii_scenario)), length(bin_mids)), ]
      data_for_prediction$s.logsize0 <-   scale(bin_mids, center= attr(TBF_long$s.logsize0,"scaled:center"), 
                                                scale= attr(TBF_long$s.logsize0,"scaled:scale"))
      data_for_prediction_rec <- cbind(i_ii_scenario)
      
      predicted_sur <- predicted_gr <-predicted_pr <- predicted_cr <- predicted_vg <- matrix(data=NA, nrow= length(my_sites), ncol= length(bin_mids))
      predicted_rec <- rep(NA, length(my_sites))
      
      ### make empty mxes before entering the site loop
      
      predicted_sur_allsites <-predicted_gr_allsites<-predicted_pr_allsites <-predicted_cr_allsites<-predicted_vg_allsites <-matrix(NA, nrow = length(my_sites), ncol = no_bins, dimnames = list(my_sites, NULL))
      predicted_rec_allsites <- matrix(NA, nrow = length(my_sites), ncol = 1, dimnames = list(my_sites, NULL))
      
      
      for (s in my_sites){ #looping through the sites
        s_data_for_prediction <- data_for_prediction
        s_data_for_prediction_rec <- data_for_prediction_rec
        s_data_for_prediction$site <- s
        s_data_for_prediction_rec$site <- s
        
        s_data_for_prediction$site     <- factor(s_data_for_prediction$site,  levels = my_sites)   ## it is important to add the levels=my_sites here to ensure that the model matrix is built correctly
        s_data_for_prediction_rec$site <- factor(s_data_for_prediction_rec$site, levels = my_sites)
        
        s_data_for_prediction <- s_data_for_prediction %>% mutate(across(c("s.logsize0", "s.sq.TSF", "s.TSF", "s.TBF"), as.numeric))   
        
        m_sur <- model.matrix(sur_term ,data=s_data_for_prediction)
        p2_sur <- rsur_coefs %*% t(m_sur) # check on whether this coef(survival_mod is a column or a row!!!!!!!!
        p2_sur <- plogis(p2_sur)   ## transform to binomial 
        predicted_sur_allsites[s,] <- p2_sur
        
        m_gr <- model.matrix(gr_term, data=s_data_for_prediction)
        p2_gr <- rgr_coefs %*% t(m_gr) # check on whether this coef(survival_mod is a column or a row!!!!!!!!
        predicted_gr_allsites[s,] <- p2_gr
        
        m_prep <- model.matrix(prep_term,data=s_data_for_prediction)
        p2_prep <- rprep_coefs %*% t(m_prep) # check on whether this coef(survival_mod is a column or a row!!!!!!!!
        p2_prep <- plogis(p2_prep)   ## transform to binomial 
        predicted_pr_allsites[s,] <- p2_prep
        
        m_crep <- model.matrix(crep_term  ,data=s_data_for_prediction)
        p2_crep <- rcrep_coefs %*% t(m_crep) # check on whether this coef(survival_mod is a column or a row!!!!!!!!
        predicted_cr_allsites[s,] <- p2_crep
        
        m_vg <- model.matrix(vargrowth_term,data=s_data_for_prediction)
        p2_vg <- rvar_coefs %*% t(m_vg) # 
        predicted_vg_allsites[s,] <- exp(p2_vg)  
        
        m_rec <- model.matrix(rec_term , data=s_data_for_prediction_rec)
        p2_rec <- rrec_coefs %*% t(m_rec) # check on whether this coef(survival_mod is a column or a row!!!!!!!!
        predicted_rec_allsites[s,1] <- p2_rec*log(1.1)            #### account for composite rec variable (rec/log.fr)
        
        
        # for mixed models, the above functions will predict site-specific results (i.e., will incorporate random effects)
        rm(s_data_for_prediction) # this part removes the site-specific dataframe used in this loop, so you do not accidentaly use the previous sites' dataframe instead of the current one-- in case there was an error somehow
      }
      predicted_sur <- apply(predicted_sur_allsites, MARGIN= 2, FUN= mean, na.rm=T) # average size-specific vital rates across sites
      predicted_gr <- apply(predicted_gr_allsites, MARGIN= 2, FUN= mean, na.rm=T)
      predicted_pr <- apply(predicted_pr_allsites, MARGIN= 2, FUN= mean, na.rm=T)
      predicted_cr <- apply(predicted_cr_allsites, MARGIN= 2, FUN= mean, na.rm=T)
      predicted_vg <- apply(predicted_vg_allsites, MARGIN= 2, FUN= mean, na.rm=T)
      predicted_vg[which(predicted_vg<0)] <- 0                 #### to prevent negative variance
      predicted_rec <- mean(predicted_rec_allsites, na.rm=T)
      
      gmx <- matrix(NA, length(bin_mids),length(bin_mids)) ## create empty gmx before entering the no_bins loop
      
      # make the TSF-specific matrix
      for (ss in 1:no_bins) {
        growcdf <- pnorm(bin_edges,predicted_gr[ss],sqrt(predicted_vg[ss])) # Yes. Vargrowth=(pred - obs)^2
        grows <- growcdf[2:length(bin_edges)]-growcdf[1:(length(bin_edges)-1)]
        if(sum(grows)>0){grows <- grows/sum(grows)
        gmx[,ss] <- grows
        } else if (sum(grows)==0 & predicted_gr[ss]< min(bin_edges)) { 
          gmx[,ss] <- c(1, rep(0, length(bin_mids)-1))  
        } else if (sum(grows)==0 & predicted_gr[ss]> max(bin_edges)) { 
          gmx[,ss] <- c(rep(0, length(bin_mids)-1), 1)  } else { # deals with complete eviction-- OUT OF THE BOUNDS  of the kernel
            gmx[,ss] <- NA} 
        # this if statement breaks the code (puts NA's into the matrix) if the sum of the PDF is zero (which happens if all the probability is outside of the size bounds)
      } # end ss loop
      
      # make the surv*growth mx
      survgmx <- gmx*t(matrix( rep(predicted_sur,no_bins),no_bins)) # Yes, this is doing element-wise multiplication 
      reprow <-  predicted_sur*predicted_pr*predicted_cr # prob that you survive, reproduce, then number fruits|reproduce; this multiplcation does it element by element
      
      mx <- matrix(0, no_bins+1, no_bins+1)
      mx[2:(no_bins+1), 2:(no_bins+1)] = survgmx
      mx[1,2:(no_bins+1)] = reprow # FRUIT ROW
      
      sdlng_cdf <- pnorm(bin_edges,mean_seedling_logsize,sqrt(var_seedling_logsize)) 
      sdlng_pdf <- sdlng_cdf[2:length(bin_edges)]-sdlng_cdf[1:(length(bin_edges)-1)]
      if(sum(sdlng_pdf)>0){
        sdlng_pdf <- sdlng_pdf/sum(sdlng_pdf)} else {
          sdlng_pdf <- NA} 
      
      mx[2:(no_bins+1),1]  = predicted_rec *sdlng_pdf 
      
      mx_prod <- mx_prod %*% mx # in the first iteration of the loop, mx just becomes mx_prod
      
    } # closes ii (TSFs) loop
    
    j_pop_gr_rate_FB_reg[i] <- (popbio::lambda(mx_prod))^(1/length(TSFs)) 
    # multiply them together to get a matrix, etc. 
  }
    j_pop_gr_rate_FB_reg
}


```

#### 2.3.3 graph

```{r}
all_pop_growth_rates_FB_reg <- cbind(possible_scenarios, all_pop_growth_rates_FB_reg)
names(all_pop_growth_rates_FB_reg) <- c(names(possible_scenarios), paste("rep", 1:no_reps, sep = ""))
rep_cols <- grep("^rep", names(all_pop_growth_rates_FB_reg), value = TRUE)

mean_lambda_across_scenarios_FB_reg <- all_pop_growth_rates_FB_reg %>%
  rowwise() %>%
  mutate(
    mean_lambda = mean(c_across(all_of(rep_cols)), na.rm = TRUE),
    lower_bound = quantile(c_across(all_of(rep_cols)), 0.025, na.rm = TRUE),
    upper_bound = quantile(c_across(all_of(rep_cols)), 0.975, na.rm = TRUE)
  ) %>%
  dplyr::select(FRI,s.TBF,mean_lambda,lower_bound, upper_bound) %>%
  mutate(TBF=NA) %>%
  ungroup()

mean_lambda_across_scenarios_FB_reg$TBF[1:11] <- "Short TBF"
mean_lambda_across_scenarios_FB_reg$TBF[12:22] <- "Long TBF"

##########. graph 
##########

plt7<- ggplot(data= mean_lambda_across_scenarios_FB_reg, aes(x= FRI, y= mean_lambda)) +
  geom_line(aes(color= TBF)) + 
  geom_ribbon(aes(ymin = mean_lambda_across_scenarios_FB_reg$lower_bound, ymax = mean_lambda_across_scenarios_FB_reg$upper_bound, fill= TBF), alpha = 0.1) + 
  #ylim(0,1) +
  #xlim(c(0, 12)) +
  labs(y= "Growth Rate", x = "Fire Return Interval",color = NULL, fill = NULL) +
  theme_bw() + theme(legend.position='none') +
  theme(legend.position = "right") +
  scale_x_continuous(breaks = 0:10) +
  theme(text = element_text(size = 16)) 
plt7
```

## 3. Result visualization

Now we have created three region specific IPMs. For the final step, we want to compare predictions of sub-IPMs to see how good our global IPM is at capturing region level responses of plant population dynamics to fire.

```{r}
lambda_regionalIPM <- bind_rows(
  mean_lambda_across_scenarios %>% mutate(dataset = "global IPM"),
  mean_lambda_across_scenarios_Cro_reg %>% mutate(dataset = "Croatan"),
  mean_lambda_across_scenarios_GSP_reg %>% mutate(dataset = "Green Swamp"),
  mean_lambda_across_scenarios_FB_reg %>% mutate(dataset = "Fort Bragg"))

plt8 <- ggplot(lambda_regionalIPM, aes(x = FRI, y = mean_lambda)) +
    geom_ribbon(
    aes(ymin = lower_bound, ymax = upper_bound,
        fill = dataset, group = interaction(dataset, TBF)),
    alpha = 0.1) +
  geom_line(
    aes(color = dataset, linetype = TBF,
        group = interaction(dataset, TBF)),
    size = 1) +
  theme_bw() +
  labs(
    y = "Growth Rate",
    x = "Fire Return Interval",
    color = "Dataset",
    fill = "Dataset",
    linetype = "TBF") +
  scale_x_continuous(breaks = 0:10) +
  theme(text = element_text(size = 16))

plt8
```
