## load packages
```{r}
library(tidyverse)
library(vegan)
library(MASS)
library(ape)
library(dplyr)
library(readxl)
library(corrplot)
library(rpart.plot)


```
## Western Ag data interpretation help link 

### https://www.westernag.ca/innovations/customer/interpretation

## load data

```{r}
data1 <- read_excel("fire nutrient/raw_data/Nutrient Supply Rate Data Project_2605_first_batch_2025.xlsx", skip = 5)
data2 <- read_excel("fire nutrient/raw_data/Nutrient Supply Rate Data_Project 2605_second_batch_2025.xlsx", skip = 5)

## extract detection limit

detection_limit <- data1 %>%
  slice(1) %>%
  dplyr::select(-(2:7))

## then remove detection limit column and merge 
data1 <- data1[-1,]
data2 <- data2[-1,]

data <- bind_rows(data1, data2)

## create site
data <- data %>%
  mutate(
    site = str_remove(`Sample ID`, " #\\d+$"),
    site = if_else(
      str_detect(site, "^HS Site"),
      str_remove(site, " B\\d+$"),
      site
    )
  )

table(data$site)

#### fix anything werid by hand 

data$site[which(data$site == "HS SIte1 B3")]<- "HS Site1"
table(data$site)

```
####add TSF and TBF (update to present)

######add fire histories for demographic site
```{r}

fire_dates_landscape <- openxlsx::read.xlsx("F:/VFT/VFT_github/zyao78VFTcode/fire nutrient/raw_data/fire history/excel fire histories.xlsx", sheet= "landscape", startRow=2,detectDates=TRUE)
sites <- unique(fire_dates_landscape$site)
fire_histories <- tidyr::expand_grid(record_type = "landscape", site = sites, startyear = 1986:2025, TSF = NA, TBF = NA)


end_date <-  openxlsx::read.xlsx("F:/VFT/VFT_github/zyao78VFTcode/fire nutrient/raw_data/fire history/excel fire histories.xlsx", sheet= "landscape", rows=1,colNames= FALSE) %>% stringr::str_remove("FILE CURRENT AS OF ") %>% as.Date("%m/%d/%y")
present_year <- as.numeric(year(end_date)[1])

## run loop

for (i in 1:length(sites)){  
  fire_dates_j_i <- fire_dates_landscape[which(fire_dates_landscape$site == sites[i]), ]
  first_fire_year <- min(fire_dates_j_i$startyear)
  
  #NB that if a fire occurs in startyear 2008, that means that a fire occurred over the 2008-2009 interval. 
  TSF <- NA
  end_startyear <- present_year
  for (ii in first_fire_year:end_startyear){
    if (ii %in% fire_dates_j_i$startyear) {TBF <- TSF # if a fire happens, TSF becomes TBF
    TSF <- 0 # a TSF of zero indicates that a fire burned over the startyear to startyear+ 1 interval
    
    } else {TSF <- TSF +1
    }
    fire_histories$TSF[which(fire_histories$ startyear == ii & fire_histories$site == sites[i] )] <- TSF
    fire_histories$TBF[which(fire_histories$ startyear == ii & fire_histories$site == sites[i])] <- TBF
    #NB that if a fire occurs in startyear 2008, that means that a fire occurred over the 2008-2009 interval.  
  }
}

# ## add TSF to community data
# data$TSF <- NA
# data$TBF <- NA
# 
# for (i in 1: nrow(data)) {
#   site_i = data$site[i]
#   startyear_i= 2025
#   
#   data$TSF[i] <- ifelse( site_i %in% sites, fire_histories$TSF[which(fire_histories$site == site_i & fire_histories$startyear == startyear_i)], NA ) 
#   data$TBF[i] <- ifelse( site_i %in% sites, fire_histories$TBF[which(fire_histories$site == site_i & fire_histories$startyear == startyear_i)], NA )
# }



```
###### add holly shelter fire records
```{r}
HS_fire <- read.csv("F:/VFT/VFT_github/zyao78VFTcode/fire nutrient/raw_data/fire history/HS fire history.csv")
June15DOY <- cropgrowdays::day_of_year(as.Date("2020-06-15"))
HS_fire_histories <- tidyr::expand_grid(record_type = "landscape", site = unique(HS_fire$Site), startyear = 1986:2025, TSF = NA, TBF = NA)


HS_fire <- HS_fire %>%
  mutate(
    Burn_Date2 = parse_date_time(Burn_Date, orders = c("mdY HMS p","mdY HMS",  "mdY HM"))) %>%
  mutate (Burn_Date2 = as.Date(Burn_Date2))

  

HS_fire$startyear <- ifelse(cropgrowdays::day_of_year(HS_fire$Burn_Date2) <= June15DOY, # if the day of year was before June 15
                                 as.numeric(format(HS_fire$Burn_Date2, "%Y"))-1,  # then return the previous years number, b/c the fire happened over the interaval of previous to current year
                                 as.numeric(format(HS_fire$Burn_Date2, "%Y")))


####
for (i in 1:length(unique(HS_fire$Site))){  
  fire_dates_j_i <- HS_fire[which(HS_fire$Site == unique(HS_fire$Site)[i]), ]
  first_fire_year <- min(HS_fire$startyear)
  
  #NB that if a fire occurs in startyear 2008, that means that a fire occurred over the 2008-2009 interval. 
  TSF <- NA
  end_startyear <- 2026
  for (ii in first_fire_year:end_startyear){
    if (ii %in% fire_dates_j_i$startyear) {TBF <- TSF # if a fire happens, TSF becomes TBF
    TSF <- 0 # a TSF of zero indicates that a fire burned over the startyear to startyear+ 1 interval
    
    } else {TSF <- TSF +1
    }
    HS_fire_histories$TSF[which(HS_fire_histories$startyear == ii & HS_fire_histories$site == unique(HS_fire$Site)[i]  )] <- TSF

    HS_fire_histories$TBF[which(HS_fire_histories$startyear == ii & HS_fire_histories$site == unique(HS_fire$Site)[i] )] <- TBF
    #NB that if a fire occurs in startyear 2008, that means that a fire occurred over the 2008-2009 interval.  
  }}


###


```

##### bind rows to form one fire history data
```{r}
fire_histories_all <- bind_rows(fire_histories, HS_fire_histories)

for (i in 1: nrow(data)) {
  site_i = data$site[i]
  startyear_i= 2025

  data$TSF[i] <- ifelse( site_i %in% sites,   fire_histories_all$TSF[which(fire_histories_all$site == site_i & fire_histories_all$startyear == startyear_i)], NA )
  

  data$TBF[i] <- ifelse( site_i %in% sites, fire_histories_all$TBF[which(fire_histories_all$site == site_i & fire_histories_all$startyear == startyear_i)], NA )
}



```




## multivariate analyses
###### prepare data into community format
```{r}
sites <- unique(fire_histories_all$site)
# fireCol <- c("TBF", "TSF")
soilnt <- setdiff(colnames(data[,-(1:7)]),c("site", "TBF", "TSF"))   


comm_data <- data %>%
  dplyr:: select(site, all_of(soilnt)) %>%
  group_by(site) %>%
  summarise(across(all_of(soilnt), ~ mean(.x, na.rm = TRUE))) %>%
  column_to_rownames(var = "site")

```
###### make env data 
```{r}
clim <- read.csv("fire nutrient/processed_data/gridMET_climate_24_25.csv",row.names = 1)

annual_clim <- clim %>%
  group_by(site) %>%
  summarize(annual_T = mean(Temp), annual_P = sum(Prec)) %>%
  mutate(site=recode(site, HS_site1="HS Site1", HS_site2="HS Site2", HS_site3="HS Site3", GSP_BI="GSP-BI", GSP_LI="GSP-LI"))

env_data <- data %>%
 dplyr::select(site, all_of(fireCol)) %>%
 group_by(site) %>%
 summarise(across(all_of(fireCol), ~ mean(.x, na.rm = TRUE))) %>%
 column_to_rownames(var = "site")

env_data <- merge(env_data, annual_clim, by.x = "row.names", by.y = "site")
env_data <- env_data %>%
 column_to_rownames(var = "Row.names")



```


###### correlation plot
```{r}
M<-cor(comm_data)

corrplot(M, method="color")
```

###### PCoA
```{r}
distance<-vegdist(comm_data, method="bray")

soil_pcoa <- pcoa(distance, correction="none", rn=NULL)
biplot(soil_pca)

```

###### PCA 

```{r}
soil_pca <- prcomp(
  comm_data,
  center = TRUE,
  scale. = TRUE
)

ordiplot(soil_pca, type="n") |>
  points("species", pch=10) |>
  text("sites", col="red", cex=0.5)
```

###### RDA (constrained ordination)

```{r}
comm_sqrt <- sqrt(comm_data)

## base rda
base_rda<-rda(comm_sqrt ~ env_data$TBF + env_data$TSF + env_data$annual_T + env_data$annual_P)
summary(base_rda)
ordiplot(base_rda, type = "t")
  
```
###### hypothesis testing:run a permanova to identify environmental driver of nutrient composition
```{r}
adonis2(distance ~ env_data$TBF + env_data$TSF + env_data$annual_T + env_data$annual_P, data=env_data, permutations=100,by ="term")


```
###### hypothesis testing: Mantel test for correlation between environment and nutrient multivariate space

Present as to 2026, this Mantel test revealed no correlation between env and nutrient compositions. 
```{r}
distance_Nutrient<-vegdist(comm_data, method="bray")
distance_env<-vegdist(env_data, method="bray")

Mantel  = mantel(distance_Nutrient, distance_env, method = "spearman", permutations  = 9999, na.rm = TRUE)

Mantel
```


###### try some regression trees  

This should be interpreted with caution since there're only 10 sites. 

```{r}
RegTree_data <- merge(comm_data, env_data, by = 0, all.x = TRUE)
RegTree_data <- column_to_rownames(RegTree_data, var = "Row.names")

fit <- rpart(`Total N`  ~ TBF+ TSF + annual_T + annual_P, method="anova", data=RegTree_data, control = rpart.control(minsplit = 2, cp=0))
rpart.plot(fit)

### try other variables

rpart.plot(rpart(`NH4-N`  ~ TBF+ TSF + annual_T + annual_P, method="anova", data=RegTree_data, control = rpart.control(minsplit = 2)))

rpart.plot(rpart(`NH4-N`  ~ TBF+ TSF + annual_T + annual_P, method="anova", data=RegTree_data, control = rpart.control(minsplit = 2)))


```

