
## load
```{r}
library(dplyr)
library(lme4)
library(MuMIn)
library(parallel)
library(doParallel)
library(foreach)
library(AICcmodavg)
library(readr)
library(cAIC4)
```

```{r}
VFT_long <- read_csv("local adaptation/data/processed data/VFT_long_upto2025_2.csv")

VFT_long$s.TSF <- scale(VFT_long$TSF)
VFT_long$s.TBF <- scale(VFT_long$TBF)

```
# explore candidate source of local adaptation (variables that vary among sites)

##plot time series
```{r}
SoilT_list <- c("T_LA", "T_LW", "T_LH", "T_LD", "T_LC")
SoilMoisture_list <- c("P_LA", "P_LD", "P_LH", "P_LW")

RegionalT_list <- c("T_RA", "T_RW", "T_RH", "T_RD", "T_RC")
precip_list <- c("P_RA", "P_RD", "P_RH", "P_RW")

Plot_clim <- function(x, data) {
  ggplot(data,
         aes(x = as.factor(startyear),
             y = .data[[x]],
             colour = site,
             group = site)) +
    geom_line() +
    geom_point() +
    labs(y = x, x = "Year") +
    theme(
      axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1, size = 8)
    ) +
    ggtitle(x)
}


p_soilT <- lapply(SoilT_list, Plot_clim, data = VFT_long)

p_soilT <- ggpubr::ggarrange(
  plotlist = p_soilT[1:5],
  ncol = 3,
  nrow = 2,
  common.legend = TRUE,
  legend = "right"
)

p_soilmoisture <- lapply(SoilMoisture_list, Plot_clim, data = VFT_long)

p_soilmoisture <- ggpubr::ggarrange(
  plotlist = p_soilmoisture[1:5],
  ncol = 3,
  nrow = 2,
  common.legend = TRUE,
  legend = "right"
)


p_precip <- lapply(precip_list, Plot_clim, data = VFT_long)

p_precip <- ggpubr::ggarrange(
  plotlist = p_precip[1:5],
  ncol = 3,
  nrow = 2,
  common.legend = TRUE,
  legend = "right"
)

p_regTemp <- lapply(RegionalT_list, Plot_clim, data = VFT_long)

p_regTemp <- ggpubr::ggarrange(
  plotlist = p_regTemp[1:5],
  ncol = 3,
  nrow = 2,
  common.legend = TRUE,
  legend = "right"
)



print(p_regTemp)
print(p_precip)
print(p_soilT)
print(p_soilmoisture)

```

random effects were selected using rePCA (random effect PCA to explore dimensions (number of random effects) supported by data given 8 sites) https://arxiv.org/pdf/1506.04967. 1) run maximal random effect model, see whether it converges. 2) run rePCA to detect whether the data support this many dimensions of random effects.  3) if not, stepwise remove smallest variance component.



# survival 

#### get best fit model
```{r}
sur_subset_R <- 
  VFT_long %>% 
  filter_at(vars(sur0_1_tp, s.logsize0, s.TSF,s.TBF, site, ), all_vars(!is.na(.)))

GM_R_sur <- glmer(
  sur0_1_tp ~ 
    s.logsize0 + 
    s.TSF * s.TBF +
    s.T_LA + s.T_LH + s.sq.T_LH + s.T_LD + s.P_RH + s.P_RD +
    s.P_RH:s.T_RH + 
    (1|site),
    # (1 + s.T_LA + s.T_LH + s.sq.T_LH + s.T_LD + s.P_RH + s.P_RD | site),
  data = sur_subset_R,
  family = "binomial",
  na.action = "na.fail"
)


n_cores <- detectCores()
n_cores
cluster <- makeCluster(n_cores - 1)
registerDoParallel(cluster)
clusterExport(cluster, c("sur_subset_R"))   ### replace with different subset (different global models)
clusterEvalQ(cluster, {library(lme4); library(MuMIn)})

sur_dredge_R <- MuMIn::dredge(
  GM_R_sur,
  cluster = cluster,
  trace   = 2
)
head(sur_dredge_R) ########### manually check for comparable models

stopCluster(cluster)
sur_mod <- get.models(sur_dredge_R, 1)[[1]]

summary(sur_mod) 

sur_model_terms <- terms(sur_mod)
sur_term <- attr(sur_model_terms, "term.labels")
sur_term <- reformulate(sur_term)

sur_term

## get the best fit model without random slope
```

#### add random slope
```{r}
sur_subset_R <- 
  VFT_long %>% 
  filter_at(vars(sur0_1_tp, s.logsize0, s.TSF,s.TBF, site, ), all_vars(!is.na(.)))

## intercept only
sur_m0 <- glmer(
  sur0_1_tp ~  s.logsize0 + s.P_RD + s.P_RH + s.sq.T_LH + s.T_LA + s.T_LD +
    s.T_LH + s.TBF + s.TSF + s.P_RH:s.T_RH + s.TBF:s.TSF +
    (1 | site),
  data = sur_subset_R,
  family = "binomial",
  na.action = "na.fail",
  # control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5))            ## increase maxfun number fron 10,0000
)

## maximal random effect complexity
sur_m1 <- glmer(
  sur0_1_tp ~  s.logsize0 + s.P_RD + s.P_RH + s.sq.T_LH + s.T_LA + s.T_LD +
    s.T_LH + s.TBF + s.TSF + s.P_RH:s.T_RH + s.TBF:s.TSF +
    (1 + s.T_LA + s.T_LH  + s.T_LD + s.P_RH + s.P_RD | site),
  data = sur_subset_R,
  family = "binomial",
  na.action = "na.fail",
  # control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5))            ## increase maxfun number fron 10,0000
)     

rePCA(sur_m1)  ## after axis 4, standard deviations drops to essentially 0.


## m2: maximal complexity with random slope on quadratic term 
sur_m2 <- glmer(
  sur0_1_tp ~  s.logsize0 + s.P_RD + s.P_RH + s.sq.T_LH + s.T_LA + s.T_LD +
    s.T_LH + s.TBF + s.TSF + s.P_RH:s.T_RH + s.TBF:s.TSF +
    (1 + s.T_LA + s.sq.T_LH  + s.T_LD + s.P_RH + s.P_RD | site),
  data = sur_subset_R,
  family = "binomial",
  na.action = "na.fail",
  # control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5))            ## increase maxfun number fron 10,0000
) 
summary(sur_m2)

## m3: only random slope on temperature covariates

sur_m3 <- glmer(
  sur0_1_tp ~  s.logsize0 + s.P_RD + s.P_RH + s.sq.T_LH + s.T_LA + s.T_LD +
    s.T_LH + s.TBF + s.TSF + s.P_RH:s.T_RH + s.TBF:s.TSF +
    (1 + s.T_LA + s.T_LH  + s.T_LD | site),
  data = sur_subset_R,
  family = "binomial",
  na.action = "na.fail",
  # control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5))            ## increase maxfun number fron 10,0000
)     

## m4: random slope on precip
sur_m4 <- glmer(
  sur0_1_tp ~  s.logsize0 + s.P_RD + s.P_RH + s.sq.T_LH + s.T_LA + s.T_LD +
    s.T_LH + s.TBF + s.TSF + s.P_RH:s.T_RH + s.TBF:s.TSF +
    (1 +    s.P_RH + s.P_RD | site),
  data = sur_subset_R,
  family = "binomial",
  na.action = "na.fail",
  # control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5))            ## increase maxfun number fron 10,0000
)

## m5: just one random slope on s.T_LA (high variance)
sur_m5 <- glmer(
  sur0_1_tp ~  s.logsize0 + s.P_RD + s.P_RH + s.sq.T_LH + s.T_LA + s.T_LD +
    s.T_LH + s.TBF + s.TSF + s.P_RH:s.T_RH + s.TBF:s.TSF +
    (1 +  s.T_LA | site),
  data = sur_subset_R,
  family = "binomial",
  na.action = "na.fail",
  # control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5))            ## increase maxfun number fron 10,0000
)

## 
# sur_modelselection <- cAIC(sur_m0, sur_m1, sur_m2, sur_m3, sur_m4, sur_m5)
# cAIC(sur_m0)

```



# growth 

#### get best fit model
```{r}
gr_subset_R_2 <- 
  VFT_long[which(VFT_long$sur0_1_tp == 1),] %>%      ## make sure to subset data to only alive plants
  filter_at(vars(logsize1, s.logsize0, s.TSF, s.TBF, site, s.sq.T_RA,s.T_RA, s.P_RA, s.T_RC, 
                 s.T_RD, s.P_RH), all_vars(!is.na(.)))
GM_R_gr <- lmer(
  logsize1 ~ s.logsize0 + s.TSF * s.TBF  + s.T_LA + s.sq.T_LA + s.P_RA + s.T_LA:s.P_RA +
    s.T_LC + s.T_LD + s.P_RH +
    (1 | site),
  data = gr_subset_R_2,
  na.action = "na.fail"
)

n_cores <- detectCores()
n_cores
cluster <- makeCluster(n_cores - 1)
registerDoParallel(cluster)
clusterExport(cluster, c("gr_subset_R_2"))   ### replace with different subset (different global models)
clusterEvalQ(cluster, {library(lme4); library(MuMIn)})

gr_dredge_R <- MuMIn::dredge(
  GM_R_gr,
  cluster = cluster,
  trace   = 2
)
head(gr_dredge_R)
growth_mod<- get.models(gr_dredge_R, 1)[[1]]

stopCluster(cluster)

gr_model_terms <- terms(growth_mod)
gr_term <- attr(gr_model_terms, "term.labels")
gr_term <- reformulate(gr_term)

gr_term

```
#### add random slope

```{r}
gr_subset_R_2 <- 
  VFT_long[which(VFT_long$sur0_1_tp == 1),] %>%      ## make sure to subset data to only alive plants
  filter_at(vars(logsize1, s.logsize0, s.TSF, s.TBF, site, s.sq.T_RA,s.T_RA, s.P_RA, s.T_RC, 
                 s.T_RD, s.P_RH), all_vars(!is.na(.)))

## m1: intercept only
gr_m0 <- lmer(
  logsize1 ~ s.logsize0 + s.P_RH + s.sq.T_LA + s.T_LA + s.T_LD + s.TBF + 
    (1  | site),
  data = gr_subset_R_2,
  na.action = "na.fail", 
  # control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5))            ## increase maxfun number fron 10,0000
) 

## m1: maximal complexity, linear

gr_m1 <- lmer(
  logsize1 ~ s.logsize0 + s.P_RH + s.sq.T_LA + s.T_LA + s.T_LD + s.TBF + 
    (1 + s.P_RH +s.T_LA+s.T_LD | site),
  data = gr_subset_R_2,
  na.action = "na.fail", 
  # control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5))            ## increase maxfun number fron 10,0000
)  

## m2: maximal complexity, quadratic
gr_m2 <- lmer(
  logsize1 ~ s.logsize0 + s.P_RH + s.sq.T_LA + s.T_LA + s.T_LD + s.TBF + 
    (1 + s.P_RH +s.sq.T_LA+s.T_LD | site),
  data = gr_subset_R_2,
  na.action = "na.fail", 
  # control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5))            ## increase maxfun number fron 10,0000
)

## m3: random temp 
gr_m3 <- lmer(
  logsize1 ~ s.logsize0 + s.P_RH + s.sq.T_LA + s.T_LA + s.T_LD + s.TBF + 
    (1 +s.T_LA+s.T_LD | site),
  data = gr_subset_R_2,
  na.action = "na.fail", 
  # control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5))            ## increase maxfun number fron 10,0000
)  

## m4: random temp, non-linear

gr_m4 <- lmer(
  logsize1 ~ s.logsize0 + s.P_RH + s.sq.T_LA + s.T_LA + s.T_LD + s.TBF + 
    (1 +s.sq.T_LA+s.T_LD | site),
  data = gr_subset_R_2,
  na.action = "na.fail", 
  # control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5))            ## increase maxfun number fron 10,0000
)

## m5: random precip

gr_m5 <- lmer(
  logsize1 ~ s.logsize0 + s.P_RH + s.sq.T_LA + s.T_LA + s.T_LD + s.TBF + 
    (1 +s.P_RH | site),
  data = gr_subset_R_2,
  na.action = "na.fail", 
  # control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5))            ## increase maxfun number fron 10,0000
)

## model selection

# anocAIC(gr_m0, gr_m1, gr_m2, gr_m3, gr_m4, gr_m5)

```


# prob of reproduction
####get best fit model

```{r}
prep_subset_R <- 
 VFT_long %>% 
  filter_at(vars(prep1, s.logsize0, s.TSF, s.TBF, site, s.T_RA,s.P_RA, s.T_RH, s.T_RC, 
                 s.P_RH, ), all_vars(!is.na(.)))

 ### this likely won't work due to convergence failure, but feel free to try (take a long time to fit)


GM_R_prep <- glmer(
  prep1 ~ s.logsize0 + s.TSF * s.TBF +
    s.T_LA + s.sq.T_LA +s.P_RA+ s.sq.P_RA + s.T_LA:s.P_RA+
    s.T_LH + s.P_RH + s.T_LH:s.P_RH + s.T_LC +
    (1|site),
  data = prep_subset_R,
  family = "binomial", 
  na.action = "na.fail"
)

n_cores <- detectCores()
n_cores
cluster <- makeCluster(n_cores - 1)
registerDoParallel(cluster)
clusterExport(cluster, c("prep_subset_R"))   ### replace with different subset (different global models)
clusterEvalQ(cluster, {library(lme4); library(MuMIn)})

prep_dredge_R_2 <- MuMIn::dredge(
  GM_R_prep,
  cluster = cluster,
  trace   = 2
)
prep_mod <- get.models(prep_dredge_R_2, 1)[[1]]

stopCluster(cluster)

prep_model_terms <- terms(prep_mod)
prep_term <- attr(prep_model_terms, "term.labels")
prep_term <- reformulate(prep_term)

prep_term
```
#### add random slope

```{r}
prep_subset_R <- 
 VFT_long %>% 
  filter_at(vars(prep1, s.logsize0, s.TSF, s.TBF, site, s.T_RA,s.P_RA, s.T_RH, s.T_RC, 
                 s.P_RH, ), all_vars(!is.na(.)))
### m0: intercept only
prep_m0 <- glmer(
  prep1 ~  s.logsize0 + s.P_RA + s.P_RH + s.sq.P_RA + s.sq.T_LA + s.T_LA + 
    s.T_LH + s.TBF + s.TSF + s.P_RH:s.T_LH + s.TBF:s.TSF + (1|site),
  data = prep_subset_R,
  family = "binomial",
  na.action = "na.fail",
  # control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5))            ## increase maxfun number fron 10,0000
)


### m1: maximal complexity, linear climate 

prep_m1 <- glmer(
  prep1 ~  s.logsize0 + s.P_RA + s.P_RH + s.sq.P_RA + s.sq.T_LA + s.T_LA + 
    s.T_LH + s.TBF + s.TSF + s.P_RH:s.T_LH + s.TBF:s.TSF + (1+s.P_RA+s.P_RH+s.T_LA+s.T_LH|site),
  data = prep_subset_R,
  family = "binomial",
  na.action = "na.fail",
  # control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5))            ## increase maxfun number fron 10,0000
)  

### m2: maximal complexity, non linear climate
prep_m2 <- glmer(
  prep1 ~  s.logsize0 + s.P_RA + s.P_RH + s.sq.P_RA + s.sq.T_LA + s.T_LA + 
    s.T_LH + s.TBF + s.TSF + s.P_RH:s.T_LH + s.TBF:s.TSF + (1+s.sq.P_RA+s.P_RH+s.sq.T_LA+s.T_LH|site),
  data = prep_subset_R,
  family = "binomial",
  na.action = "na.fail",
  # control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5))            ## increase maxfun number fron 10,0000
) 

### m3: linear temperature only

prep_m3 <- glmer(
  prep1 ~  s.logsize0 + s.P_RA + s.P_RH + s.sq.P_RA + s.sq.T_LA + s.T_LA + 
    s.T_LH + s.TBF + s.TSF + s.P_RH:s.T_LH + s.TBF:s.TSF + (1+s.T_LA+s.T_LH|site),
  data = prep_subset_R,
  family = "binomial",
  na.action = "na.fail",
  # control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5))            ## increase maxfun number fron 10,0000
) 

### m4: non linear temperature only
prep_m4 <- glmer(
  prep1 ~  s.logsize0 + s.P_RA + s.P_RH + s.sq.P_RA + s.sq.T_LA + s.T_LA + 
    s.T_LH + s.TBF + s.TSF + s.P_RH:s.T_LH + s.TBF:s.TSF + (1+s.sq.T_LA+s.T_LH|site),
  data = prep_subset_R,
  family = "binomial",
  na.action = "na.fail",
  # control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5))            ## increase maxfun number fron 10,0000
) 

### m5: linear precip only
prep_m5 <- glmer(
  prep1 ~  s.logsize0 + s.P_RA + s.P_RH + s.sq.P_RA + s.sq.T_LA + s.T_LA + 
    s.T_LH + s.TBF + s.TSF + s.P_RH:s.T_LH + s.TBF:s.TSF + (1+s.P_RA+s.P_RH|site),
  data = prep_subset_R,
  family = "binomial",
  na.action = "na.fail",
  # control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5))            ## increase maxfun number fron 10,0000
)
### m6: non linear precip only
prep_m6 <- glmer(
  prep1 ~  s.logsize0 + s.P_RA + s.P_RH + s.sq.P_RA + s.sq.T_LA + s.T_LA + 
    s.T_LH + s.TBF + s.TSF + s.P_RH:s.T_LH + s.TBF:s.TSF + (1+s.sq.P_RA+s.P_RH|site),
  data = prep_subset_R,
  family = "binomial",
  na.action = "na.fail",
  # control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5))            ## increase maxfun number fron 10,0000
) 

# prep_modelselection <- model.sel(prep_mod, prep_m1, prep_m2, prep_m3, prep_m4, prep_m5, prep_m6)

```

# number of fruits

since previous work has shown that Crep does not depend on fire, I remove fire indices and instead create a new global model that is purely climate driven (with a complete set of climate variables).
#### get best fit model
```{r}
crep_subset_R <- 
  VFT_long %>% 
  filter_at(vars(logcrep1, s.logsize0, TSF, TBF, site, s.T_RA, s.T_RH), all_vars(!is.na(.)))

GM_R_crep <- lmer(
  logcrep1 ~ s.logsize0 + s.P_RA  + s.sq.P_RA + s.P_RH + s.P_RW + s.P_RD +s.sq.P_RD + s.sq.T_LA + s.T_LA + s.T_LH +s.sq.T_LH + s.T_LD + s.T_LC + s.sq.T_LC+ s.P_RH:s.T_LH + s.T_LD:s.P_RD +
    (1 | site),
  data = crep_subset_R,
  na.action = "na.fail"
)

n_cores <- detectCores()
n_cores
cluster <- makeCluster(n_cores - 1)
registerDoParallel(cluster)
clusterExport(cluster, c("crep_subset_R"))   ### replace with different subset (different global models)
clusterEvalQ(cluster, {library(lme4); library(MuMIn)})

crep_dredge_R <- MuMIn::dredge(
  GM_R_crep,
  cluster = cluster,
  trace   = 2
)
head(crep_dredge_R)
crep_mod <- get.models(crep_dredge_R, 1)[[1]]

stopCluster(cluster)

crep_model_terms <- terms(crep_mod)
crep_term <- attr(crep_model_terms, "term.labels")
crep_term <- reformulate(crep_term)

crep_term
```

#### add random slope

```{r}

crep_subset_R <- 
  VFT_long %>% 
  filter_at(vars(logcrep1, s.logsize0, TSF, TBF, site, s.T_RA, s.T_RH), all_vars(!is.na(.)))

### m0: intercept only
crep_m0 <- lmer(
  logcrep1 ~ s.logsize0 + s.P_RA + s.P_RH + s.sq.P_RD + s.sq.T_LA + s.sq.T_LH + 
    s.T_LA + s.T_LH + 
    (1  | site),
  data = crep_subset_R,
  na.action = "na.fail"
)

#### m1: maximal complexity, linear (delete if is singular fit)
crep_m1 <- lmer(
  logcrep1 ~ s.logsize0 + s.P_RA + s.P_RH + s.sq.P_RD + s.sq.T_LA + s.sq.T_LH + 
    s.T_LA + s.T_LH + 
    (1  + s.P_RA + s.P_RH + s.sq.P_RD +s.T_LA + s.T_LH | site),
  data = crep_subset_R,
  na.action = "na.fail"
)

summary(crep_m1)

#### m2: maximal complexity, non linear 

crep_m2 <- lmer(
  logcrep1 ~ s.logsize0 + s.P_RA + s.P_RH + s.sq.P_RD + s.sq.T_LA + s.sq.T_LH + 
    s.T_LA + s.T_LH + 
    (1  + s.P_RA + s.P_RH + s.sq.P_RD + s.sq.T_LA     | site),
  data = crep_subset_R,
  na.action = "na.fail"
)
summary(crep_m2)

lme4::rePCA(sur_m2)

```

## model selection 

```{r}
getwd()
```

