
## load
```{r}
library(dplyr)
library(lme4)
library(MuMIn)
library(parallel)
library(doParallel)
library(foreach)
library(AICcmodavg)
library(readr)
library(cAIC4)
library(performance)
library(DHARMa)
library(sjPlot)
library(insight)
library(MASS)
library(ggplot2)
library(merTools)
library(merTools)


```

```{r}
VFT_long <- read_csv("local adaptation/data/processed data/VFT_long_upto2025_2.csv")

VFT_long$s.TSF <- scale(VFT_long$TSF)
VFT_long$s.TBF <- scale(VFT_long$TBF)

```
## explore candidate source of local adaptation (variables that vary among sites)

#### plot time series
```{r}
SoilT_list <- c("T_LA", "T_LW", "T_LH", "T_LD", "T_LC")
SoilMoisture_list <- c("P_LA", "P_LD", "P_LH", "P_LW")

RegionalT_list <- c("T_RA", "T_RW", "T_RH", "T_RD", "T_RC")
precip_list <- c("P_RA", "P_RD", "P_RH", "P_RW")

Plot_clim <- function(x, data) {
  ggplot(data,
         aes(x = as.factor(startyear),
             y = .data[[x]],
             colour = site,
             group = site)) +
    geom_line() +
    geom_point() +
    labs(y = x, x = "Year") +
    theme(
      axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1, size = 8)
    ) +
    ggtitle(x)
}


p_soilT <- lapply(SoilT_list, Plot_clim, data = VFT_long)

p_soilT <- ggpubr::ggarrange(
  plotlist = p_soilT[1:5],
  ncol = 3,
  nrow = 2,
  common.legend = TRUE,
  legend = "right"
)

p_soilmoisture <- lapply(SoilMoisture_list, Plot_clim, data = VFT_long)

p_soilmoisture <- ggpubr::ggarrange(
  plotlist = p_soilmoisture[1:5],
  ncol = 3,
  nrow = 2,
  common.legend = TRUE,
  legend = "right"
)


p_precip <- lapply(precip_list, Plot_clim, data = VFT_long)

p_precip <- ggpubr::ggarrange(
  plotlist = p_precip[1:5],
  ncol = 3,
  nrow = 2,
  common.legend = TRUE,
  legend = "right"
)

p_regTemp <- lapply(RegionalT_list, Plot_clim, data = VFT_long)

p_regTemp <- ggpubr::ggarrange(
  plotlist = p_regTemp[1:5],
  ncol = 3,
  nrow = 2,
  common.legend = TRUE,
  legend = "right"
)



print(p_regTemp)
print(p_precip)
print(p_soilT)
print(p_soilmoisture)

```

random effects were selected using rePCA (random effect PCA to explore dimensions (number of random effects) supported by data given 8 sites) https://arxiv.org/pdf/1506.04967. 1) run maximal random effect model, see whether it converges. 2) run rePCA to detect whether the data support this many dimensions of random effects.  3) if not, stepwise remove smallest variance component.



# survival 

#### get best fit model
```{r}
sur_subset_R <- 
  VFT_long %>% 
  filter_at(vars(sur0_1_tp, s.logsize0, s.TSF,s.TBF, site, ), all_vars(!is.na(.)))

GM_R_sur <- glmer(
  sur0_1_tp ~ 
    s.logsize0 + 
    s.TSF * s.TBF +
    s.T_LA + s.T_LH + s.sq.T_LH + s.T_LD + s.P_RH + s.P_RD +
    s.P_RH:s.T_RH + 
    (1|site),
    # (1 + s.T_LA + s.T_LH + s.sq.T_LH + s.T_LD + s.P_RH + s.P_RD | site),
  data = sur_subset_R,
  family = "binomial",
  na.action = "na.fail"
)


n_cores <- detectCores()
n_cores
cluster <- makeCluster(n_cores - 1)
registerDoParallel(cluster)
clusterExport(cluster, c("sur_subset_R"))   ### replace with different subset (different global models)
clusterEvalQ(cluster, {library(lme4); library(MuMIn)})

sur_dredge_R <- MuMIn::dredge(
  GM_R_sur,
  cluster = cluster,
  trace   = 2
)
head(sur_dredge_R) ########### manually check for comparable models

stopCluster(cluster)
sur_mod <- get.models(sur_dredge_R, 1)[[1]]

summary(sur_mod) 

sur_model_terms <- terms(sur_mod)
sur_term <- attr(sur_model_terms, "term.labels")
sur_term <- reformulate(sur_term)

sur_term

## get the best fit model without random slope
```

#### add random slope
```{r}
sur_subset_R <- 
  VFT_long %>% 
  filter_at(vars(sur0_1_tp, s.logsize0, s.TSF,s.TBF, site, ), all_vars(!is.na(.)))

## intercept only
sur_m0 <- glmer(
  sur0_1_tp ~  s.logsize0 + s.P_RD + s.P_RH + s.sq.T_LH + s.T_LA + s.T_LD +
    s.T_LH + s.TBF + s.TSF + s.P_RH:s.T_LH + s.TBF:s.TSF +
    (1 | site),
  data = sur_subset_R,
  family = "binomial",
  na.action = "na.fail",
  # control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5))            ## increase maxfun number fron 10,0000
)

## maximal random effect complexity
sur_m1 <- glmer(
  sur0_1_tp ~  s.logsize0 + s.P_RD + s.P_RH + s.sq.T_LH + s.T_LA + s.T_LD +
    s.T_LH + s.TBF + s.TSF + s.P_RH:s.T_LH + s.TBF:s.TSF +
    (1 + s.T_LA + s.T_LH  + s.T_LD + s.P_RH + s.P_RD | site),
  data = sur_subset_R,
  family = "binomial",
  na.action = "na.fail",
  # control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5))            ## increase maxfun number fron 10,0000
)     



## m2: maximal complexity with random slope on quadratic term 
sur_m2 <- glmer(
  sur0_1_tp ~  s.logsize0 + s.P_RD + s.P_RH + s.sq.T_LH + s.T_LA + s.T_LD +
    s.T_LH + s.TBF + s.TSF + s.P_RH:s.T_LH + s.TBF:s.TSF +
    (1 + s.T_LA + s.sq.T_LH  + s.T_LD + s.P_RH + s.P_RD | site),
  data = sur_subset_R,
  family = "binomial",
  na.action = "na.fail",
  # control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5))            ## increase maxfun number fron 10,0000
) 
summary(sur_m2)

## m3: only random slope on temperature covariates

sur_m3 <- glmer(
  sur0_1_tp ~  s.logsize0 + s.P_RD + s.P_RH + s.sq.T_LH + s.T_LA + s.T_LD +
    s.T_LH + s.TBF + s.TSF + s.P_RH:s.T_LH + s.TBF:s.TSF +
    (1 + s.T_LA + s.T_LH  + s.T_LD | site),
  data = sur_subset_R,
  family = "binomial",
  na.action = "na.fail",
  # control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5))            ## increase maxfun number fron 10,0000
)     

## m4: random slope on precip
sur_m4 <- glmer(
  sur0_1_tp ~  s.logsize0 + s.P_RD + s.P_RH + s.sq.T_LH + s.T_LA + s.T_LD +
    s.T_LH + s.TBF + s.TSF + s.P_RH:s.T_LH + s.TBF:s.TSF +
    (1 +    s.P_RH + s.P_RD | site),
  data = sur_subset_R,
  family = "binomial",
  na.action = "na.fail",
  # control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5))            ## increase maxfun number fron 10,0000
)

## m5: just one random slope on s.T_LA (high variance)
sur_m5 <- glmer(
  sur0_1_tp ~  s.logsize0 + s.P_RD + s.P_RH + s.sq.T_LH + s.T_LA + s.T_LD +
    s.T_LH + s.TBF + s.TSF + s.P_RH:s.T_LH + s.TBF:s.TSF +
    (1 +  s.T_LA | site),
  data = sur_subset_R,
  family = "binomial",
  na.action = "na.fail",
  # control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5))            ## increase maxfun number fron 10,0000
)

sur_list <- list(
  sur_intercept        = sur_m0,
  sur_max_linear       = sur_m1,
  sur_max_quad         = sur_m2,
  sur_temp_only        = sur_m3,
  sur_precip_only      = sur_m4,
  sur_highest_var_only = sur_m5
)


```



# growth 

#### get best fit model
```{r}
gr_subset_R_2 <- 
  VFT_long[which(VFT_long$sur0_1_tp == 1),] %>%      ## make sure to subset data to only alive plants
  filter_at(vars(logsize1, s.logsize0, s.TSF, s.TBF, site, s.sq.T_RA,s.T_RA, s.P_RA, s.T_RC, 
                 s.T_RD, s.P_RH), all_vars(!is.na(.)))
GM_R_gr <- lmer(
  logsize1 ~ s.logsize0 + s.TSF * s.TBF  + s.T_LA + s.sq.T_LA + s.P_RA + s.T_LA:s.P_RA +
    s.T_LC + s.T_LD + s.P_RH +
    (1 | site),
  data = gr_subset_R_2,
  na.action = "na.fail"
)

n_cores <- detectCores()
n_cores
cluster <- makeCluster(n_cores - 1)
registerDoParallel(cluster)
clusterExport(cluster, c("gr_subset_R_2"))   ### replace with different subset (different global models)
clusterEvalQ(cluster, {library(lme4); library(MuMIn)})

gr_dredge_R <- MuMIn::dredge(
  GM_R_gr,
  cluster = cluster,
  trace   = 2
)
head(gr_dredge_R)
growth_mod<- get.models(gr_dredge_R, 1)[[1]]

stopCluster(cluster)

gr_model_terms <- terms(growth_mod)
gr_term <- attr(gr_model_terms, "term.labels")
gr_term <- reformulate(gr_term)

gr_term

```
#### add random slope

```{r}
gr_subset_R_2 <- 
  VFT_long[which(VFT_long$sur0_1_tp == 1),] %>%      ## make sure to subset data to only alive plants
  filter_at(vars(logsize1, s.logsize0, s.TSF, s.TBF, site, s.sq.T_RA,s.T_RA, s.P_RA, s.T_RC, 
                 s.T_RD, s.P_RH), all_vars(!is.na(.)))

## m1: intercept only
gr_m0 <- lmer(
  logsize1 ~ s.logsize0 + s.P_RH + s.sq.T_LA + s.T_LA + s.T_LD + s.TBF + 
    (1  | site),
  data = gr_subset_R_2,
  na.action = "na.fail", 
  # control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5))            ## increase maxfun number fron 10,0000
) 

## m1: maximal complexity, linear

gr_m1 <- lmer(
  logsize1 ~ s.logsize0 + s.P_RH + s.sq.T_LA + s.T_LA + s.T_LD + s.TBF + 
    (1 + s.P_RH +s.T_LA+s.T_LD | site),
  data = gr_subset_R_2,
  na.action = "na.fail", 
  # control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5))            ## increase maxfun number fron 10,0000
)  

## m2: maximal complexity, quadratic
gr_m2 <- lmer(
  logsize1 ~ s.logsize0 + s.P_RH + s.sq.T_LA + s.T_LA + s.T_LD + s.TBF + 
    (1 + s.P_RH +s.sq.T_LA+s.T_LD | site),
  data = gr_subset_R_2,
  na.action = "na.fail", 
  # control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5))            ## increase maxfun number fron 10,0000
)

## m3: random temp 
gr_m3 <- lmer(
  logsize1 ~ s.logsize0 + s.P_RH + s.sq.T_LA + s.T_LA + s.T_LD + s.TBF + 
    (1 +s.T_LA+s.T_LD | site),
  data = gr_subset_R_2,
  na.action = "na.fail", 
  # control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5))            ## increase maxfun number fron 10,0000
)  

## m4: random temp, non-linear

gr_m4 <- lmer(
  logsize1 ~ s.logsize0 + s.P_RH + s.sq.T_LA + s.T_LA + s.T_LD + s.TBF + 
    (1 +s.sq.T_LA+s.T_LD | site),
  data = gr_subset_R_2,
  na.action = "na.fail", 
  # control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5))            ## increase maxfun number fron 10,0000
)

## m5: random precip

gr_m5 <- lmer(
  logsize1 ~ s.logsize0 + s.P_RH + s.sq.T_LA + s.T_LA + s.T_LD + s.TBF + 
    (1 +s.P_RH | site),
  data = gr_subset_R_2,
  na.action = "na.fail", 
  # control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5))            ## increase maxfun number fron 10,0000
)

## m6: highest variance only
gr_m6 <- lmer(
  logsize1 ~ s.logsize0 + s.P_RH + s.sq.T_LA + s.T_LA + s.T_LD + s.TBF + 
    (1 + s.T_LA | site),
  data = gr_subset_R_2,
  na.action = "na.fail", 
  # control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5))            ## increase maxfun number fron 10,0000
)

## model selection

gr_list <- list(
  gr_intercept        = gr_m0,
  gr_max_linear       = gr_m1,
  gr_max_quad         = gr_m2,
  gr_temp_only        = gr_m3,
  gr_temp_only_quad   = gr_m4,
  gr_precip_only      = gr_m5,
  gr_highest_var_only = gr_m6
)
```


# prob of reproduction
####get best fit model

```{r}
prep_subset_R <- 
 VFT_long %>% 
  filter_at(vars(prep1, s.logsize0, s.TSF, s.TBF, site, s.T_RA,s.P_RA, s.T_RH, s.T_RC, 
                 s.P_RH, ), all_vars(!is.na(.)))

 ### this likely won't work due to convergence failure, but feel free to try (take a long time to fit)


GM_R_prep <- glmer(
  prep1 ~ s.logsize0 + s.TSF * s.TBF +
    s.T_LA + s.sq.T_LA +s.P_RA+ s.sq.P_RA + s.T_LA:s.P_RA+
    s.T_LH + s.P_RH + s.T_LH:s.P_RH + s.T_LC +
    (1|site),
  data = prep_subset_R,
  family = "binomial", 
  na.action = "na.fail"
)

n_cores <- detectCores()
n_cores
cluster <- makeCluster(n_cores - 1)
registerDoParallel(cluster)
clusterExport(cluster, c("prep_subset_R"))   ### replace with different subset (different global models)
clusterEvalQ(cluster, {library(lme4); library(MuMIn)})

prep_dredge_R_2 <- MuMIn::dredge(
  GM_R_prep,
  cluster = cluster,
  trace   = 2
)
prep_mod <- get.models(prep_dredge_R_2, 1)[[1]]

stopCluster(cluster)

prep_model_terms <- terms(prep_mod)
prep_term <- attr(prep_model_terms, "term.labels")
prep_term <- reformulate(prep_term)

prep_term
```
#### add random slope

```{r}
prep_subset_R <- 
 VFT_long %>% 
  filter_at(vars(prep1, s.logsize0, s.TSF, s.TBF, site, s.T_RA,s.P_RA, s.T_RH, s.T_RC, 
                 s.P_RH, ), all_vars(!is.na(.)))
### m0: intercept only
prep_m0 <- glmer(
  prep1 ~  s.logsize0 + s.P_RA + s.P_RH + s.sq.P_RA + s.sq.T_LA + s.T_LA + 
    s.T_LH + s.TBF + s.TSF + s.P_RH:s.T_LH + s.TBF:s.TSF + (1|site),
  data = prep_subset_R,
  family = "binomial",
  na.action = "na.fail",
  # control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5))            ## increase maxfun number fron 10,0000
)


### m1: maximal complexity, linear climate 

prep_m1 <- glmer(
  prep1 ~  s.logsize0 + s.P_RA + s.P_RH + s.sq.P_RA + s.sq.T_LA + s.T_LA + 
    s.T_LH + s.TBF + s.TSF + s.P_RH:s.T_LH + s.TBF:s.TSF + (1+s.P_RA+s.P_RH+s.T_LA+s.T_LH|site),
  data = prep_subset_R,
  family = "binomial",
  na.action = "na.fail",
  # control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5))            ## increase maxfun number fron 10,0000
)  

### m2: maximal complexity, non linear climate
prep_m2 <- glmer(
  prep1 ~  s.logsize0 + s.P_RA + s.P_RH + s.sq.P_RA + s.sq.T_LA + s.T_LA + 
    s.T_LH + s.TBF + s.TSF + s.P_RH:s.T_LH + s.TBF:s.TSF + (1+s.sq.P_RA+s.P_RH+s.sq.T_LA+s.T_LH|site),
  data = prep_subset_R,
  family = "binomial",
  na.action = "na.fail",
  # control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5))            ## increase maxfun number fron 10,0000
) 

### m3: linear temperature only

prep_m3 <- glmer(
  prep1 ~  s.logsize0 + s.P_RA + s.P_RH + s.sq.P_RA + s.sq.T_LA + s.T_LA + 
    s.T_LH + s.TBF + s.TSF + s.P_RH:s.T_LH + s.TBF:s.TSF + (1+s.T_LA+s.T_LH|site),
  data = prep_subset_R,
  family = "binomial",
  na.action = "na.fail",
  # control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5))            ## increase maxfun number fron 10,0000
) 

### m4: non linear temperature only
prep_m4 <- glmer(
  prep1 ~  s.logsize0 + s.P_RA + s.P_RH + s.sq.P_RA + s.sq.T_LA + s.T_LA + 
    s.T_LH + s.TBF + s.TSF + s.P_RH:s.T_LH + s.TBF:s.TSF + (1+s.sq.T_LA+s.T_LH|site),
  data = prep_subset_R,
  family = "binomial",
  na.action = "na.fail",
  # control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5))            ## increase maxfun number fron 10,0000
) 

### m5: linear precip only
prep_m5 <- glmer(
  prep1 ~  s.logsize0 + s.P_RA + s.P_RH + s.sq.P_RA + s.sq.T_LA + s.T_LA + 
    s.T_LH + s.TBF + s.TSF + s.P_RH:s.T_LH + s.TBF:s.TSF + (1+s.P_RA+s.P_RH|site),
  data = prep_subset_R,
  family = "binomial",
  na.action = "na.fail",
  # control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5))            ## increase maxfun number fron 10,0000
)
### m6: non linear precip only
prep_m6 <- glmer(
  prep1 ~  s.logsize0 + s.P_RA + s.P_RH + s.sq.P_RA + s.sq.T_LA + s.T_LA + 
    s.T_LH + s.TBF + s.TSF + s.P_RH:s.T_LH + s.TBF:s.TSF + (1+s.sq.P_RA+s.P_RH|site),
  data = prep_subset_R,
  family = "binomial",
  na.action = "na.fail",
  # control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5))            ## increase maxfun number fron 10,0000
) 

### m7: highest variance term 
prep_m7 <- glmer(
  prep1 ~  s.logsize0 + s.P_RA + s.P_RH + s.sq.P_RA + s.sq.T_LA + s.T_LA + 
    s.T_LH + s.TBF + s.TSF + s.P_RH:s.T_LH + s.TBF:s.TSF + (1+s.T_LH|site),
  data = prep_subset_R,
  family = "binomial",
  na.action = "na.fail",
  # control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5))            ## increase maxfun number fron 10,0000
)


###
prep_list <- list(
  prep_intercept        = prep_m0,
  prep_max_linear       = prep_m1,
  prep_max_quad         = prep_m2,
  prep_temp_only        = prep_m3,
  prep_temp_only_quad   = prep_m4,
  prep_precip_only      = prep_m5,
  prep_precip_only_quad = prep_m6,
  prep_highest_var_only = prep_m7
)
```

# number of fruits

since previous work has shown that Crep does not depend on fire, I remove fire indices and instead create a new global model that is purely climate driven (with a complete set of climate variables).
#### get best fit model
```{r}
crep_subset_R <- 
  VFT_long %>% 
  filter_at(vars(logcrep1, s.logsize0, TSF, TBF, site, s.T_RA, s.T_RH), all_vars(!is.na(.)))

GM_R_crep <- lmer(
  logcrep1 ~ s.logsize0 + s.P_RA  + s.sq.P_RA + s.P_RH + s.P_RW + s.P_RD +s.sq.P_RD + s.sq.T_LA + s.T_LA + s.T_LH +s.sq.T_LH + s.T_LD + s.T_LC + s.sq.T_LC+ s.P_RH:s.T_LH + s.T_LD:s.P_RD +
    (1 | site),
  data = crep_subset_R,
  na.action = "na.fail"
)

n_cores <- detectCores()
n_cores
cluster <- makeCluster(n_cores - 1)
registerDoParallel(cluster)
clusterExport(cluster, c("crep_subset_R"))   ### replace with different subset (different global models)
clusterEvalQ(cluster, {library(lme4); library(MuMIn)})

crep_dredge_R <- MuMIn::dredge(
  GM_R_crep,
  cluster = cluster,
  trace   = 2
)
head(crep_dredge_R)
crep_mod <- get.models(crep_dredge_R, 1)[[1]]

stopCluster(cluster)

crep_model_terms <- terms(crep_mod)
crep_term <- attr(crep_model_terms, "term.labels")
crep_term <- reformulate(crep_term)

crep_term
```

#### add random slope

```{r}

crep_subset_R <- 
  VFT_long %>% 
  filter_at(vars(logcrep1, s.logsize0, TSF, TBF, site, s.T_RA, s.T_RH), all_vars(!is.na(.)))

### m0: intercept only
crep_m0 <- lmer(
  logcrep1 ~ s.logsize0 + s.P_RA + s.P_RH + s.sq.P_RD + s.sq.T_LA + s.sq.T_LH + 
    s.T_LA + s.T_LH + 
    (1  | site),
  data = crep_subset_R,
  na.action = "na.fail"
)

#### m1: maximal complexity, linear (delete if is singular fit)
crep_m1 <- lmer(
  logcrep1 ~ s.logsize0 + s.P_RA + s.P_RH + s.sq.P_RD + s.sq.T_LA + s.sq.T_LH + 
    s.T_LA + s.T_LH + 
    (1  + s.P_RA + s.P_RH + s.sq.P_RD +s.T_LA + s.T_LH | site),
  data = crep_subset_R,
  na.action = "na.fail"
)

summary(crep_m1)

#### m2: maximal complexity, non linear 

crep_m2 <- lmer(
  logcrep1 ~ s.logsize0 + s.P_RA + s.P_RH + s.sq.P_RD + s.sq.T_LA + s.sq.T_LH + 
    s.T_LA + s.T_LH + 
    (1  + s.P_RA + s.P_RH + s.sq.P_RD + s.sq.T_LA     | site),
  data = crep_subset_R,
  na.action = "na.fail"
)
summary(crep_m2)

lme4::rePCA(sur_m2)

```

## model diagnostics

```{r}


view_resid_dm <- function(mod) {
  mod_name <- deparse(substitute(mod))
  plot(simulateResiduals(fittedModel = mod))
}

view_hist <- function(mod) {
  mod_name <- deparse(substitute(mod))
  hist(resid(mod))
}
```
####sur
```{r}
sur_list <- list(sur_m0, sur_m1, sur_m2, sur_m3, sur_m4, sur_m5)

lapply(sur_list, view_resid_dm)
```
####gr
```{r}
gr_list <- list(gr_m0, gr_m1, gr_m2, gr_m3, gr_m4, gr_m5)

lapply(gr_list, view_hist)
```
####prep
```{r}
prep_list <- list(prep_m0, prep_m1, prep_m2, prep_m3, prep_m4, prep_m5,prep_m6)

lapply(prep_list, view_resid_dm)
```



## model selection

#### check singularity
```{r}


for (i in 1: length(sur_list)) { if (lme4::isSingular(sur_list[[i]]) == "TRUE") { print(names(sur_list[i]))} else {print("no singularity issue")}
}
for (i in 1: length(gr_list)) { if (lme4::isSingular(gr_list[[i]]) == "TRUE") { print(names(gr_list[i]))} else {print("no singularity issue")}
}
for (i in 1: length(prep_list)) { if (lme4::isSingular(prep_list[[i]]) == "TRUE") { print(names(prep_list[i]))} else {print("no singularity issue")}
}

## delete singular models

prep_list <- prep_list[-which(names(prep_list) == "prep_max_linear")]
```


#### check convergence

```{r}

check_mod_convergence <- function(m) {
  w <- m@optinfo$conv$lme4$messages
  if (is.null(w))  {return("no warning message")}
  else if (any(grepl("failed to converge", w))) {return("fail to converge")} 
  else {return("check warning message")}
}

lapply(sur_list, check_mod_convergence)
lapply(gr_list, check_mod_convergence)
lapply(prep_list, check_mod_convergence)

### delete problematic models or refer to the 00_fix_lme4_warnings scripts under Legacy effect/scripts to try to fix it

```


#### compare AICc 

```{r}
## use sapply since we want the output to be a vector not a list
## use double [[]] to extract model itself in a list

sur_bestfit <- sur_list[[which(names(sur_list) == names(which.min(sapply(sur_list, AICc))))]]

gr_bestfit <- gr_list[[which(names(gr_list) == names(which.min(sapply(gr_list, AICc))))]]

prep_bestfit <- prep_list[[which(names(prep_list) == names(which.min(sapply(prep_list, AICc))))]]

```







## graph

#### coefficient plots with random effects (gave up on plotting random effect variance CI). The most likely issue is that the model is too delicate to run Confint() on. 


```{r}
### make random effect plot (visualize random effect sign for each site). Not always useful.

# plot_model(sur_m2, type =  "re")
# # plot_model(gr_bestfit, type =  "re")
# # plot_model(prep_bestfit, type =  "re")

### plot random effect variance

# varcorr <- as.data.frame(VarCorr(sur_bestfit, comp = "Variance"))
# attributes(VarCorr(sur_bestfit)$site)$stddev
# 
# confint(sur_m0)
# confint(sur_m0, parm = "theta_", method = "profile")
# 
# confint(sur_bestfit, method = "boot", nsim = 500)
# 
# 
# mod <- glmer(
#   sur0_1_tp ~  s.logsize0 + s.P_RD + s.P_RH + s.sq.T_LH + s.T_LA + s.T_LD +
#     s.T_LH + s.TBF + s.TSF + s.P_RH:s.T_RH + s.TBF:s.TSF +
#     (1 + s.T_LA + s.sq.T_LH  + s.T_LD + s.P_RH + s.P_RD | site),
#   data = sur_subset_R,
#   family = "binomial",
#   na.action = "na.fail",
#   # control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5))            ## increase maxfun number fron 10,0000
# ) 
# 
# confint(mod, method = "boot", nsim = 500)
# 
# boot_fun <- function(x) fixef(x)
# 
# set.seed(1)
# bootMer(mod, FUN = boot_fun, nsim = 2)


```

```{r, extract random effect variance without CI}

sur_revar <- as.data.frame(VarCorr(sur_bestfit))
sur_revar <- sur_revar %>%
  filter(is.na(var2)) %>%
  dplyr::select(var1,sdcor) %>% 
  rename(random_effect=var1, standard_deviation=sdcor)

gr_revar <- as.data.frame(VarCorr(gr_bestfit))
gr_revar <- gr_revar %>%
  filter(is.na(var2)) %>%
  dplyr::select(var1,sdcor) %>% 
  rename(random_effect=var1, standard_deviation=sdcor)
  
prep_revar <- as.data.frame(VarCorr(prep_bestfit))
prep_revar <- prep_revar %>%
  filter(is.na(var2)) %>%
  dplyr::select(var1,sdcor) %>% 
  rename(random_effect=var1, standard_deviation=sdcor)

revar_list <- list(
  survival = sur_revar,
  growth = gr_revar,
  Prob_reproduction = prep_revar)

revar_allVR <- bind_rows(revar_list, .id = "vital rate")
revar_allVR <- revar_allVR[which(!is.na(revar_allVR$random_effect)),]


plt1 <- ggplot(revar_allVR, aes(x=random_effect, y=standard_deviation, fill = `vital rate`)) +
  geom_col(position = position_dodge(width = 0.8)) +
   theme(axis.text.x = element_text(angle = 45, hjust = 1))

```


highest var random effect: 
sur: T_LA
gr: T_LA
prep: T_LH

## MerTools::Predict, graph VR against the highest variance random effect term and the highest fixed effect coefficient size term
```{r}
VFT_site <- c("B1", "B2", "CM", "CH", "IA", "ME", "GSP-BI", "GSP-LI")
```


#### sur ~ T_LA
```{r, make new data}
attr(terms(sur_bestfit), "term.labels")
data_new1 <- as.data.frame((rep(mean(sur_subset_R$s.logsize0, na.rm=TRUE), 50)))
names(data_new1)[1] <- "s.logsize0"
data_new1[c("s.P_RD", "s.P_RH", "s.sq.T_LH", "s.T_LD", "s.T_LH", "s.TBF", "s.TSF")] <- list(rep(mean(VFT_long$s.P_RD, na.rm=TRUE), 50), 
        rep(mean(VFT_long$s.P_RH, na.rm=TRUE), 50),
        rep(mean(VFT_long$s.sq.T_LH, na.rm=TRUE), 50), 
        rep(mean(VFT_long$s.T_LD, na.rm=TRUE), 50),
        rep(mean(VFT_long$s.T_LH, na.rm=TRUE), 50),
        rep(mean(VFT_long$s.TBF, na.rm=TRUE), 50),
        rep(mean(VFT_long$s.TSF, na.rm=TRUE), 50)
        )

mean_TA <- mean(VFT_long$T_LA, na.rm=TRUE)
sd_TA <- sd(VFT_long$T_LA, na.rm=TRUE)
new_T_LA <- ((seq(min(VFT_long$T_LA), max(VFT_long$T_LA)+2, length.out = 50))-mean_TA)/sd_TA
## upper bound of prediction set to be past max + 2, which is a rather extreme case
data_new1$s.T_LA <- new_T_LA      
data_new1$T_LA <- seq(min(VFT_long$T_LA), max(VFT_long$T_LA)+2, length.out = 50)

## replicate 8 times to store site effect
data_new1 <- bind_rows(replicate(8, data_new1, simplify = FALSE))
data_new1$site <- rep(VFT_site, each = 50)

data_new1$sur <- merTools::predictInterval(sur_bestfit, data_new1, type = "probability",level= 0.95,n.sims= 1000)

plt2 <- ggplot(data= data_new1, aes(x= T_LA, y= sur$fit)) +
  geom_line(aes(color = site)) + 
  # geom_ribbon(aes(ymin = data_new1$sur$lwr, ymax = data_new1$sur$upr, fill = site), alpha = 0.1) + 
  #ylim(0,1) +
   labs(y= "probability of survival", x = "annual mean soil temperature", size = 1.5) +
  theme_bw() + theme(legend.position='none') +
  theme(legend.position = "right") +
  theme(text = element_text(size = 13))                                                                                             
```

#### sur ~ T_LH (highest coef est)

```{r}
data_new1 <- as.data.frame((rep(mean(sur_subset_R$s.logsize0, na.rm=TRUE), 50)))
names(data_new1)[1] <- "s.logsize0"
data_new1[c("s.P_RD", "s.P_RH", "s.T_LD", "s.T_LA", "s.TBF", "s.TSF")] <- list(rep(mean(VFT_long$s.P_RD, na.rm=TRUE), 50), 
        rep(mean(VFT_long$s.P_RH, na.rm=TRUE), 50),
        rep(mean(VFT_long$s.T_LD, na.rm=TRUE), 50),
        rep(mean(VFT_long$s.T_LA, na.rm=TRUE), 50),
        rep(mean(VFT_long$s.TBF, na.rm=TRUE), 50),
        rep(mean(VFT_long$s.TSF, na.rm=TRUE), 50)
        )

mean_TLH <- mean(VFT_long$T_LH, na.rm=TRUE)
sd_TLH <- sd(VFT_long$T_LH, na.rm=TRUE)
mean_TLH2 <- mean(VFT_long$T_LH^2, na.rm=TRUE)
sd_TLH2 <- sd(VFT_long$T_LH^2, na.rm=TRUE)

new_T_LH <- seq(min(VFT_long$T_LH), max(VFT_long$T_LH)+2, length.out = 50)
## upper bound of prediction set to be past max + 2, which is a rather extreme case
data_new1$T_LH <- new_T_LH      
data_new1$s.T_LH <- (seq(min(VFT_long$T_LH), max(VFT_long$T_LH)+2, length.out=50)-mean_TLH)/sd_TLH

data_new1$s.sq.T_LH <- (seq(min(VFT_long$T_LH)^2, (max(VFT_long$T_LH)+2)^2, length.out=50)-mean_TLH2)/sd_TLH2


## replicate 8 times to store site effect
data_new1 <- bind_rows(replicate(8, data_new1, simplify = FALSE))
data_new1$site <- rep(VFT_site, each = 50)

data_new1$sur <- merTools::predictInterval(sur_bestfit, data_new1, type = "probability",level= 0.95,n.sims= 1000)

plt3 <- ggplot(data= data_new1, aes(x= T_LH, y= sur$fit)) +
  geom_line(aes(color = site)) + 
  # geom_ribbon(aes(ymin = data_new1$sur$lwr, ymax = data_new1$sur$upr, fill = site), alpha = 0.1) + 
  #ylim(0,1) +
   labs(y= "probability of survival", x = " hottest month soil temperature", size = 1.5) +
  theme_bw() + theme(legend.position='none') +
  theme(legend.position = "right") +
  theme(text = element_text(size = 13))    
```
#### gr ~ T_LA 

```{r}
data_new1 <- as.data.frame((rep(mean(gr_subset_R_2$s.logsize0, na.rm=TRUE), 50)))
names(data_new1)[1] <- "s.logsize0"
data_new1[c("s.P_RH", "s.T_LD", "s.TBF")] <- list(rep(mean(VFT_long$s.P_RH, na.rm=TRUE), 50), 
        rep(mean(VFT_long$s.T_LD, na.rm=TRUE), 50),
        rep(mean(VFT_long$s.TBF, na.rm=TRUE), 50)
        )
new_T_LA <- seq(min(VFT_long$T_LA), max(VFT_long$T_LA)+2, length.out = 50)
## upper bound of prediction set to be past max + 2, which is a rather extreme case
mean_T_LA2 <- mean(VFT_long$T_LA^2, na.rm=TRUE)
sd_T_LA2 <- sd(VFT_long$T_LA^2, na.rm=TRUE)

data_new1$T_LA <- new_T_LA      
data_new1$s.T_LA <- (seq(min(VFT_long$T_LA), max(VFT_long$T_LA)+2, length.out=50)-mean_TA)/sd_TA
data_new1$s.sq.T_LA <- (seq(min(VFT_long$T_LA)^2, (max(VFT_long$T_LA)+2)^2, length.out=50)-mean_T_LA2)/sd_T_LA2


## replicate 8 times to store site effect
data_new1 <- bind_rows(replicate(8, data_new1, simplify = FALSE))
data_new1$site <- rep(VFT_site, each = 50)

data_new1$gr <- merTools::predictInterval(gr_bestfit, data_new1, type = "linear.prediction",level= 0.95,n.sims= 1000)

plt4 <- ggplot(data= data_new1, aes(x= T_LA, y= gr$fit)) +
  geom_line(aes(color = site)) + 
  # geom_ribbon(aes(ymin = data_new1$sur$lwr, ymax = data_new1$sur$upr, fill = site), alpha = 0.1) + 
  #ylim(0,1) +
   labs(y= "growth", x = " mean annual soil temperature", size = 1.5) +
  theme_bw() + theme(legend.position='none') +
  theme(legend.position = "right") +
  theme(text = element_text(size = 13))   
```


#### prep ~ T_LH 

```{r}
data_new1 <- as.data.frame((rep(mean(prep_subset_R$s.logsize0, na.rm=TRUE), 50)))
names(data_new1)[1] <- "s.logsize0"
data_new1[c("s.P_RA", "s.P_RH", "s.sq.P_RA", "s.sq.T_LA","s.T_LA", "s.TBF", "s.TSF")] <- list(rep(mean(VFT_long$s.P_RA, na.rm=TRUE), 50), 
        rep(mean(VFT_long$s.P_RH, na.rm=TRUE), 50),
        rep(mean(VFT_long$s.sq.P_RA, na.rm=TRUE), 50),
        rep(mean(VFT_long$s.sq.T_LA, na.rm=TRUE), 50),
        rep(mean(VFT_long$s.T_LA, na.rm=TRUE), 50),
        rep(mean(VFT_long$s.TBF, na.rm=TRUE), 50),
        rep(mean(VFT_long$s.TSF, na.rm=TRUE), 50)
        )


## upper bound of prediction set to be past max + 2, which is a rather extreme case
data_new1$T_LH <- new_T_LH      
data_new1$s.T_LH <- (seq(min(VFT_long$T_LH), max(VFT_long$T_LH)+2, length.out=50)-mean_TLH)/sd_TLH

data_new1$s.sq.T_LH <- (seq(min(VFT_long$T_LH)^2, (max(VFT_long$T_LH)+2)^2, length.out=50)-mean_TLH2)/sd_TLH2


## replicate 8 times to store site effect
data_new1 <- bind_rows(replicate(8, data_new1, simplify = FALSE))
data_new1$site <- rep(VFT_site, each = 50)

data_new1$prep <- merTools::predictInterval(prep_bestfit, data_new1, type = "probability",level= 0.95,n.sims= 1000)

plt5 <- ggplot(data= data_new1, aes(x= T_LH, y= prep$fit)) +
  geom_line(aes(color = site)) + 
  # geom_ribbon(aes(ymin = data_new1$sur$lwr, ymax = data_new1$sur$upr, fill = site), alpha = 0.1) + 
  #ylim(0,1) +
   labs(y= "probability of reproduction", x = " hottest month soil temperature", size = 1.5) +
  theme_bw() + theme(legend.position='none') +
  theme(legend.position = "right") +
  theme(text = element_text(size = 13))  
```

#### prep ~ T_LA
```{r}
data_new1 <- as.data.frame((rep(mean(prep_subset_R$s.logsize0, na.rm=TRUE), 50)))
names(data_new1)[1] <- "s.logsize0"
data_new1[c("s.P_RA", "s.P_RH", "s.sq.P_RA","s.T_LH", "s.TBF", "s.TSF")] <- list(rep(mean(VFT_long$s.P_RA, na.rm=TRUE), 50), 
        rep(mean(VFT_long$s.P_RH, na.rm=TRUE), 50),
        rep(mean(VFT_long$s.sq.P_RA, na.rm=TRUE), 50),
        rep(mean(VFT_long$s.T_LH, na.rm=TRUE), 50),
        rep(mean(VFT_long$s.TBF, na.rm=TRUE), 50),
        rep(mean(VFT_long$s.TSF, na.rm=TRUE), 50)
        )


## upper bound of prediction set to be past max + 2, which is a rather extreme case
data_new1$T_LA <- new_T_LA      
data_new1$s.T_LA <- (seq(min(VFT_long$T_LA), max(VFT_long$T_LA)+2, length.out=50)-mean_TA)/sd_TA

data_new1$s.sq.T_LA <- (seq(min(VFT_long$T_LA)^2, (max(VFT_long$T_LA)+2)^2, length.out=50)-mean_T_LA2)/sd_T_LA2


## replicate 8 times to store site effect
data_new1 <- bind_rows(replicate(8, data_new1, simplify = FALSE))
data_new1$site <- rep(VFT_site, each = 50)

data_new1$prep <- merTools::predictInterval(prep_bestfit, data_new1, type = "probability",level= 0.95,n.sims= 1000)

plt6 <- ggplot(data= data_new1, aes(x= T_LA, y= prep$fit)) +
  geom_line(aes(color = site)) + 
  # geom_ribbon(aes(ymin = data_new1$sur$lwr, ymax = data_new1$sur$upr, fill = site), alpha = 0.1) + 
  #ylim(0,1) +
   labs(y= "probability of reproduction", x = " mean annual soil temperature", size = 1.5) +
  theme_bw() + theme(legend.position='none') +
  theme(legend.position = "right") +
  theme(text = element_text(size = 13)) 
```

### plot time series for only T_LA and T_LH

```{r}
ggplot(VFT_long,
         aes(x = as.factor(startyear),
             y = T_LA,
             colour = site,
             group = site)) +
    geom_line() +
    geom_point() +
    labs(y = "T_LA", x = "Year") +
    theme(
      axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1, size = 8)
    ) 
```

```{r}
ggplot(VFT_long,
         aes(x = as.factor(startyear),
             y = T_LH,
             colour = site,
             group = site)) +
    geom_line() +
    geom_point() +
    labs(y = "T_LH", x = "Year") +
    theme(
      axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1, size = 8)
    ) 
```



